### 2019-02-01 Low_latency_control/response/response_main.tex
Add files via upload

\documentclass[11pt]{article}
\usepackage{pgfplots,graphics, subfigure,vmargin, amsfonts, amsbsy, amsmath, color}
\usepackage{for_authors_reply}
\input{mysymbol.sty}

\setpapersize{USletter}
%\setmarginsrb{leftmargin}{topmargin}{rightmargin}{bottommargin}{headheight}{headsep}{footheight}{footskip}
 \setmarginsrb{1.0in}     {0.8in}    {1.0in}      {.8in}        {.2in}      {0.3in}   {.2in}       {0.3in}

\pagestyle{fancy}
\footrulewidth 0.5pt
\lfoot{\small \today}
\cfoot{}
\rfoot{\small \thepage}
\lhead{IEEE Internet of Things Journal {\bf Paper:OoT-5520-2018  } - Authors' Response}

\newcommand {\mytitledproof}[1] {\medskip\noindent{\bf Proof (#1): }}
\def         \myproof           {\noindent{\bf Proof: }}
\def \myQED       {\hfill\QED \\ \vspace{-12pt} \\ \noindent}
\def \myQEDnotext {\hfill\QED}
\def \notation    {\medskip\noindent{\bf Notation: }}
\def \SINR        {{\rm SINR}}
\newcommand \Eh[1]    {{\mathbb E}_{\bbh}\left[ #1 \right]}
\newcommand \EhBig[1] {{\mathbb E}_{\bbh}\Big[ #1 \Big]}
\newcommand \Eht[1]   {{\mathbb E}_{\bbh(t)}\left[ #1 \right]}
\newcommand \Ep[1]    {{\mathbb E}_{m(\bbp(\bbh))}\left( #1 \right)}
\newcommand \Ehp[1]   {{\mathbb E}_{\bbh,\bbp(\bbh)}\left[ #1 \right]}

\def \distlamt        {\big\|\bblam(t)     - \bblam^* \big\|^2}
\def \distlamtns      {\big\|\bblam(t)     - \bblam^* \big\|}
\def \distlamtpone    {\big\|\bblam(t+1)   - \bblam^* \big\|^2}

\newcommand \Cif[1] {C_{if}\big(#1\big)}
%\newcommand \inlineCif[1] {C_{if}\big(#1\big)}
\def \Cift {\Cif{h_{if}(t),p_{if}(t)}}
\def \Cifh {\Cif{h_{if},p_{if}(\bbh)}}
\def \Cifhtpnot {\Cif{h_{if}(t),p_{if}}}
%\def \inlineCift      {\Cif{h_{if}(t),p_{if}(t)}}
%\def \inlineCifh      {\Cif{h_{if},p_{if}(\bbh)}}
%\def \inlineCifhtpnot {\Cif{h_{if}(t),p_{if}}}

\newcommand{\e}{\mathtt{e}}
\newcommand{\R}{\mathbb{R}} 
\newcommand{\N}{\mathcal{N}}
\def\forall{\text{for all\ }}
\def\N{n}
\def \sech {\text{sech}}

\def \hbS {{\bf \hat{S}}}

\def \mph         {m\big(\bbp(\bbh)\big)}
\def \mphstar         {m^{*}\big(\bbp(\bbh)\big)}
\def \mpht        {m\big(\bbp(\bbh(t))\big)}
\def \moph        {m_{0}\big(\bbp(\bbh)\big)}
\def \fphh        {\bbf_{1}\big(\bbp(\bbh);\bbh\big)}
\def \fphhu       {\bbf_{1}\big(\bbp(\bbh(u));\bbh(u)\big)}
\def \fphht       {\bbf_{1}\big(\bbp(\bbh(t));\bbh(t)\big)}
\def \fptht       {\bbf_{1}\big(\bbp(t);\bbh(t)\big)}
\def \fpuhu       {\bbf_{1}\big(\bbp(u);\bbh(u)\big)}
\def \fphtlamtht  {\bbf_{1} \big(\bbp(\bbh(t),\bblam(t));\bbh(t) \big)}
\def \fpohh    {\bbf_{1}\big(\bbp(\bbh);\bbh\big)}
\def \lagrangian  {\ccalL[\bblam,\bbx,\mph]}
\def \lagrangianforreviewer  {\ccalL[\bblam_{2},\bbx]}
\def \lagrangiansubzero  {\ccalL_{0}\left[\bblam,\bbx,\mph\right]}
\def \lagrangiansubzerot  {\ccalL_{0}\left[\bblam(t),\bbx,\mph\right]}
\def \lagrangiant {\ccalL\left[\bblam(t),\bbx, \mph\right]}
\def \dest {{\rm dest}}

\def\E{\mathbb{E}}

\begin{document}

\title{Control Aware Radio Resource Allocation in Low Latency Wireless Control Systems}
\author{Mark Eisen$^*$ \quad Mohammad M. Rashid$^\dagger$ \quad Konstantinos Gatsis$^*$ \\ Dave Cavalcanti$^\dagger$\quad Nageen Himayat$^{\dagger}$ \quad Alejandro Ribeiro$^*$ \\ $^{*}$University of Pennsylvania \quad $^{\dagger}$Intel Corporation} \maketitle                                    %%%%

\editor

\red{
We thank the associate editor and the reviewers for the time and effort invested in providing constructive suggestions. The requests most common among the reviewers concerned numerous points of clarification that add to the understandability and readability of the manuscript. Two reviewers raised questions and comments regarding the derivation of the problem in Example 1. Reviewers also raised key questions regarding some assumptions made in the analysis.  A more detailed account of the major changes implemented in the manuscript follows.

\begin{itemize}

\item[{\bf[C1]}] Reviewers 1 and 3 expressed concerns about the derivation of Example 1 and the stability analysis in Section V-C. We have thus expanded upon Example 1 that makes the derivations more explicit. We have revised Section V-C to give further details on the stability claim. We have also identified an omission in the previous version that has been corrected. More specifically, assuming we have a statistically accurate dual variable solution, using Proposition 2 and Theorem 2 we argue that the actual packet success rate selected by the algorithm can be made sufficiently close to the optimal one. Since the latter leads to system stability then the packet success rate selected by our algorithm also leads to stability. We thank the reviewers for these comments.

\item[{\bf[C2]}] Reviewer 1 has suggested some additional numerical experiments to explore the effects of different parameters in the performance. Per the reviewer's suggestion, we have added Figure 4 to the paper, which compares both the suboptimality and constraint violation of the resource allocations generated by Newton's method over a non-stationary channel using two different accuracies $\hat{V}$. The results demonstrate an interesting effect that can occur if the choice of $\hat{V}$ is too strict. We thank the reviewer for this suggestion.

\item[{\bf[C3]}] Reviewers raised numerous questions regarding some of the practical considerations in Section VI, such as backtracking parameter selection as well as satisfying the non-stationary assumption in Assumption 4. To address these concerns, we have added additional details to Section VI to clarify many of these points, and an additional remark after Assumption 4. We thank the reviewers for raising these questions.

\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

\red{We thank the reviewer for the comments, feedback, and questions. The reviewer raised thoughtful questions about the modeling of the problem and some of the assumptions made in the paper. The reviewer also suggests to make the numerical results more thorough. Each concern is addressed individually below:}

\issue{ The authors need to recast the framework within ultra-reliable and low-latency control and communication.}
\answer{\red{To be written}}

\issue{ It was unclear whether the focus is on communication for a given control or the other way around. Taht is currently it looks like the solution is based on sequential design.}
\answer{\red{To be written}}

\issue{ The article would improve a lot if specific use cases are defined.}
\answer{\red{To be written}}

\issue{ Linear control was adopted (for tractability) but clearly the nonlinearity is needed}
\answer{\red{To be written}}

\issue{Reliability as not addressed however it is intertwined with latency}
\answer{\red{To be written}}

\issue{Authors did not cite recent and fundamental articles within URLLC:}
\answer{\red{To be written}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

\red{We thank the reviewer for the comments, feedback, and questions. The reviewer raised thoughtful questions about the modeling of the problem and some of the assumptions made in the paper. The reviewer also suggests to make the numerical results more thorough. Each concern is addressed individually below:}

\issue{ Currently, it is not easy to follow the abstract. It should be revised. The contributions of the this paper should be highlighted.}
\answer{\red{To be written}}

\issue{  In simulation part, it would be great if the authors could compare the proposed CALLS method with other methods if possible, since there are also some papers discussing communication resource allocation for multiple control loops.}
\answer{\red{To be written}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

\red{We thank the reviewer for the comments, feedback, and questions. The reviewer raised thoughtful questions about the modeling of the problem and some of the assumptions made in the paper. The reviewer also suggests to make the numerical results more thorough. Each concern is addressed individually below:}

\issue{ Why do you only consider the AWGN channels?  How will your scheme performance when considering fading channels? Is the fading channel also a typical scenario for WiFi based in-door wireless control system?  }
\answer{\red{To be written}}

\issue{In your system, how do you obtain accurate channel state information? E.g. for AWGN channels, you need to test the noise variance. Is your scheme sensitive to channel estimation errors?}
\answer{\red{To be written}}

\issue{The proposed system shall be compared to more existed typical scheduling scheme employed in control system.}
\answer{\red{To be written}}

\end{document}

### 2019-02-01 Low_latency_control/arxiv/problem_formulation.tex
Add files via upload

%%%%%%%%%%%%%%%% F I G U R E %%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\input{tikz_wncs4}
\caption{Wireless control system with $m$ independent systems. Each system contains a sensor that measure state information, which is transmitted to the controller over a wireless channel. The state information is used by the controller to determine control policies for each of the systems.The communication is assumed to be wireless in the uplink and ideal in the downlink.}
\label{fig_wcs}
\end{figure}

Consider a system of $m$ independent linear control systems, or devices, where each system $i=1,\hdots,m$ maintains a state variable $\bbx_{i} \in \reals^p$. The dynamics are discretized so that the state evolves over time index $k$.  Applying an input $\bbu_{i,k} \in \reals^q$ causes the state and output to evolve based on the discrete-time  state space equations,
%
\begin{align}\label{eq_control_orig}
\bbx_{i,k+1} &= \bbA_{i} \bbx_{i,k} + \bbB_{i} \bbu_{i,k} + \bbw_k
\end{align}
%
where $\bbA_{i} \in \reals^{p \times p}$ and $\bbB_{i} \in \reals^{p \times q}$ are matrices that define the system dynamics, and $\bbw_{k} \in \reals^{p}$ is Gaussian noise with co-variance $\bbW_i$ that captures the errors in the linear model (due to, e.g., unknown dynamics or from linearizion of non-linear dynamics). We further assume the state transition matrix $\bbA_i$ is on its own unstable, i.e. has at least one eigenvalue greater than 1. This is to say that, without an input, the dynamics will drive the state $\bbx_{i,k} \rightarrow \infty$ as $k \rightarrow \infty$.

In the wireless control system model is presented in Figure \ref{fig_wcs}. Each system is closed over a wireless medium, over which the sensor located at the control system sends state information to the controller located at a wireless access point (AP) shared among all systems. Using the state information $\bbx_{i,k}$ received from device $i$ at time $k$, the controller determines the input $ \bbu_{i,k}$ to be applied. We stress in Figure \ref{fig_wcs} we restrict our attention to the wireless communications at the sensing, or ``uplink'', while the control actuation, or ``downlink, is assumed to occur over an ideal channel. We point out that while a more complete model may include packet drops in the downlink, in practice the more significant latency overhead occurs in the uplink. We therefore keep this simpler model for mathematical coherence. In low-latency applications, a high state sampling rate is required be able to adapt to the fast-moving dynamics  This subsequently places a tight restriction on the latency in the wireless transmission, so as to avoid losing sampled state information. This specific latency requirement between the sensor and AP we denote by $\tau_{\max}$, and is often considered to be in the order of milliseconds.

Because the control loop in Figure \ref{fig_wcs} is closed over a wireless channel, there exists a possibility at each cycle $k$ that the transmission fails and state information is not received by the controller. We refer to this as the ``open-loop' configuration; when state information is received, the system operates in ``closed-loop.'' As such, it is necessary to define the system dynamics in both configurations. Consider a generic linear control, in which the input being determined as $ \bbu_{i,k} = \bbK_i \bbx_{i,k}$ for some matrix $\bbK_i \in \reals^{q \times p}$. Many common control policies indeed can be formulated in such a manner, such as LQR control. In general, this matrix $\bbK$ is chosen such as that the closed loop dynamic matrix $\bbA + \bbB \bbK$ is stable, i.e. has all eigenvalues less that 1. Thus, application of this control over time will drive the state $\bbx_{i,k} \rightarrow 0$ as $k \rightarrow \infty$. As the controller does not always have access to state information, we alternatively consider the estimate of state information of device $i$ known to the controller at time $k$ as
%
\begin{align}\label{eq_state_est}
\hbx^{(l_i)}_{i,k} :=(\bbA_i + \bbB_i\bbK_i)^{l_i} \bbx_{i,k-l_i},
\end{align}
%
where $k-l_i \geq k-1$ is the last time instance in which control system $i$ was closed. There are two important things to note in \eqref{eq_state_est}. First, this is the estimated state \emph{before} a transmission has been attempted at time $k$; hence, $l_i = 1$ when state information was received at the previous time. Second, observe that in \eqref{eq_state_est} we assume that the AP/controller has knowledge of the dynamics $\bbA_i$ and $\bbB_i$, as well as the linear control matrix $\bbK_i$. Any gap in this knowledge of dynamics is captured in the noise $\bbw_k$ in the actual dynamics in \eqref{eq_control_orig}. Note that the estimated state \eqref{eq_state_est} is used in place of the true state in both the determination of the control \emph{and} the radio resource allocation decisions as discussed later in this paper.

At time $k$, if the state information is received, the controller can apply the input $\bbu_{i,k} = \bbK_i \bbx_{i,k}$ exactly, otherwise it applies an input using the estimated state, i.e. $\bbu_{i,k} = \bbK_i \hbx_{i,k}$. Thus, in place of \eqref{eq_control_orig}, we obtain the following switched system dynamics for $\bbx_{i,k}$ as
%
\begin{align}\label{eq_control_switch}
\bbx_{i,k+1} &= \begin{cases}
(\bbA_i + \bbB_i \bbK_i) \bbx_{i,k} + \bbw_k, \ \text{in closed-loop}, \\
\bbA_i \bbx_{i,k} + \bbB_i\bbK_i\hbx^{(l_i)}_{i,k} + \bbw_k, \ \text{in open-loop}.
\end{cases}
\end{align}
%
The transmission counter $l_i$ is updated at time $k$ as
%
\begin{align}
l_i &\leftarrow \begin{cases}
1, \ \text{in closed-loop}, \\
l_1 + 1, \ \text{in open-loop}.
\end{cases} \label{eq_time_switch}
\end{align}
%
Observe in \eqref{eq_control_switch} that, when the system operates in open loop, the control is not applied relative to the current state $\bbx_{i,t}$ but on the estimated state $\hbx^{(l_i)}_{i,k}$, which indeed may not be close to the true state. In this case, the state may not be driven to zero as in the closed-loop configuration. To see the effect of operating in open loop for many successive iterations, we can write the error between the true and estimated state as
%
\begin{align}\label{eq_diff}
\bbe_{i,k} := \bbx_{i,k} - \hbx^{(l_i)}_{i,k} = \sum_{j=0}^{l_i-1}\bbA_i^{j} \bbw_{i,k-j-1}.
\end{align}
%
In \eqref{eq_diff}, it can be seen that as $l_i$ grows, the error $\bbe_{i,k}$ grows with the accumulation of the noise present in the actual state but not considered in the estimated state. Thus, if $l_i$ is large and $\bbw_{i,k}$ is large (i.e., high variance), this error will become large as well.

To conclude the development of the wireless control formulation, we define a quadratic Lyapunov function $L(\bbx) := \bbx^T \bbP \bbx$ for some positive definite $\bbP \in \reals^{p \times p}$ that measures the performance of the system as a function of the state. Because the scheduler only has access to estimated state info, we consider the expected value of the Lagrangian given the state estimate, which can be found via \eqref{eq_diff} as
%
\begin{align}
\E [L(\bbx_{i,k}) \mid &\hbx^{(l_i)}_{i,k}]  \label{eq_perf} \\
%&=\E \left[ \left(\hbx^{(l_i)}_{i,k} +\sum_{j=0}^{l-1}\bbA^{i} \bbw_{i,k-j-1} \right)^T  \left(\hbx^{(l_i)}_{i,k} + \sum_{i=0}^{l-1}\bbA_i^{j} \bbw_{i,k-j-1}\right) \right] \nonumber \\
&=  (\hbx^{(l_i)}_{i,k})^T \bbP (\hbx^{(l_i)}_{i,k}) + \sum_{j=0}^{l_i-1} \Tr[(\bbA_i^T\bbP^{\frac{1}{j}}\bbA_i)^{j} \bbW_i]. \nonumber
\end{align}
%
Thus, the control-specific goal is to keep $\E [L(\bbx_{i,k}) \mid \hbx^{(l_i)}_{i,k}]$ within acceptable bounds for each system $i$. We now proceed to discuss the wireless communication model that determines the resource allocations necessary to close the loop.

\subsection{IEEE 802.11ax communication model}\label{sec_comm_model}

%%%%%%% F I G U R E %%%%%%%%%%%
\begin{figure}
%\centering
%\input{multiplex.tex}
\includegraphics[width=0.5\textwidth]{schedule2.png}
\caption{Multiplexing of frequencies (RU) and time (PPDU) in IEEE 802.11ax transmission window (formally referred as Transmission Opportunity or TXOP in the standard. The total transmission time is the time of all PPDUs, including the overhead of trigger frames (TF) and acknowledgments.}
\label{fig_multiplex}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%

We consider the communication model provided in the next-generation Wi-Fi standard IEEE 802.11ax. While 3GPP wireless systems such as LTE~\cite{sesia2011lte} or the next generation 5G~\cite{agiwal2016next} can also be considered as alternate communication models, most factory floors are already equipped with Wi-Fi connectivity and, moreover, Wi-Fi can operate in the unlicensed band. It is generally considered to be cost-effective to operate and maintain.

Traditional Wi-Fi systems rely only on contention-based channel access and may introduce high or variable latency in congested or dense deployment scenarios even in a fully managed Wi-Fi network, which is typically available in industrial control and automation scenarios. To address the problems with dense deployment, the draft 802.11ax amendment has defined scheduling capability for Wi-Fi access points (APs). Wi-Fi devices can now be scheduled for accessing the channel in addition to the traditional contention-based channel access. Such scheduled access enables more controlled and deterministic behavior in the Wi-Fi networks. Within each transmission window (formally referred as transmission opportunity or TXOP in the standard), the AP may schedule devices through both frequency and time division multiplexing using the multi-user (MU) OFDMA technique. This to say that devices can be slotted in various frequency bands---formally called resource units (RUs)---and in different timed transmission slots---formally called PPDUs. An example of the multiplexing of devices across time and frequency is demonstrated in Figure \ref{fig_multiplex}. The AP additionally sends a trigger frame (TF) indicating which devices should transmit data in the current TXOP and the time/frequency resources these triggered devices should use in their transmissions.

To state this model formally, the scheduling parameters assigned by the AP to each device consist of a frequency-slotted RU, time-slotted PPDU, and an associated modulation and coding scheme (MCS) to determine the transmission format. The transmission power is assumed to be fixed and equally divided amongst all devices. We define the following notations to formulate these parameters.  To specify an RU, we first notate by $f^1, f^2, \hdots, f^b$, where $n$ is the number of discrete frequency bands of fixed bandwidth (typically 2 MHz) in which a device can transmit; in a 20MHz channel, for example, there are $n=10$ such bands. For each device, we then define a set of binary variables $\varsigma^j_{i} \in \{0,1\}$ if device $i$ transmits in band $f^j$ and collect all such variables for device $i$ in $\bbsigma_i = [\varsigma^1_{i};\hdots;\varsigma^b_{i}] \in \{0,1\}^{b}$ and for all devices in  $\bbSigma := [\bbsigma_i, \hdots, \bbsigma_m] \in \{0,1\}^{b \times m}$. A device may transmit in bands in certain multiples of 2MHz as well, which would be notated as, e.g. $\bbsigma_i = [1; 1; 0; \hdots; 0]$ for transmission in an RU of size 4MHz. Note, however, that allowable RU's contain only sizes of certain multiples of 2MHz---namely, 2MHz, 4MHz, 8MHz, and 20MHz in the 802.11ax standard. Furthermore, it is only permissible to transmit in \emph{adjacent} bands, e.g. $f^{j}$ and $f^{j+1}$. We therefore define the set $\ccalS \subset \{0,1\}^{b}$ as the set of binary vectors that define permissible RUs and consider only $\bbsigma_i \in \ccalS$ for all devices $i$.  Finally, note that the RU assignment $\bb0 \in \ccalS$ signifies a device does not transmit in this particular transmission window.

%%%%%%% T A B L E %%%%%%%%%%%
\begin{table}
\input{mcs.tex}
\caption{Data rates for MCS configurations in IEEE 802.11ax for 20MHz channel. The modulation type and coding rate in the first 2 columns together specify a PDR function $q(\bbmu,\bbsigma)$ for RU $\bbsigma$. The data rate in the third column specifies the associated transmission time $\tau(\mu, \bbsigma)$. }
\label{tab_mcs}
\end{table}
%%%%%%%%%%%%%%%%%%%%%

To specify the PPDU of all scheduled devices, we define for device $i$ a positive integer value $\alpha_i \in \mathbb{Z}_{++}$ that denotes the PPDU slot in which it transmits and collect such variables for all devices in $\bbalpha = [\alpha_1; \hdots; \alpha_m] \in \mathbb{Z}_{++}^m$. Likewise, device $i$ is given an MCS $\mu_i$ from the discrete space $\ccalM = \{0,1,2,\hdots,10\}$. The MCS in particular defines a pair of modulation scheme and coding rate that subsequently determine both the data rate and packet error rate of the transmission. The allowable MCS settings provided in 802.11ax are provided in Table \ref{tab_mcs}. Finally, we notate by $\bbh_{i} := [h_i^1; h_i^2; \hdots; h_i^b] \in \reals^b_+$ a set of channel states experienced by device $i$, where $h_i^j$ is the gain of a wireless fading channel in frequency band $f^j$. We assume that channel conditions are constant within a single TXOP, i.e. do not vary across PPDUs.

We now proceed to define two functions that describe the wireless communications over the channel. Firstly, we define a function  $q: \reals_+^b \times \ccalM \times \ccalS \rightarrow [0,1]$ which, given a set of channel conditions $\bbh$, MCS $\mu$, and RU $\bbsigma$, returns the probability of successful transmission, otherwise called packet delivery rate (PDR). Furthermore, define by $\tau: \ccalM \times \ccalS  \rightarrow \reals_+$ a function that, given an MCS $\mu$ and RU $\bbsigma$, returns the maximum time taken for a single transmission attempt. Assuming a fixed packet size, such a function can be determined from the data rates associated with each MCS in Table \ref{tab_mcs}.  Observe that all functions just defined are determined independent of the PPDU slot the transmission takes place in, while transmission time is also independent of the channel state. Because a PPDU cannot finish until all transmissions within the PPDU have been completed, the total transmission time of a single PPDU $s$ is the maximum transmission time taken by all devices within that time slot. We define the transmission time of PPDU slot $s$ as
%
\begin{equation}\label{eq_time_slot}
\hat{\tau}(\bbSigma, \bbmu, \bbalpha, s) := \max_{i: \alpha_i = s} \tau(\mu_i, \bbsigma_i) + \tau_0(\bbalpha,s),
\end{equation}
%
where $\tau_0: \mathbb{Z}_{++}^m \times \mathbb{Z}_{++} \rightarrow \reals_+$ is a function that specifies the communication overhead of PPDU $s$. This overhead may consist of, e.g., the time required to send TFs to scheduled users, as seen in Figure \ref{fig_multiplex}.

%\begin{remark}\normalfont
%Note that, while in the next sections we proceed to derive the scheduling problem and method specifically for the IEEE 802.11ax standard, the concepts we discuss and overall approach can be applied generically to any communication standard that permits centralized resource allocation and scheduling, e.g., LTE. We discuss these details as they pertain to IEEE 802.11ax primarily for its applicability in industrial control settings, but a corresponding scheduling protocol for other standards can be immediately obtained by substituting the above scheduling parameters, e.g. RUs, MCS, etc., with the available parameters for the communication standard of interest.
%\end{remark}

### 2019-02-01 Low_latency_control/arxiv/root.tex
Add files via upload

%\documentclass[12pt, draftclsnofoot, onecolumn]{IEEEtran}
\documentclass[journal]{IEEEtran}
\usepackage{etex}
\reserveinserts{100}
%\usepackage{spconf}
\usepackage{mathtools}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphics, theorem, times, amsfonts, graphicx, amssymb, cite}
\usepackage[outdir=./]{epstopdf}
\usepackage{tikz}
%\usepackage{fix2col}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{matrix} % for block alignment
\usetikzlibrary{decorations.markings} % for arrow heads
\usetikzlibrary{calc} % for manimulation of coordinates
\usetikzlibrary{decorations.pathmorphing,snakes, positioning, decorations.pathreplacing}

\usetikzlibrary{fit}
%\input{figures/tikh_styles}
\usepackage{pgfplots}
\usepackage{color}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{comment}
%\usepackage{flushend}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[affil-it]{authblk}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumerate}
\usepackage{morefloats}
\usepackage{setspace}

\input{mysymbol.sty}
\input{my_sections}
%\input{TikZGallery-master/Sty/TikZGallery-Preamble}
%\renewcommand{\baselinestretch}{.995}
\newcommand{\QED}{\hfill\ensuremath{\blacksquare}}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}

\def \vneg {\vspace{-2cm}}

\def\whp{\text{w.h.p}}
\def\E{\mathbb{E}}
\def\P{\mathbb{P}}
\def\supp{\text{support}}
\def\Tr{\text{Tr}}

\def\keywords{
{\bfseries\textit{Index Terms}---\,\relax%
}}
\def\endkeywords{\par}

%\addtolength{\textwidth}{10mm}
%\addtolength{\evensidemargin}{-5mm}
%\addtolength{\oddsidemargin}{-5mm}
%\addtolength{\textheight}{10mm}
%\addtolength{\topmargin}{-5mm}

%\renewcommand \blue[1]{}  \renewcommand \red[1]{}

\newtheorem{assumption}{\hspace{0pt}\bf Assumption}
\newtheorem{lemma}{\hspace{0pt}\bf Lemma}
\newtheorem{proposition}{\hspace{0pt}\bf Proposition}
\newtheorem{example}{\hspace{0pt}\bf Example}
\newtheorem{observation}{\hspace{0pt}\bf Observation}
\newtheorem{theorem}{\hspace{0pt}\bf Theorem}
\newtheorem{corollary}{\hspace{0pt}\bf Corollary}
\newtheorem{fact}{\hspace{0pt}\bf Fact}
\newtheorem{remark}{\hspace{0pt}\bf Remark}
\newtheorem{test}{\hspace{0pt}\it Test Case}
\newtheorem{definition}{\hspace{0pt}\bf Definition}
\newtheorem{conj}{\hspace{0pt}\bf Conjecture}
\newtheorem{method}{\hspace{0pt}\bf Method}

% Title.
% ------

\title{Control Aware Radio Resource Allocation in Low Latency Wireless Control Systems}
%
% Single address.
% ---------------

\author{Mark Eisen$^*$ \quad Mohammad M. Rashid$^\dagger$ \quad Konstantinos Gatsis$^*$ \\ \textup{Dave Cavalcanti$^\dagger$\quad Nageen Himayat$^{\dagger}$ \quad Alejandro Ribeiro$^*$}
\thanks{Supported by the Intel Science and Technology Center for Wireless Autonomous Systems and ARL DCIST CRA W911NF-17-2-0181. The authors are with the ($^*$)Department of Electrical and Systems Engineering, University of Pennsylvania and ($^\dagger$)Wireless Communications Research, Intel Corporation. Email: maeisen@seas.upenn.edu, mamun.rashid@intel.com, kgatsis@seas.upenn.edu, dave.cavalcanti@intel.com, nageen.himayat@intel.com, aribeiro@seas.upenn.edu.}}

\begin{document}

\thispagestyle{empty}
\maketitle

%
\begin{abstract}
%We consider the problem of allocating radio resources, such as frequency, bandwidth, data rates and time slots, over wireless communication links to control a series of independent wireless control systems. Low-latency transmissions are necessary in control systems that contain high sampling rates, thus necessitating fast transmissions. Compared to wired links, wireless links are highly susceptible to communication errors arising from channel noise and interference. Achieving fast data rates over wireless links thus comes at the cost of reliability in the form of high packet error rates. However, the effect of the communication link errors on the control system performance is not static but depends dynamically on the control system state. Traditional resource allocation methods over wireless links are unaware of control system states and thus are designed with static link error performance, which could be very inefficient if  error performance assumption is too conservative (e.g. too low error rate target), or unacceptably unreliable if the assumption is too relaxed. In this paper, we propose a novel control-communication co-design approach to this low-latency resource allocation problem. We develop methods that incorporate control state information to make scheduling decisions over time on frequency, bandwidth and data rates across the next-generation Wi-Fi based wireless communication links connecting the different components of the control systems. Control systems that are closer to instability or further from a desired range in a given control cycle are given priority in the scheduling. Rather than a simple priority ranking, we derive precise packet error rate targets for each system needed to satisfy respective Lyapunov control targets and make scheduling decisions to best meet targets while minimizing total transmission time. The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.
We consider the problem of allocating radio resources over wireless communication links to control a series of independent wireless control systems. Low-latency transmissions are necessary in enabling time-sensitive control systems to operate over wireless links with high reliability. Achieving fast data rates over wireless links thus comes at the cost of reliability in the form of high packet error rates compared to wired links due to channel noise and interference.  However, the effect of the communication link errors on the control system performance depends dynamically on the control system state. We propose a novel control-communication co-design approach to the low-latency resource allocation problem. We incorporate control and channel state information to make scheduling decisions over time on frequency, bandwidth and data rates across the next-generation Wi-Fi based wireless communication links that close the control loops.  Control systems that are closer to instability or further from a desired range in a given control cycle are given higher packet delivery rate targets to meet. Rather than a simple priority ranking, we derive precise packet error rate targets for each system needed to satisfy stability targets and make scheduling decisions to meet such targets while reducing total transmission time. The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.

\end{abstract}
%

\begin{keywords}
wireless control, low-latency, codesign, IEEE 802.11ax
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% S E C T I O N %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec_intro}
\input{introduction.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% S E C T I O N %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Wireless Control Sysyem}\label{sec_problem_formulation}
\input{problem_formulation.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% S E C T I O N %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Optimal Control Aware Scheduling}\label{sec_optimal}
\input{optimal_scheduling.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% S E C T I O N %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Control-Aware Low-Latency Scheduling (CALLS)}\label{sec_calls}
\input{scheduling_alg.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% S E C T I O N %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation Results}\label{sec_simulation}
\input{simulations.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% S E C T I O N %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Conclusions}\label{sec_conclusion}
\input{conclusions.tex}

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\urlstyle{same}
\bibliographystyle{IEEEtran}
\bibliography{wireless_ll_control,scheduling_control}

\end{document}

### 2019-02-01 Low_latency_control/arxiv/simulations.tex
Add files via upload

In this section, we simulate the implementation of both the control-aware CALLS method and a standard ``control-agnostic'' scheduling methods for various low-latency control systems over a simulated wireless channel. We point out the low-latency based scheduling/assignment approaches of both methods being compared are identical, with the distinguishing features being the dynamic control-aware packet delivery rates incorporated in the CALLS method. In doing so, we may analyze the performance of the control-aware design outlined in the previous section relative to a standard latency-aware approach in terms of, e.g., number of users supported with fixed latency threshold or best latency achieved with fixed number of users. As we are interested primarily in low latency settings that tightly restrict the communication resources, we consider two standard control systems whose rapidly changing state requires high sampling rates, and consequently a communication latency on the order of milliseconds. The parameters for the simulation setup are provided in Table \ref{tab_simulation}. The packet delivery rate function $q(\bbh,\mu,\bbsigma)$ is computed using the standard AWGN noise curves for wireless channels. The transmission time $\tau(\mu,\bbsigma)$ is computed in the simulations using the associated data rates of an MCS in Table \ref{tab_mcs} for a 100 byte packet and overhead (e.g. TFs) of the 802.11ax specifications. The latency overhead for this setting amounts to approximately $\tau_0 \approx 100 \mu$s.
%
\begin{table}
\begin{tabular}{ l | l } \hline
  Channel model & IEEE Model E (indoor) \cite{liu2014ieee} \\
  Sensor to AP distances & Random (1 to 50 meters)\\
  Transmit power & 23 dbm \\
  Channel bandwidth & 20 MHz \\
  RU sizes & 2, 4, 8, 20 MHz \\
  \# of antennas at AP & 2 \\
  \# of antennas at sensors & 1 \\
  MCS options & See Table \ref{tab_mcs} \\
  State sampling period & 10 ms \\
  \hline
\end{tabular}
\caption{Simulation setting parameters.}
\label{tab_simulation}
\end{table}
%

\subsection{Inverted pendulum system}

\begin{figure}
%\input{pendulum.tex}
\centering
\includegraphics[height=.2\textheight]{pendulum.PNG}
\caption{Inverted pendulum-cart system $i$. The state $\bbx_{i,k} = [x_{i,k}, \dot{x}_{i,k}, \theta_{i,k}, \dot{\theta}_{i,k}]$ contains specifies angle $ \theta_{i,k}$ of the pendulum to the vertical, while the input $u_{i,k}$ reflects a horizontal force on the cart.}
\label{fig_inverted_pendulum}
\end{figure}

We perform an initial set of simulations on the well-studied problem of controlling a series of inverted pendulums on a horizontal cart. While conceptually simple, the highly unstable dynamics of the inverted pendulum make it a representative example of control system that requires fast control cycles, and subsequently low-latency communications when being controlled over a wireless medium. Consider a series of $m$ identical inverted pendulums, as pictured in Figure \ref{fig_inverted_pendulum}. Each pendulum of length $L$ is attached at one end to a cart that can move along a single, horizontal axis. The position of the pendulum changes by the effects of gravity and the force applied to the linear cart. For our experiments, we use the modeling of the inverted pendulum as provided by Quanser \cite{quanser}. The state is $p=4$ dimensional vector that maintains the position and velocity of the cart along the horizontal axis, and the angular position and velocity of the pendulum, i.e. $\bbx_{i,k} := [x_{i,k}, \dot{x}_{i,k}, \theta_{i,k}, \dot{\theta}_{i,k}]$. The system input $u_{i,k}$ reflects a horizontal force placed upon $i$th pendulum. By applying a zeroth order hold on the continuous dynamics with a state sampling rate of $0.01$ seconds and linearizing, we obtained the following discrete linear dynamic matrices of the pendulum system
%
\begin{align}\label{eq_control_orig}
\bbA_i =
\begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & 2.055 & -0.722 & 4.828 \\
0 & 0.023 & 0.91 & 0.037 \\
0 & 0.677 & -0.453 & 2.055
\end{bmatrix},
\bbB_i =
\begin{bmatrix}
0.034 \\ 0.168 \\ 0.019 \\ 0.105
\end{bmatrix}.
\end{align}
%

Because the state $\bbx_{i,k}$ measures the angle of the $i$th pendulum at time $k$, the goal is to keep this close to zero, signifying that the pendulum remains upright. The input matrix $\bbK$ is computed to be a standard LQR-controller.

We perform a set of simulations scheduling the transmissions to control a series of inverted pendulums, varying both the latency threshold $\tau_{\max}$ and number of devices $m$. We perform the scheduling using the proposed CALLS method for control-aware low latency scheduling an, as a point of comparison, consider scheduling using a fixed ``high-reliability'' PDR of $0.99$ for all devices. Each simulation is run for a total of $1000$ seconds and is deemed ``successful'' if all pendulums remain upright for the entire run. We perform 100 such simulations for each combination of latency threshold and number of devices to determine how many devices we can support at each latency threshold using both the CALLS and fixed-PDR methods for scheduling.

In Figure \ref{fig_ip_dist} we show the results of a representative simulation of the control of $m=25$ pendulum systems with a latency bound of $\tau_{\max} = 10^{-3}$ seconds. In both graphs we show the average distance from the center vertical of each pendulum over the course of 1000 seconds. In the top figure, we see by using the control-aware CALLS method we are able to keep each of the 25 pendulums close to the vertical for the whole simulation. Meanwhile, using the standard fixed PDR, we are unable to meet the scheduling limitations imposed by the latency threshold, and many of the pendulums swing are unable to be kept upright, as signified by the large deviations from the origin. This is due to the fact that certain pendulums were not scheduled when most critical, and they subsequently became unstable.

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{ip_dist2.pdf}
\caption{Average pendulum distance to center vertical for $m=25$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max}= 1$ ms latency threshold. The proposed control aware scheme keeps all pendulums close to the vertical, while fixed-PDR scheduling cannot.}
\label{fig_ip_dist}
\end{figure}
%%

We present in Figure \ref{fig_ip_bar} the final capacity results obtained over all the simulations. We say that a scheduling method was able to successfully serve $m'$ devices if it keeps all devices within a  $|\theta_{i,k}| \leq 0.05$ error region for 100 independent simulations. Observe that the proposed approach is able to increase the number of devices supported in each case, with up to $1.5$ factor increase over the standard fixed PDR approach. Indeed, the proposed CALLS method is able to allocate the available resource in a more principled manner, which allows for the support of more devices simultaneously being controlled.

%%%%%%
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    ybar,
    enlargelimits=0.15,
    legend style={at={(0.5,0.85)},
      anchor=north,legend columns=-1},
    ylabel={\# devices supported},
    symbolic x coords={0.5 ms,1 ms,1.5 ms},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot coordinates {(0.5 ms,9) (1 ms,19) (1.5 ms,36)};
\addplot coordinates {(0.5 ms,12) (1 ms,28) (1.5 ms,55)};
\legend{Fixed PDR, CALLS}
\end{axis}
\end{tikzpicture}
\caption{Total number of inverted pendulum devices that can be controlled using Fixed-PDR and CALLS scheduling for various latency thresholds.}
\label{fig_ip_bar}
\end{figure}
%%%%%%

\subsection{Balancing board ball system}

We perform another series of experiments on the wireless control of a series of balancing board ball systems developed by Acrome \cite{acrome}. In such a system, a ball is kept on a rectangular board with a single point of stability in the center of the board. Two servo motors underneath the board are used to push the board in the horizontal and vertical directions, with the objective to keep the ball close to the center of the board. The state here reflects the position and velocity in the horizontal and vertical axes, i.e. $\bbx_{i,k} := [x_{i,k}, \dot{x}_{i,k}, y_{i,k}, \dot{y}_{i,k}]$.  The input $\bbu_{i,k} = [v_x, v_y]$ reflects the voltage applied to the horizontal and vertical motors. As before, we apply a zeroth order hold on the continuous dynamics with a state sampling rate of $0.01$ seconds and linearize, thus obtaining the following dynamic system matrices,
%

\begin{align}\label{eq_control_orig}
\bbA_i =
\begin{bmatrix}
1 & 0.01 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0.01 \\
0 & 0 & 0 & 1
\end{bmatrix},
\bbB_i =
\begin{bmatrix}
-0.0001 & 0\\ -0.02  & 0\\ 0 &-0.00008 \\ 0 & -0.01
\end{bmatrix}.
\end{align}
%
As before, we compute the control matrix $\bbK$ using standard LQR-control computation.

In Figure \ref{fig_bbb_dist} we show the results of a representative simulation of the control of $m=50$ balancing board ball systems with a latency bound of $\tau_{\max} = 10^{-3}$ seconds. Observe that, in this system, even with a large number of users the CALLS method can keep all systems very close to the center of the board, while the fixed PDR scheduler loses a few of the balls due to the agnosticism of the scheduler.

 To dive deeper into the benefits provided by control aware scheduling, we present in Figure \ref{fig_bbb_psr} a histogram of the actual  packet delivery rates each of the devices achieved over the representative simulation. It is interesting to observe that, in the CALLS method, the achieved PDRs are closely concentrated, ranging from 0.3 to 0.44. On the other hand, using a fixed PDR scheduling scheme, the non-variable rates are too strict for the low-latency system to support, and without control-aware scheduling the achieved PDRs range wildly from close to 0 to close to 1. In this case, some devices are able to transmit almost every cycle while others are almost never able to successfully transmit their packets. This suggests that, by using control aware scheduling, we indirectly achieve a sense of fairness across users over the long term. Further note that the PDRs required to keep the balancing board ball stable, e.g. 0.4, are relatively small. This is due to the fact that the balancing board ball features relatively slow moving dynamics, making it easier to control with less frequent transmissions. This is comparison to the inverted pendulum system, in which the pendulums were kept stable with PDRs in the range 0.6-0.75.

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{bbb_dist.pdf}
\caption{Average ball distance to center for $m=50$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme keeps all balancing balls close to center, while fixed-PDR scheduling cannot.}
\label{fig_bbb_dist}
\end{figure}
%%

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{bbb_psr.pdf}
\caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme achieves similar PDRs for all devices, while the control agnostic scheduling results in large variation in packet delivery.}
\label{fig_bbb_psr}
\end{figure}
%%

We present in Figure \ref{fig_bbb_bar} the final capacity results obtained over all the simulations for the balancing board ball system. Observe that proposed approach increases the number of supported devices by factor of 2 relative to the standard fixed PDR approach. The even greater improvement here relative to the inverted pendulum simulations can be attributed to the slower dynamics of the balancing board ball, which allows for even more gains using control-aware PDRs due to the lower PDR requirements of the system.

%%%%%%
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    ybar,
    enlargelimits=0.15,
    legend style={at={(0.5,0.85)},
      anchor=north,legend columns=-1},
    ylabel={\# devices supported},
    symbolic x coords={0.5 ms,1 ms},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot coordinates {(0.5 ms,9) (1 ms,27)};
\addplot coordinates {(0.5 ms,25) (1 ms,60)};
\legend{Fixed PDR, CALLS}
\end{axis}
\end{tikzpicture}
\caption{Total number of balancing ball board devices that can be controlled using Fixed-PDR and CALLS scheduling for various latency thresholds.}
\label{fig_bbb_bar}
\end{figure}
%%%%%%

### 2019-02-01 Low_latency_control/icassp/icassp2.tex
Add files via upload

%\documentclass[letterpaper, 9 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
\documentclass{article}
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
%\IEEEoverridecommandlockouts                              % This command is only needed ifyou want to use the \thanks command
%\overrideIEEEmargins                                      % Needed to meet printer requirements.
%\documentclass{article}
\usepackage{spconf}
\usepackage{cite}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\graphicspath{{fig/}{jpeg/}}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{epstopdf}
\usepackage{algorithmic}
\usepackage{subcaption}
%\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage{comment}
\input{../mysymbol.sty}
\input{../my_sections.tex}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{matrix} % for block alignment
\usetikzlibrary{decorations.markings} % for arrow heads
\usetikzlibrary{calc} % for manimulation of coordinates
\usetikzlibrary{decorations.pathmorphing,snakes, positioning, decorations.pathreplacing}

\usetikzlibrary{fit}
%\input{figures/tikh_styles}

\usepackage{pgfplots}
\usepackage{color}
%\usepackage{setspace}
%\usepackage{hyperref}
\usepackage{multirow}
%\usepackage{rotating}
%\usepackage{comment}
%\usepackage{flushend}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage[affil-it]{authblk}
%\usepackage{bm}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{enumerate}
%\usepackage{morefloats}
%\usepackage{setspace}

\newcommand{\QED}{\hfill\ensuremath{\blacksquare}}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}

%\IEEEoverridecommandlockouts 
\def \vneg {\vspace{-.5cm}}

\def\whp{\text{w.h.p}}
\def\E{\mathbb{E}}
\def\P{\mathbb{P}}
\def\supp{\text{support}}
\def\Tr{\text{Tr}}

%\addtolength{\textwidth}{10mm}
%\addtolength{\evensidemargin}{-5mm}
%\addtolength{\oddsidemargin}{-5mm}
%\addtolength{\textheight}{10mm} 
%\addtolength{\topmargin}{-5mm}

%\renewcommand \blue[1]{}  \renewcommand \red[1]{}

\newtheorem{assumption}{\hspace{0pt}\bf Assumption}
\newtheorem{lemma}{\hspace{0pt}\bf Lemma}
\newtheorem{proposition}{\hspace{0pt}\bf Proposition}
\newtheorem{example}{\hspace{0pt}\bf Example}
\newtheorem{observation}{\hspace{0pt}\bf Observation}
\newtheorem{theorem}{\hspace{0pt}\bf Theorem}
\newtheorem{corollary}{\hspace{0pt}\bf Corollary}
\newtheorem{fact}{\hspace{0pt}\bf Fact}
\newtheorem{remark}{\hspace{0pt}\bf Remark}
\newtheorem{test}{\hspace{0pt}\it Test Case}
\newtheorem{definition}{\hspace{0pt}\bf Definition}
\newtheorem{conj}{\hspace{0pt}\bf Conjecture}

\addtolength{\textwidth}{10mm}
\addtolength{\evensidemargin}{-5mm}
\addtolength{\oddsidemargin}{-5mm}
\addtolength{\textheight}{10mm} 
\addtolength{\topmargin}{-5mm}

% Title.
% ------

%
% Single address.
% ---------------
\name{\begin{tabular}{c}Mark Eisen$^*$, Mohammad M. Rashid$^\dagger$, Konstantinos Gatsis$^*$ \\
			Dave Cavalcanti$^\dagger$, Nageen Himayat$^\dagger$, Alejandro Ribeiro$^*$\end{tabular}
\thanks{Supported by Intel Science and Technology Center for Wireless Autonomous Systems.}}
\address{$^*$Department of Electrical and Systems Engineering, University of Pennsylvania \\
$^\dagger$Wireless Communications Research, Intel Corporation}
\title{Control Aware Communication Design for \\ Time Sensitive Wireless Systems}

\begin{document} 
\maketitle
\ninept

%\thispagestyle{empty}
%
\begin{abstract}
We consider the problem of allocating radio resources over wireless communication links to control a series of independent low-latency wireless control systems common in industrial settings. Supporting wireless control in time sensitive settings requires fast data rates over wireless links, which comes at the cost of reliability. It is challenging to meet both latency and reliability requirements with an equal or arbitrary allocation of resources. We thus propose a novel control-aware approach to the low-latency scheduling problem in which we incorporate control and channel state information in allocating bandwidth and data rates across the wireless links. Control systems that are in desirable states are given modest requirements on error rates, while systems in undesirable states are given more priority. We derive control-aware packet error rate targets for each system to satisfy stability goals and make scheduling decisions to meet such targets while reducing total transmission time. The resulting control-aware based method is tested in simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.
\end{abstract}
%

\begin{keywords}
wireless control, low-latency, control-aware, scheduling
\end{keywords}

\section{Introduction}
\input{introduction}

%\input{Problem_formulation_icassp}

\section{Time Sensitive Wireless Control Systems}\label{sec_problem}

\input{problem_formulation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%  S  E  C  T  I  O  N     %%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Control-Communication Co-Design}\label{sec_codesign}

\input{scheduling_alg}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   S   E   C   T   I   O   N   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \section{Simulation Results}\label{sec_numerical_results}

We perform a series of experimets on the canonical latency-constrained control problem of controlling a series of inverted pendulums on a horizontal cart. While conceptually simple, the highly unstable dynamics of the inverted pendulum make it a representative example of control system that requires fast control cycles, and subsequently low-latency communications when being controlled over a wireless medium. Consider a series of $m$ identical inverted pendulums---where each pendulum is attached at one end to a cart that can move along a single, horizontal axis---using the modeling of the inverted pendulum as provided by Quanser \cite{quanser}. The state is $p=4$ dimensional vector that maintains the position and velocity of the cart along the horizontal axis, and the angular position and velocity of the pendulum, i.e. $\bbx_{i,k} := [x_{i,k}, \dot{x}_{i,k}, \theta_{i,k}, \dot{\theta}_{i,k}]$. The system input $u_{i,k}$ reflects a horizontal force placed upon $i$th pendulum. By applying a zeroth order hold on the continuous dynamics with a state sampling rate of $0.01$ seconds and linearizing, we obtained the following discrete linear dynamic matrices of the pendulum system
%
\begin{align}\label{eq_control_orig}
\bbA_i =
\begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & 2.055 & -0.722 & 4.828 \\
0 & 0.023 & 0.91 & 0.037 \\
0 & 0.677 & -0.453 & 2.055
\end{bmatrix},
\bbB_i =
\begin{bmatrix}
0.034 \\ 0.168 \\ 0.019 \\ 0.105
\end{bmatrix}.
\end{align}
%

Because the state $\bbx_{i,k}$ measures the angle of the $i$th pendulum at time $k$, the goal is to keep this close to zero, signifying that the pendulum remains upright. The input matrix $\bbK$ is computed to be a standard LQR-controller. For the definition of FDs, TDs, and data rates, we use the parameters of IEEE 802.11ax Wi-Fi protocol \cite{liu2014ieee}, referred to as resource units, PPDUs, and MCS, respectivley.

We perform a set of experiments with the inverted pendulum varying both the latency threshold and number of devices $m$. We perform the scheduling using the proposed control-aware method for control-aware low latency scheduling and, as a point of comparison, consider scheduling using a standard fixed PDR of $0.99$ for all devices. Each simulation is run for a total of $1000$ seconds and is deemed ``successful'' if all pendulums remain upright for the entire run. We perform 100 such simulations for each combination of latency threshold and number of devices to determine how many devices we can support at each latency threshold using both the control-aware and fixed-PDR methods for scheduling.

In the left of Figure \ref{fig_sim_results} we show the results of a representative simulation of the control of $m=25$ pendulum systems with a latency bound of $\tau_{\max} = 10^{-3}$ seconds. In both graphs we show the average distance from the center vertical of each pendulum over the course of 1000 seconds. In the top figure, we see by using the control-aware control-aware method we are able to keep each of the 25 pendulums close to the vertical for the whole simulation. Meanwhile, using the standard fixed PDR, we are unable to meet the scheduling limitations imposed by the latency threshold, and many of the pendulums swing are unable to be kept upright, as signified by the large deviations from the origin. This is due to the fact that certain pendulums were not scheduled when most critical, and they subsequently became unstable, while the control-aware adapted successfully to the needs of each system.

We present in the right of Figure \ref{fig_sim_results} the maximum number of uses each scheduler could support under different latency requirements without any instability. Each device must be kept with a  $|\theta_{i,k}| \leq 0.05$ error region for 100 independent simulations. We observe that the control-aware approach is able to increase the number of devices supported in each case, with up to $1.5$ factor increase over the fixed PDR scheduling.

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   S   E   C   T   I   O   N   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \section{Conclusion}
In this paper we develop a control-communication control-aware approach towards scheduling for low-latency, or time sensitive, wireless control systems. Because many control systems in industrial settings require very low latency transmission to operate effectively, there is an intrinsic challenge in trading off the data rates necessary to achieve low latency with the packet error rates necessary for high reliability. We demonstrate that packet delivery rates required to meet control-specific goals changes dynamically with the state of the systems. This allows to develop a scheduling method that smartly allocates bandwidth and data rates among many independent systems based upon their own current PDR requirements. We demonstrate the improvements relative to control-agnostic scheduling procedures in simulation experiments.

\newpage 
% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\urlstyle{same}
\bibliographystyle{../IEEEbib}
\bibliography{../wireless_ll_control,../scheduling_control}

\end{document}

### 2019-02-01 Low_latency_control/icassp/introduction.tex
Add files via upload

The increasing scale of modern IoT and industrial control systems has motivated the development of wireless control system technology that can achieve reliable performance in these settings~\cite{varghese2014wireless, li2017review}. One of the primary challenge of wireless control in industrial settings, however, is the time sensitive nature of the systems, thus requiring low latency wireless transmissions~\cite{varghese2014wireless}.  The noise of the wireless channel makes it difficult to simultaneously maintain high reliability while achieving low latency. This motivates the design of resource allocation and scheduling strategies that can both meet reliability \emph{and} latency requirements of the industrial control system.

In the wireless communications research and industry, many radio resource allocation schemes in the form of wireless scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}. For time-sensitive applications, delay-aware schedulers such as EDF~\cite{wu2014analysis} and WFQ~\cite{lu1999fair} have been developed,  while M-LWDF \cite{andrews2001providing} extends these ideas to include channel state information.

Likewise, in the context of wireless control systems, dynamic schedulers have been developed that provide access to the communication medium dynamically at each step. %, for example by dynamically assigning priorities to the competing tasks.
Initial approaches make scheduling decisions based on abstractions of control performance~\cite{Branicky_RM, Liu_Real_time_systems}. More recently, ``control-aware'' scheduling approaches make decisions explicitly based on current control system states ,~\cite{Donkers_switched, Walsh_stability, Hristu_Kumar_interrupt_based, Egerstedt_queue, Hirche_Scheduling_Price, Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
%
Further work takes into account current wireless channel conditions when attempting to meet target control system reliability requirements~\cite{GatsisEtal15}. The work here takes a similar channel-based opportunistic approach, while further incorporating current control states with the explicit goal of meeting low-latency requirements. The proposed approach has been developed specifically for the IEEE 802.11ax protocol in \cite{EisenEtal18}.

%In this paper, we develop a control-aware, high reliability, low-latency IEEE 802.11ax WiFi protocol~\cite{bellalta2016ieee} that is designed to reduce the total transmission time at each uplink cycle in the control loop. This is done through the mathematical formulation of the control system design goal in the form of a Lyapunov function that ensures stability of the control system. This formulation naturally induces a bound on the packet delivery rate each control system needs to achieve to meet the control-based goal. Such packet delivery rates depend upon current control and channel states and thus dynamically change over the course of the system life time. This can be viewed as an opportunistic protocol with respect to both the current control states \emph{and} channel conditions. Furthermore, because these control-based success rate requirements may be significantly lower than traditional, high reliability communication demands, the proposed method is better suited to find scheduling configurations that can moreover meet the strict latency requirements imposed by the physical system. This is in contrast to the control-aware approach taken in \cite{GatsisEtal15}, which focuses on low-power and infrequent transmissions rather than the low-latency setting of interest in industrial control.

This paper is organized as follows. We formulate the wireless control system in which state information is communicated to the control over a wireless channel as a switched dynamical system (Section \ref{sec_problem}). We then discuss the communication architecture that determines the speed and error rate of transmissions (Section \ref{sec_comm_model}). With this formulation, we adapt concepts of control-communication control-aware for low latency settings by using current control states and channel conditions to derive dynamic packet success rates necessary for each user (Section \ref{sec_codesign}). In this manner, control systems with the most critical communication needs are given higher packet delivery rate targets to meet. The scheduling procedure leverages these dynamic, more liberal rate requirements to reduce total latency, incorporating both a selective scheduling procedure  (Section \ref{sec_rss}) and an assignment-method based that attempts to further reduce total transmission time (Section \ref{sec_assignment}). The performance of the control-aware method is analyzed in a representative low-latency simulation experiment in which its performance is compared against a control-agnostic procedure (Section \ref{sec_numerical_results}).

### 2019-02-01 Low_latency_control/icassp/optimal_scheduling.tex
Add files via upload

Using the communication model of 802.11ax just outlined and the control-based Lyapunov metric of \eqref{eq_perf}, we can formulate an optimization problem that characterizes the exact optimal scheduling of transmissions with a transmission window to maximize control performance. The optimal scheduling and allocation selects the set of RUs $\bbSigma$. MCS $\bbmu$, and PPDUs $\bbalpha$ for all devices--which in effect fully determine the schedule---such to minimize a cost subject to scheduling design and feasibility constraints. In particular, we discuss two related, alternative formulations of the low-latency scheduling problem.

\subsection{Latency-constrained scheduling}\label{sec_optimal_a}
In the latency-constrained formulation, we are interested in minimizing a common control cost subject to strict latency requirements. In particular, in the low-latency setting we set a bound $\tau_{\max}$ on the total transmission time across all PPDUs in a TXOP. This constraint is relevant in design of MAC-layer protocols that set strict limits on transmission times. In addition, the RU and PPDU allocation across devices must be feasible, i.e., two devices cannot be transmitting in the same frequency band in the same PPDU.

Recall the PDR function $q(\bbh,\mu,\bbsigma)$ and consider that this can alternatively be interpreted as the probability of closing the control loop under certain channel conditions and scheduling parameters. From there, we can now write the expected Lyapunov value for at time $k+1$ given its current state $\bbx_{i,k}$, channel state $\bbh_{i,k}$, MCS $\mu_i$, and RU $\bbsigma_i$ using the expected cost in \eqref{eq_perf}. By defining $\bbx_{i,k+1}^c$ and $\bbx_{i,k+1}^o$ as the closed loop and open loop states, respectively, as determined by the switched system in \eqref{eq_control_switch}, this is written as
%
\begin{align}
J_i(\hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i) &:= \E (L(\bbx_{i,k+1}) \mid \hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i) \nonumber \\
=& (1-q(\bbh_{i,k},\mu_i,\bbsigma_i))  \E L(\bbx_{i,k+1}^o \mid \hbx^{(l_i)}_{i,k}) \nonumber \\
&+q(\bbh_{i,k},\mu_i,\bbsigma_i) \E L(\bbx_{i,k+1}^c \mid \hbx^{(l_i)}_{i,k}). \label{eq_lyap_p}
\end{align}
%

For notational convenience, we collect all current estimated control states at time $k$ as $\hbX_k := [\hbx^{(l_1)}_{1,k}, \hdots, \hbx^{(l_m)}_{m,k}]$ and channel states $\bbH_k = [\bbh_{1,k}, \hdots, \bbh_{m,k}]$. Now, define the total control cost, given the current states and scheduling parameters as some aggregation of the combined expected future Lyapunov costs across all devices, i.e.,
%
\begin{align}\label{eq_cost_total}
 \tdJ( \hbX_{k},&\bbH_{k}, \bbmu, \bbSigma) := \\
 &g(J_1(\hbx^{(l_1)}_{1,k},\bbh_{1,k},\mu_1,\bbsigma_1), \hdots, J_m(\hbx^{(l_m)}_{m,k},\bbh_{m,k},\mu_m,\bbsigma_m)). \nonumber
 \end{align}
 %
 Natural choices of the aggregation function $g(\cdot)$ are, for example, either the sum or maximum of its arguments.

The optimal scheduling at transmission time $k$ is formulated as the one which minimizes this cost $\tdJ$ while satisfying low-latency and feasibility requirements of the schedule, expressed formally with the following optimization problem.
%
\begin{align}\label{eq_problem}
[&\bbSigma_k^*, \bbmu_k^*,\bbalpha_k^*]  := \argmin_{\bbSigma, \bbmu, \bbalpha, S} \tdJ(\hbX_{k},\bbH_{k}, \bbmu, \bbSigma) \\
& \quad \st \sum_{i: \alpha_i = s} \varsigma_{i}^j \leq 1, \quad \forall j, s, \label{eq_c1} \\
& \qquad \quad \sum_{s=1}^S \hat{\tau}(\bbSigma, \bbmu, \bbalpha, s)  \leq \tau_{\max}, \label{eq_c2}\\
& \qquad \quad 1 \leq \alpha_i \leq S, \quad \forall i, \label{eq_c3}\\
& \qquad \quad \bbsigma_i \in \ccalS, \ \forall i, \quad \bbmu \in \ccalM^m, \quad \bbalpha \in \mathbb{Z}_+^m, \quad S \in \mathbb{Z}_+. \label{eq_c4}
\end{align}
%
The optimization problem in \eqref{eq_problem} provides a precise and instantaneous selection of frequency allocations between devices given their current control states $\hbX_k$ and communication states $\bbH_k$. The constraints in \eqref{eq_c1}-\eqref{eq_c4} encode the following scheduling conditions. The constraint \eqref{eq_c1} ensures that for every PPDU $s$, there is only one device transmitting on a frequency slot $j$. In \eqref{eq_c2}, we set the low-latency transmission time constraint in terms of the sum of all transmission times for each PPDU $s$. The constraint in \eqref{eq_c3} bounds each transmission slot by the total number of PPDU's $S$ while \eqref{eq_c4} constrains each variable to its respective feasible set. Note that $S$ is itself treated as an optimization variable in the above problem, so that the number of PPDUs may vary as needed.

Observe in the objective in \eqref{eq_problem} that, by minimizing an aggregate of local control costs, the devices with the highest cost $J_i$ as described by \eqref{eq_lyap_p} will be given the most bandwidth or most favorable frequency bands to increase probability of successful transmission $q(\bbh_{i,k},\mu_i,\bbsigma_i)$. This in effect increases the chances those devices will close their control loops and be driven towards a more favorable state. Likewise, a device who is experiencing very adverse channel conditions may not be allocated prime transmission slots to reserve such resources who have more favorable channel conditions. In this way, we say this is \emph{control-aware} scheduling, as it considers both the control and channel states of the devices to determine optimal scheduling. However, we stress that the optimization problem described in \eqref{eq_problem}-\eqref{eq_c4} is by no means easy to solve. In fact, the optimization over multiple discrete variables makes this problem combinatorial in nature. In the following section, we discuss a practical reformulation of the problem above and develop heuristic methods to approximate the solutions in realistic low-latency wireless applications.

%\begin{remark}\label{remark_per}\normalfont
%The value in the formulation above is not only in its adaptability to the control system itself, but informs the choosing of transmission success rates---or, effectively, the packet error rates (PER)---that are necessary to keep the control systems in favorable states. Depending on the particular system dynamics described in \eqref{eq_control_orig}, such PERs may be considerably more lenient than the default target transmission success rates used in practical wireless systems (e.g. $10^{-3}$). This notion is explored more explicitly in Section \ref{sec_per}.
%\end{remark}

\subsection{Control-constrained scheduling}\label{sec_optimal_b}
We reformulate the problem in \eqref{eq_problem}-\eqref{eq_c4} to an alternative formulation that more directly informs the control-aware, low-latency scheduling method to be developed. To do so, we introduce a \emph{control-constrained} formulation, in which the Lyapunov decrease goals are presented as explicit requirement, i.e. constraints in the optimization problem. We are interested, then, in constraint of the form
%
\begin{align}\label{eq_control_constraint}
\tdJ(\hbX_{k},\bbH_{k}, \bbmu, \bbSigma) \leq J_{\max},
\end{align}
%
where $J_{\max}$ is a limiting term design to enforce desired system performance. Determining this constant is largely dependent on the particular application of interest, needs of the control systems, and also may be related to the choice aggregation function $g(\cdot)$ in \eqref{eq_cost_total}. For example, $J_{\max}$ may represent a point at which control systems become volatile, unsafe, or unstable.

For the scheduling procedure developed in this paper, we focus on a particular formulation of the control constraint in \eqref{eq_control_constraint} that constrains the expected future Lyapunov value of each system by a rate decrease of its current value. In particular the following rate-decrease condition for each device $i$,
%
\begin{align}\label{eq_lyap_constraint}
J_i(\hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i) \leq \rho_i \E [L(\bbx_{i,k}) \mid \hbx^{(l_i)}_{i,k}] + c_i,
\end{align}
%
where $\rho_i \in (0,1]$ is a decrease rate and $c_i \geq 0$ is a constant. Recall the definition of $J_i(\hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i)$ in \eqref{eq_lyap_p} as the expected Lyapunov value of time $k+1$ given its current estimate and scheduling $\mu_i, \bbsigma_i$. The constraint in \eqref{eq_lyap_constraint} ensures the future Lyapunov cost will exhibit a decrease of at least a rate of $\rho_i$ for device $i$ in expectation. The constant $c_i$ is included to ensure this condition is satisfied by default if the state $\hbx^{(l_i)}_{i,k}$ is already sufficiently small.

We formulate the control-constrained scheduling problem by substituting the latency constraint with the control constraint in \eqref{eq_lyap_constraint}, i.e.,
%
\begin{align}\label{eq_problem2}
[&\bbSigma_k^*, \bbmu_k^*,\bbalpha_k^*]  := \argmin_{\bbSigma, \bbmu, \bbalpha, S} \sum_{s=1}^S \hat{\tau}(\bbSigma, \bbmu, \bbalpha, s) \\
& \quad \st \sum_{i: \alpha_i = s} \varsigma_{i}^j \leq 1, \quad \forall j, s, \label{eq_c12} \\
& \quad   J_i(\hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i) \leq \rho \E [L(\bbx_{i,k}) \mid \hbx^{(l_i)}_{i,k}] + c_i \ \forall i, \label{eq_c22}\\
& \qquad \quad 1 \leq \alpha_i \leq S, \quad \forall i, \label{eq_c32}\\
& \qquad \quad \bbsigma_i \in \ccalS, \ \forall i, \quad \bbmu \in \ccalM^m, \quad \bbalpha \in \mathbb{Z}_+^m, \quad S \in \mathbb{Z}_+. \label{eq_c42}
\end{align}
%
Observe that the objective in \eqref{eq_problem2} is now to minimize the total transmission time, rather than being forced as an explicit constraint. In this way, the optimization problem defined in 
\eqref{eq_problem2}-\eqref{eq_c42} can be viewed as an alternative to the latency constrained problem in \eqref{eq_problem}-\eqref{eq_c4}. Because the scheduling algorithm we develop in this paper requires the ability to quickly identify feasible solutions, we focus our attention on the control-constrained formulation in \eqref{eq_problem2}-\eqref{eq_c42}. Before presenting the details of the scheduling algorithm, we present a brief remark regarding the  addition of ``safety'', or worst-case, constraints to either problem formulation.

\begin{remark}\label{remark_worst_case}\normalfont
The control constraint in \eqref{eq_c22} is formulated to guarantee an average decrease of expected Lyapunov value by a rate of $\rho$. This is of interest to ensure the system states are driven to zero over time. However, in practical systems we may also be interested in protecting against worst-case behavior, e.g. entering an unsafe or unstable region. Consider a vector $\bbb_i \in \reals^p$ as the boundary of safe operation of system $i$. A constraint that protects against exceeding this boundary can be written as
%
\begin{align}\label{eq_worst_case}
\P [ |\bbx_{i,k+1}| \geq \bbb_i \mid \hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i ] \leq \delta,
\end{align}
%
where $\delta \in (0,1)$ is small. The expression in \eqref{eq_worst_case} can be included as an additional constraint to either the latency-constrained or control-constrained scheduling problems previously discussed.
\end{remark}

### 2019-02-01 Low_latency_control/icassp/problem_formulation.tex
Add files via upload

%%%%%%%%%%%%%%%%% F I G U R E %%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\input{../images/tikz_wncs4}
%\caption{Wireless control system with $m$ independent systems. Each system contains a sensor that measure state information, which is transmitted to the controller over a wireless channel. The state information is used by the controller to determine control policies for each of the systems.The communication is assumed to be wireless in the uplink and ideal in the downlink.}
%\label{fig_wcs}
%\end{figure}

Consider a system of $m$ independent linear control systems, or devices, where each system $i=1,\hdots,m$ maintains a state variable $\bbx_{i} \in \reals^p$. The dynamics evolve over a discrete time index $k$.  Applying an input $\bbu_{i,k} \in \reals^q$ causes the state and output to evolve based on the generic state space representation,
%
\begin{align}\label{eq_control_orig}
\bbx_{i,k+1} &= \bbA_{i} \bbx_{i,k} + \bbB_{i} \bbu_{i,k} + \bbw_k
\end{align}
%
where $\bbA_{i} \in \reals^{p \times p}$ and $\bbB_{i} \in \reals^{p \times q}$ are matrices that define the system dynamics, and $\bbw_{k} \in \reals^{p}$ is Gaussian noise with co-variance $\bbW_i$ that captures the noise in the model. We further assume the state transition matrix $\bbA_i$ is on its own unstable, i.e. has at least one eigenvalue greater than 1. This is to say that, without an input, the dynamics will drive the state $\bbx_{i,k} \rightarrow \infty$ as $k \rightarrow \infty$.

In the time sensitive wireless control system, each system is closed over a wireless medium through which the sensor located at the control plant sends state information to the controller located at a common wireless base station (BS). Using the state information $\bbx_{i,k}$ received from device $i$ at time $k$, the controller determines the input $ \bbu_{i,k}$ to be applied. We stress that, due to the latency constraints of the control system operation, the BS gives each device only a short time window, or transmission opportunity, to finish transmitting its state information. This model restricts its attention to wireless connections in uplink of the control loop, while downlink is assumed to occur over an ideal channel.

To derive the mathematical model, consider a generic linear control $ \bbu_{i,k} = \bbK_i \bbx_{i,k}$ for some matrix $\bbK_i \in \reals^{q \times p}$. Many common control policies indeed can be formulated in such a manner, such as LQR control. This matrix $\bbK$ is chosen such that the closed loop dynamics $\bbA + \bbB \bbK$ has all eigenvalues less that 1. Due to noise in the channel, there is potential for state information packets  to be dropped in the uplink, which is modeled as ``open-loop'' configuration. Meanwhile, successful transmission are modeled as a ``closed-loop'' configuration. To account for incomplete state information at the BS due to packet drops, we consider the estimate of state information of device $i$ known to the BS/controller at time $k$ as
%
\begin{align}\label{eq_state_est}
\hbx^{(l_i)}_{i,k} :=(\bbA_i + \bbB_i\bbK_i)^{l_i} \bbx_{i,k-l_i},
\end{align}
%
where $k-l_i \geq k-1$ is the last time instance in which control system $i$ was closed. Observe that in \eqref{eq_state_est} we assume that the BS/controller has knowledge of $\bbA_i, \bbB_i$, and $\bbK_i$, but not the noise $\bbw_k$ present in \eqref{eq_control_orig}.

At time $k$, if the state information is received, the controller applies the input $\bbu_{i,k} = \bbK_i \bbx_{i,k}$ using the exact state, and otherwise applies input $\bbu_{i,k} = \bbK_i \hbx_{i,k}$. We obtain then the following switched system dynamics for $\bbx_{i,k}$ as
%
\begin{align}\label{eq_control_switch}
\bbx_{i,k+1} &= \begin{cases}
(\bbA_i + \bbB_i \bbK_i) \bbx_{i,k} + \bbw_k, \ \text{in closed-loop}, \\
\bbA_i \bbx_{i,k} + \bbB_i\bbK_i\hbx^{(l_i)}_{i,k} + \bbw_k, \ \text{in open-loop}.
\end{cases}
\end{align}
%
The transmission counter $l_i$ is updated at time $k$ as
%
\begin{align}
l_i &\leftarrow \begin{cases}
1, \ \text{in closed-loop}, \\
l_1 + 1, \ \text{in open-loop}.
\end{cases} \label{eq_time_switch}
\end{align}
%
Observe that the successive error between the true and estimated state can be written as
$\bbe_{i,k} := \bbx_{i,k} - \hbx^{(l_i)}_{i,k} = \sum_{j=0}^{l_i-1}\bbA_i^{j} \bbw_{i,k-j-1}$. It is evident that this error grows with the transmission counter $l_i$. We proceed now to describe the architecture of the wireless communications that determines both the speed and error rate of the state information transmissions over the wireless channel.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Communication architecture}\label{sec_comm_model}

We consider a standard communication architecture in which, within a transmission window of a single cycle in the control loop, devices are scheduled by the BS across discrete time division (TD) and frequency division (FD) slots, both of which may vary in size. To adapt transmission lengths, each device is given a data rate (DR) parameter. Collectively, the assignment of device to their respective TDs, FDs, and with determined DR fully specifies the scheduling for the given transmission window. This architecture reflects that used scheduling-based multiple access wireless protocols, such as LTE \cite{sesia2011lte} , 5G \cite{agiwal2016next}, and next-generation WiFi IEEE 802.11ax \cite{liu2014ieee}. The transmission power is assumed fixed and equal across all devices.

We state this model formally with the following variable definitions. Consider that the total allowable bandwidth is divided into $b$ discrete bands of equal size,  the FD slot of device $i$ is specified by a binary vector $\bbsigma_{i} \in \{0,1\}^b$, where the $j$ element $\bbsigma_i(j) = 1$ if device $i$ transmits in the $j$th frequency band. Note that a device may transmit in multiple \emph{adjacent} bands simultaneously to indicate a FD slot of larger bandwidth. Because not all binary vectors of length $b$ define a possible FD, we define the set $\ccalS \subset \{0,1\}^{b}$ to collect such a set of definable FD vectors.  The FD assignment $\bb0 \in \ccalS$ is defined to reflect that a device does not transmit.

We further define for device $i$ a positive integer value $\alpha_i \in \{1,2,\hdots,S\}$ that denotes its TD slot and a real-valued $\mu_i \geq \mu_0$ to denote its DR, where $\mu_0$ is a minimum allowable rate. Finally, we define $\bbh_{i,k} \in \reals^b_+$ to be the set of fading channel states experienced by device $i$ at cycle $k$, where $\bbh_{i,k}(j)$ is the fading channel gain in frequency band $j$. We assume that channel conditions do not change across different TDs within a single cycle/transmission window $k$.

The variables $\{\bbsigma_i, \alpha_i, \mu_i\}$ define the scheduling specification for user $i$ and, with the channel conditions $\bbh_{i,k}$, determine the communication performance obtained with such a scheduling decision. We first define a function  $q(\bbh,\mu,\bbsigma)$ which returns the probability of successful transmission/closing loop, otherwise called packet delivery rate (PDR)---given the channel conditions, DR and FD selections (this is independent of TD selection). Likewise, define by $\tau(\mu,\bbsigma)$ a function that, given an DR $\mu$ and FD $\bbsigma$, returns the maximum time taken for a single transmission attempt (this is independent of TD selection and channel conditions). Both of these functions play a critical role in determining scheduling decisions in time-sensitive wireless control system settings. We are, in particular, interested in exploring the trade-off between PDR and transmission time that comes from varying $\mu$. Generally speaking, the functions $q(\bbh,\mu,\bbsigma)$ and $\tau(\mu,\bbsigma)$ relate to $\mu$ by 
%
\begin{equation}
\mu' > \mu \implies q(\bbh,\mu',\bbsigma) \leq q(\bbh,\mu,\bbsigma), \quad \tau(\mu',\bbsigma) \leq \tau(\mu,\bbsigma).
\end{equation}
%

The goal then is to determine schedule $\{\bbsigma_i, \alpha_i, \mu_i\}_{i=1}^m$ at every cycle $k$ that keeps all control systems in a region of desirable performance, while keeping the total transmission time across all TDs small to meet latency requirements.

%Because a TD cannot finish until all transmissions within the TD have been completed, the total transmission time of a single TD $s$ is the maximum transmission time taken by all devices within that time slot. We define the transmission time of TD slot $s$ as
%%
%\begin{equation}\label{eq_time_slot}
%\hat{\tau}(\bbSigma, \bbmu, \bbalpha, s) := \max_{i: \alpha_i = s} \tau(\mu_i, \bbsigma_i) + \tau_0(\bbalpha,s),
%\end{equation}
%%
%where $\tau_0: \mathbb{Z}_{++}^m \times \mathbb{Z}_{++} \rightarrow \reals_+$ is a function that specifies the communication overhead of TD $s$. This overhead may consist of, e.g., the time required to send TFs to scheduled users, as seen in Figure \ref{fig_multiplex}.

### 2019-02-01 Low_latency_control/icassp/scheduling_alg.tex
Add files via upload

We develop a control-communication control-aware approach towards scheduling in time-sensitive settings. Due to the tight latency constraints placed on the communications, we leverage knowledge of the control state and dynamics to determine a more principled and opportunistic method of scheduling devices. In particular, we use control information to identify maximum data rates we can achieve while maintaining strong control performance so as to meet latency targets. We first derive a manner in which we can evaluate control performance. Consider a quadratic Lyapunov function $L(\bbx) := \bbx^T \bbP \bbx$ for some positive definite $\bbP \in \reals^{p \times p}$ that measures the performance of a system as a function of the state. For the system to both remain stable and over time be driver to zero, it is necessary for value of $L(\bbx_{k+1})$ to decrease relative to $L(\bbx_k)$ at all times $k$. We cannot guarantee this occurs deterministically, but instead consider a condition on the estimated future Lyapunov cost, i.e.
%
\begin{align} \label{eq_lyop}
\E [L(\bbx_{i,k+1}) \mid \hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i) \leq \rho \E [L(\bbx_{i,k}) \mid \hbx^{(l_i)}_{i,k}] + c_i,
\end{align}
%
for some $\rho \in (0,1)$. The condition in \eqref{eq_lyop} specifies that the expected Lyapunov cost for system $i$ should decrease by a factor of $\rho$ from cycle $k$ to $k+1$ (up to constant $c_i$). Observe that this expectation is conditioned upon the estimated state $\hbx^{(l_i)}_{i,k}$, channel conditions $\bbh_{i,k}$, as well as a scheduling decision $\{\mu_i, \bbsigma_i\}$. The scheduling decision impacts this expected value through the resulting PDR $q(\bbh_i, \mu_i, \bbsigma_i)$ which determines the probability of closing the control loop and thus diminishing this Lyapunov cost. We may derive an explicit or equivalent condition on  $q(\bbh_i, \mu_i, \bbsigma_i)$ to satisfy the condition in \eqref{eq_lyop}, which we present in the following proposition. The proof can be found in \cite{EisenEtal18}.
%%%%%%%%%%%%%%%%%%%
\begin{proposition}\label{prop_pdr_constraint}
Consider the switched dynamics in \eqref{eq_control_switch}. Define the closed-loop state transition matrix $\bbA^c_i := \bbA_i + \bbB_i \bbK_i$ and $j$-accumulated noise $\omega_i^j := \Tr[(\bbA_i^T\bbP^{1/j} \bbA_i)^{j} \bbW_i]$. The control constraint in \eqref{eq_lyop} is satisfied for device $i$ if and only if the following condition on PDR $q(\bbh_{i,k},\mu_i,\bbsigma_i)$ holds, 
%
\begin{align}\label{eq_pdr_constraint}
&q(\bbh_{i,k},\mu_i,\bbsigma_i) \geq \tdq_i(\hbx^{(l_i)}_{i,k}) :=  \\ &\ \frac{1}{\Delta_i} \left[  \left\| (\bbA^c_i - \rho_i\bbI)\hbx^{(l_i)}_{i,k} \right\|_{\bbP^{\frac{1}{2}}}^2+ (1-\rho_i)\sum_{j=0}^{l_i-1} \omega_i^j + \omega_i^{l_i} - c_i  \right],  \nonumber
\end{align}
%
where we have further defined the constant
%
\begin{equation}\label{eq_noise_diff}
\Delta_i :=  \sum_{j=0}^{l_i-1}[\omega^{j+1}_i - \Tr(\bbA^{cT}_i (\bbA_i^T\bbP^{1/j}\bbA_i)^j \bbA^c_i\bbW_i)  ].
\end{equation}
\end{proposition}
%

Proposition \ref{prop_pdr_constraint} formally establishes lower bound $\tdq_i(\hbx^{(l_i)}_{i,k})$ on the PDR of device $i$ such that its Lyapunov condition in \eqref{eq_lyop} is met in expectation. This bound is determined based upon the current estimated state, overall system dynamics, and transmission history and effectively sets a constraint on the scheduling bandwidth of FD $\bbsigma_i$ and DR $\mu_i$. We point out the relevant components of the expression in \eqref{eq_pdr_constraint}. First, note that the first term on the right hand will grow larger as the state gets larger, or closer to instability. Likewise, the latter two terms on the right hand side together grow larger as the estimating noise increases due to successive dropped packets, due to both the noise variance $\bbW_i$ and last-update counter $l_i$. In this manner, it is both the current control state and transmission history, as they relate to the dynamics of the system, that set a delivery rate requirement for each device.

The PDR condition in \eqref{eq_pdr_constraint} is valuable in the time-constrained settings because it allows us to dynamically adapt the data rate needs of each device relative to their control state. While the latency requirement effectively constrains the total resources available, the ability to properly identify the users to be given scheduling slots is and important consideration in maintaining overall reliable performance. It is worth pointing out, that depending upon the system dynamics of a particular control system, the PDRs derived in \eqref{eq_pdr_constraint} may often be in practice significantly lower than the standard, fixed PDR requirements used in high-reliability systems, e.g. $\geq 0.99$. In such cases, the identification of PDR requirements necessary for proper operation of the control system can reduce a large amount of resource constraints, as is later seen in the simulations in Section \ref{sec_numerical_results}. We proceed now to discuss the ways in which the dynamic and more lenient PDR targets in \eqref{eq_pdr_constraint} may be leveraged by the scheduling to further reduce the total transmission times.

%\begin{remark}\normalfont
%It is worthwhile to note that by placing a stricter Lyapunov decrease constraint with smaller  rate $\rho_i$ in \eqref{eq_c22}, then the first term on the right hand side of \eqref{eq_pdr_constraint} also grows larger and increases the necessary PDR. Generally, selecting a smaller $\rho$ will result in a faster convergence to stability but will require stricter communication requirements. In fact, we may use the inherent bound on the probability $q(\bbh_{i,k},\mu_i,\bbsigma_i) \leq 1$ to find a lower bound on the Lyapunov decrease rate $\rho_i$ that can be feasibly obtained based upon current control state and system dynamics. This bound, however, may not be obtainable in practice due to the scheduling constraints. In practice, we select $\rho_i$ to be in the interval $[0.90,0.1)$. 
%\end{remark}

\subsection{Selective scheduling}\label{sec_rss}
We first consider a stochastically \emph{selective scheduling} protocol, whereby we do not attempt to schedule every device at each transmission cycle, but instead select a subset to schedule a principled random manner. Define by $\nu_{i,k} \in [0,1]$ the probability that device $i$ is included in the transmission schedule at time $k$ and further recall by $q(\bbh_{i,k},\mu_i,\bbsigma_i)$ to be the packet delivery rate with which it transmits. Then, we may consider the \emph{effective} packet delivery rate $\hat{q}$ as 
%
\begin{align}\label{eq_effective_pdr}
\hat{q}(\bbh_{i,k},\mu_i,\bbsigma_i) = \nu_{i,k} q(\bbh_{i,k},\mu_i,\bbsigma_i)
\end{align}
%
Observe that in order to meet the PDR target defined in \eqref{eq_effective_pdr}, device $i$ would need to meet a \emph{modified} PDR  target $q(\bbh_{i,k},\mu_i,\bbsigma_i) \geq \tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k}$. While imposing a tighter PDR requirement will indeed require longer transmission times, this added time cost is generally less than the transmission overhead of scheduling all users. In particular, the scheduling probability of device $i$ is defined relative to its PDR requirement $\tdq_i(\hbx^{(l_i)}_{i,k})$ as 
%
\begin{align}\label{eq_prob_c}
\nu_{i,k} := e^{\tdq_i(\hbx^{(l_i)}_{i,k})-1} .
\end{align}
%
Notice that, when a transmission is necessary, i.e. $\tdq_i(\hbx^{(l_i)}_{i,k}) = 1$, then device $i$ is included in the scheduling with probability $\nu_{i,k} = 1$.

\subsection{Assignment-based scheduling}\label{sec_assignment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%   A   L   G   O   R   I   T   H   M   %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\linespread{1.3}
\begin{algorithm}[t] \begin{algorithmic}[1]
\STATE \textbf{Parameters:} Lyapunov decrease rate $\rho$
\STATE \textbf{Input:} Channel conditions $\bbh_{i,k}$ and estimated states $\hbx^{(l_i)}_{i,k}$ for all $i$
\STATE Compute target PDR $\tdq_i(\hbx^{(l_i)}_{i,k})$ for each device $i$ [cf. \eqref{eq_pdr_constraint}].
\STATE Determine selection probabilities $\nu_{i,k}$ for each device [cf. \eqref{eq_prob_c}].
\STATE Select devices $\ccalI_k$ with probs. $\{\nu_{1,k},\hdots,\nu_{m,k}\}$
\STATE Determine set of FDs/TDs $\ccalS'_k$ [cf. \eqref{eq_ru_sets}].
\STATE Determine max. DR for each device/FD assignment [cf. \eqref{eq_mcs_select}].
\STATE Schedule selected devices via assignment method \cite{kuhn1955hungarian}.
\STATE \textbf{Return:} Scheduling variables $\{\bbsigma_{i}, \mu_i, \alpha_i\}_{i=1}^m$ 
\end{algorithmic}
\caption{Control-aware scheduling for low-latency at cycle $k$}\label{alg_calls} \end{algorithm}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%
\begin{table}[]
\centering
\begin{tabular}{|c|c|c|}
\hline 
\textbf{TD 1} & \textbf{TD 2} & \textbf{TD 3}        \\ \hline \hline
\multicolumn{1}{|c|}{FD 1} & \multicolumn{1}{c|}{\multirow{2}{*}{FD 8}} & \multicolumn{1}{c|}{\multirow{4}{*}{FD 11}} \\ \cline{1-1}
\multicolumn{1}{|c|}{FD 2} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \cline{1-2}
\multicolumn{1}{|c|}{FD 3} & \multicolumn{1}{c|}{\multirow{2}{*}{FD 9}} & \multicolumn{1}{c|}{}                       \\ \cline{1-1}
\multicolumn{1}{|c|}{FD 4} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \hline
\multicolumn{1}{|c|}{FD 5} & \multicolumn{1}{c|}{\multirow{4}{*}{FD 10}} & \multicolumn{1}{c|}{\multirow{4}{*}{FD 12}} \\ \cline{1-1}
\multicolumn{1}{|c|}{FD 6} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \cline{1-1}
\multicolumn{1}{|c|}{FD 7} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \cline{1-1} \hline
\end{tabular}
\caption{Example of FD selection with $m_k= 12$ devices. There are a total of $S_k = 3$ TDs, given $n_1=9$, $n_2 = 3$, $n_3 = 2$ FDs, respectively.}
\label{tab_rus}
\end{table}
%%%%%

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% F I G U R E %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[t]
\centering
\includegraphics[width=0.4\linewidth,height=0.25\linewidth]{../images/ip_dist2.pdf} \qquad
\input{ip_bar.tex} 
\caption{Simulation results for a series of inverted pendulums controlled over shared wireless channel. (left) The average distances from center vertical for $m=25$ pendulums. The control-aware, or ``co-design'', scheduler keeps all pendulums close, unlike the fixed PDR scheduler. (right) For different latency bounds, the control-aware can support more pendulums than the fixed PDR scheduler.}\label{fig_sim_results}
\end{figure*}

Given the set of devices selectively scheduled via \eqref{eq)prob_c}, we proceed to discuss an assignment-based formulation that can be employeed to select a low-latency schedule. Define the set of $m_k$ devices to selected be scheduled as $\ccalI_k \subseteq \{1,2,\hdots,m\}$ where  $| \ccalI_k| = m_k$ and device $j \in \ccalI_k$ with probability $\nu_{i,k}$. Further define $\ccalS_{(n)} \subset \ccalS$ to be an arbitrary set of $n$ FDs that do not intersect over any frequency bands, i.e. $\sum_{j \in \ccalS_{(n)}} \bbsigma_j \leq \mathbf{1}$. To accommodate the $m_k$ devices to be scheduled, we consider a set of $S$ such sets  $\{\ccalS^s_{(n_s)}\}_{s=1}^S$ with size $n_s$, whose combined elements total $\sum_{s=1}^{S} n_s = m_k$. We define this full set of assignable FDs at cycle $k$ as
%
\begin{align}\label{eq_ru_sets}
\ccalS'_k :=\ccalS^1_{(n_1)} \cup \ccalS^1_{(n_2)} \cup \hdots \cup \ccalS^{S_k}_{(n_{S_k})}.
\end{align}
%
Observe that an FD $\bbsigma$ is further superindexed by its TD slot $s$ y. In this way \eqref{eq_ru_sets} defines a complete set of  combinations of frequency-allocated FD and \emph{time}-allocated TDs to assign users during this cycle. An example of a possible $\ccalS'_k$ for scheduling $m_k = 12$ devices is shown in Table \ref{tab_rus}.

For all $i \in \ccalI_k$ and FD $\bbsigma \in \ccalS'_k$, define the largest affordable DR given the \emph{modified} PDR requirement $\tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k}$ by
 %
 \begin{align}\label{eq_mcs_select}
 \mu_{i,k}(\bbsigma) := \begin{cases}
 \max \{\mu \mid q(\bbh_{i,k},\mu,\bbsigma) \geq \tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k}\} \\
 \mu_0,\quad  \text{ if } q(\bbh_{i,k},\mu,\bbsigma) < \tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k} \ \forall \mu
 \end{cases}
 \end{align}
 %
Observe in \eqref{eq_mcs_select} that, when no DR achieves the desired PDR in a particular FD, this value is set to $\mu=\mu_0$ by default.  The DR defined in \eqref{eq_mcs_select} subsequently incurs a time cost $\tau(\mu_{i,k}(\bbsigma), \bbsigma)$ for assigning device $i$ to FD $\bbsigma$. Define an \emph{assignment} $\ccalV=\{v_{ij}^s\}$ that assigns each device $i \in \ccalI_k$ to an FD/TD pair $(j,s)$ corresponding to an element in $\ccalS'_k$.  For time sensitive applications, the goal is to minimize the total transmission time across all TDs. The transmission time of a single TD is limited by the slowest device (i.e. a TD cannot finish until all devices finish transmitting). Thus, we can write the total transmission time as 
%
\begin{align}\label{eq_assignment}
T(\ccalV) = \sum_{s=1}^S \max_{j} \left[v^s_{ij} \tau(\mu_{i,k}(\bbsigma^s_j), \bbsigma_j^s)\right].
\end{align}
% 
The problem of minimizing $T(\ccalV)$ is a particular version of a non-linear assignment problem, where the goal is to choose an assignment---or schedule---that minimizes transmission time while meeting the control-aware PDR targets in \eqref{eq_pdr_constraint}. These problems are combinatorial in nature and challenging to solve exactly. We may approximate this problem by applying, e.g., the Hungarian method \cite{kuhn1955hungarian}, a well-known method for solving linear-cost assignment problems. Other heuristic assignment approaches may be designed to approximate the solution to \eqref{eq_assignment}. For the simulations performed later in this paper, we apply such a heuristic method, the details of which are left out for proprietary reasons.

The complete control-aware scheduling procedure for low-latency settings is present in Algorithm \ref{alg_calls}. At each cycle $k$, the BS uses current channel states $\bbh_{i,k}$ (obtained via pilot signals) and the current estimated control states $\hbx^{(l_i)}_{i,k}$  (obtained via \eqref{eq_state_est} for each device $i$) to compute control-aware target PDRs  $\tdq_i(\hbx^{(l_i)}_{i,k})$ for each device via \eqref{eq_pdr_constraint} in Step 3. In Step 4, the target PDRs are used to establish selection probabilities $\nu_{i,k}$ for each agent with \eqref{eq_prob_c}. After randomly selecting devices $\ccalI_k$ in Step 5, an appropriate set of FDs and TDs $\ccalS'_k$ are selected in Step 6. In Step 7, the associated DR values are determined for each possible assignment of device to FDs via \eqref{eq_mcs_select}. Finally, in Step 8 the assignment is performed using, e.g., the Hungarian method or other user-designed heuristic assignment method. The resulting assignment determines the scheduling parameters $\{\bbsigma_i, \mu_i, \alpha_i\}$ for all devices $i$ in the current cycle.

### 2019-02-01 Low_latency_control/images/multiplex.tex
Add files via upload

\begin{tikzpicture}
\draw[step=.5cm,gray,thick] (-2,-2) grid (3,2);
\node (u1) at (-2,-2) {};
\draw [ultra thick, draw=black, fill=black!20!white] (-2,-2) rectangle  (0,0);
\node (u2) at (-2,0) {};
\draw [ultra thick, draw=black, fill=red!20!white] (-2,0)  rectangle  (0,1);
\draw [
    thick,
    decoration={
        brace,
        raise=0.1cm
    },
    decorate
] (u1.west) -- (u2.west) 
node [pos=0.5,anchor=north,rotate=90,yshift=.7cm] {RU};
\draw [ultra thick, draw=black, fill=blue!20!white] (-2,1)  rectangle  (0,2);
\draw [ultra thick, draw=black, fill=green!20!white] (0,-2) rectangle (1,2);
\draw [ultra thick, draw=black, fill=cyan!20!white] (1,-2)  rectangle  (3,0);
\draw [ultra thick, draw=black, fill=magenta!20!white] (1,0)  rectangle  (3,2);
\node[align=center, scale =0.9] at (0,-2.5) {Transmission time};
%\node[rotate=90] at (-2.5,0) {Resource units (RU)};
\end{tikzpicture}

### 2019-02-01 Low_latency_control/images/tikz_random_access.tex
Add files via upload

%!TEX root = root.tex
\documentclass[]{standalone}

\usepackage{amsmath} % nice math symbols

\usepackage{color} % change text color        
\usepackage{tikz}
\usetikzlibrary{matrix} % for block alignment
\usetikzlibrary{arrows, decorations.markings} % for arrow heads
\usetikzlibrary{calc} % for manimulation of coordinates
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{shapes}

\usetikzlibrary{fit}

%\usetikzlibrary{backgrounds}

\begin{document}
% TikZ styles for drawing

\pgfdeclarelayer{bg0}    % declare background layer
\pgfdeclarelayer{bg1}    % declare background layer
\pgfsetlayers{bg0,bg1,main}  % set the order of the layers (main is the standard layer)

\tikzstyle{block} = [draw,rectangle,thick,
%minimum height=0.7cm, minimum width=0.3cm, 
text height=0.2cm, text width=0.7cm, 
fill=blue!30, outer sep=0pt, inner sep=0pt]
\tikzstyle{dots} = [font = \large, minimum width=2pt]
\tikzstyle{dash_block} = [draw,rectangle,dashed,minimum height=1cm,minimum width=1cm]
\tikzstyle{smallblock} = [draw,rectangle,minimum height=0.5cm,minimum width=0.5cm,fill= blue!30, font =  \scriptsize]
\tikzstyle{smallcircle} = [draw,ellipse,minimum height=0.1cm,minimum width=0.3cm,fill= yellow!40, font =  \scriptsize ]
\tikzstyle{connector} = [->]
\tikzstyle{dash_connector} = [->,thick,decorate,decoration={snake, amplitude =1pt, segment length=8pt}, magenta]
\tikzstyle{branch} = [circle,inner sep=0pt,minimum size=1mm,fill=black,draw=black]

\tikzstyle{vecArrow} = [thick, decoration={markings,mark=at position
   1 with {\arrow[semithick]{open triangle 60}}},
   double distance=1.4pt, shorten >= 5.5pt,
   preaction = {decorate},
   postaction = {draw,line width=1.4pt, white,shorten >= 4.5pt}]

\tikzset{>=latex}

\begin{tikzpicture}[scale=1, blocka/.style ={rectangle,text width=0.6cm,text height=0.5cm, outer sep=0pt}]
 \small

    % node placement with matrix library: 5x4 array
    \matrix(M)[ampersand replacement=\&, row sep=0.25cm, column sep=0.25cm] {

    \node[smallblock, align=center, fill= green!30] (P1) {Plant {1}};\&

		\node[smallblock, align=center] (S1) {Sensor {1}};\&
		\&
		\&
		\node[smallblock, align=center, fill= green!30] (P2) {Plant {2}};\&

		\node[smallblock, align=center] (S2) {Sensor {2}};\&

%    \&
    %\node(d0) {$\cdots$};\&
%    \&

    \\

    %\node(d1) {$\cdots$};\&
%    \&

    \\
		\\
		\\
		\\
		\\
		\\
		\\

		\&
		\node[smallblock, align=center] (C1) {Controller {1}};\&
		\&
		\&
		\&
		\node[smallblock, align=center] (C2) {Controller {2}};\&

		\\
    %%
    %\node[blocka] (R1) {};\&
%%    \node[smallcircle] (R2) {R2};\&
    %%\node[blocka] (d3) {};\&
    %\node[blocka] (R2) {};\&
    %\\
    };

    %\node[block] (outer) [fit=(R1.north west)  (R2.south east)] {};
    %
    %\node[align=center, font =  \scriptsize] at (outer.center) {Access Point/ \\Controller};

    \draw [->,  red] (S1.south) -- node[left, pos = 0.1]{ $\alpha_1$ } (C1.north);
%    \draw [->, thick, magenta] (T2) -- node[left]{ \scriptsize $h_2$} (R2);
    \draw [->, red] (S2.south) -- node[right, pos = 0.1]{ $\alpha_2$ } (C2.north);
%

		\draw [->, dashed,  red] (S1.south) -- node[right, pos=0.2]{ } (C2.north);

		\draw [->, dashed, red] (S2.south) -- node[left, pos=0.2]{ } (C1.north);

		%\begin{pgfonlayer}{bg0}    % select the background layer
		%\draw [->, black] (C1.west) |- ($(C1) + (-25pt, 0pt)$) node(down_right){} 
		%-- ($(P1) + (-25pt,+0pt)$) node(up_right){} -- (P1);
		%\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}    % select the background layer
		\draw [->, black] (C1.west) -|   (P1.south);
		\end{pgfonlayer}

		%\begin{pgfonlayer}{bg0}
		%\draw [->,  black] (R2) |- ($(R2) + (-25pt,-15pt)$) node(down_right){} 
		%--($(P2) + (-25pt,+0pt)$) node(up_right){} -- (P2);
		%\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}    % select the background layer
		\draw [->, black] (C2.west) -|   (P2.south);
		\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}
		\draw [->, black] (P1) -- (S1);
		\end{pgfonlayer}

		\begin{pgfonlayer}{bg0}
		\draw [->,  black] (P2) -- (S2);
		\end{pgfonlayer}

		%
		\begin{pgfonlayer}{bg1}
		%\begin{scope}[on background layer]
		\node(shared) [fill=red!10, fit={
		($(S1.south) + (-32pt, -12pt)$) 
		($(S2.south) + (+8pt, -12pt)$)
		($(C1.north) + (-32pt, +10pt)$)
		($(C2.north) + (+8pt, +10pt)$)
		}] {};
		%\end{scope}
		\end{pgfonlayer}

		\node[align=center, red!40, font =  \footnotesize](shared_medium) at ($(shared.center)-(50pt, 0)$) {Shared \\ Wireless \\ Medium};

\coordinate (FIRST NE) at (current bounding box.north east);
   \coordinate (FIRST SW) at (current bounding box.south west);

	\pgfresetboundingbox
   \useasboundingbox ($(FIRST SW) + (+52pt,+3pt)$) rectangle ($(FIRST NE) + (-10pt,-2pt)$);

\end{tikzpicture}
\end{document}

### 2019-02-01 Low_latency_control/introduction.tex
Add files via upload

The reliability and robustness of wireless communication systems has become a key component in the design and implementation of large scale systems in the Internet-of-Things (IoT). For these systems to operate successfully, it is important that the data collected by sensors can be communicated throughout the IoT network. While the most reliable form of data communication is through wired connections, the scale and mobility of modern IoT settings has rendered the cost of installing and maintaining a wired network a significant challenge~\cite{wollschlaeger2017future}. Consequently, there is great interest in the design of wireless control systems that can achieve reliable performance~\cite{varghese2014wireless, li2017review}. Low-latency wireless transmissions are however a necessary feature in many wireless IoT systems, particularly those in industrial control~\cite{varghese2014wireless}.  The primary challenge is then the inherent trade-off that occurs between reliability and latency, as it is difficult to maintain high reliability while using higher data rates due to the stochasticity of the wireless channel. It is necessary to design resource allocation and scheduling protocols for wireless control systems that can both meet reliability \emph{and} latency requirements of the control system.

%In the wireless communications research and industry, many radio resource allocation schems in the form of wireless scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}. For instance, round-robin scheduling is a common approach due to its simplicity inherent fairness. Algorithms such as proportional fair (PF)~\cite{kim2005proportional}, max-SNR~\cite{knopp1995information} and their variations have been designed for maximizing the overall system throughput and/or fairness. However, these algorithms can perform very poorly with regards to ensuring that all of the admitted users' packet delivery rates and latency targets will be met. For time-sensitive applications such as industrial control and automation, the experiences of individual users/devices are highly dependent on the ability of the network to deliver their packets within a latency bound. If this objective is not met, the high overall system throughput or throughput-fairness means little to the end users. On the other hand, the generic delay-aware schedulers such as EDF~\cite{wu2014analysis} and WFQ~\cite{lu1999fair} do not properly adapt to wireless channel conditions and may make poor radio resource allocation decisions. These algorithms have also been modified to adapt to wireless channel conditions, such as in M-LWDF \cite{andrews2001providing}. However, all such wireless scheduling algorithms are unable to leverage the information on the control system dynamics and thus, perform inefficiently when deployed in time-sensitive wireless control systems.

In the wireless communications research and industry, many radio resource allocation schemes in the form of wireless scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}. For instance, round-robin scheduling is a common approach due to its simplicity inherent fairness. Algorithms such as proportional fair (PF)~\cite{kim2005proportional}, max-SNR~\cite{knopp1995information} and their variations have been designed for maximizing the overall system throughput and/or fairness. For time-sensitive applications such as industrial control and automation, the experiences of individual users/devices are highly dependent on the ability of the network to deliver their packets within a latency bound. Generic delay-aware schedulers such as EDF~\cite{wu2014analysis} and WFQ~\cite{lu1999fair} do not adapt to wireless channel conditions, while channel state information is considered in M-LWDF \cite{andrews2001providing}. However, all such methods are unable to leverage the information on the control system dynamics and thus can make unsuitable radio resource allocation decisions when deployed in time-sensitive wireless control systems.

In the context of wireless control systems, there have been a range of works that incorporates control system information in the networking and communication policies.
%
The mechanisms usually examined are either static or dynamic. Typical examples of the former type are periodically protocols where the wireless devices transmit in a predefined repeating order, e.g., round-robin. Control system stability under such protocols can be analyzed -- see, e.g.,~\cite{Hespanha_survey, Schenato_foundations, Donkers_switched, Branicky_stability}. Periodic sequences leading to stability~\cite{Hristu_shared_feedback}, controllability and observability~\cite{Hristu_communication_control}, or optimizing control objectives~\cite{LeNy_resource_LQR,Meier_measurement_control, Scheduling_control_combinatorics} have been proposed.
%
Dynamic schedulers do not rely on a predefined sequence but decide access to the communication medium dynamically at each step. %, for example by dynamically assigning priorities to the competing tasks.
Initial approaches abstract control performance requirements in the time/frequency domain, e.g., how often a task needs resource access, employing algorithms from real-time scheduling theory~\cite{Branicky_RM, Liu_Real_time_systems}. More recent scheduling approaches often depend on the current control system states, i.e., informally the subsystem with the largest state discrepancy is scheduled to communicate -- see, e.g.,~\cite{Donkers_switched, Walsh_stability, Hristu_Kumar_interrupt_based, Egerstedt_queue, Hirche_Scheduling_Price, Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
%
Alternatively scheduling can take into account current wireless channel conditions opportunistically to meet target control system reliability requirements~\cite{GatsisEtal15}. None of these approaches, however, are explicitly designed to achieve strong performance it low-latency scenarios.

In this paper, we develop a control-aware, high reliability, low-latency IEEE 802.11ax WiFi protocol~\cite{bellalta2016ieee} that is designed to reduce the total transmission time at each uplink cycle in the control loop. This is done through the mathematical formulation of the control system design goal in the form of a Lyapunov function that ensures stability of the control system. This formulation naturally induces a bound on the packet delivery rate each control system needs to achieve to meet the control-based goal. Such packet delivery rates depend upon current control and channel states and thus dynamically change over the course of the system life time. This can be viewed as an opportunistic protocol with respect to both the current control states \emph{and} channel conditions. Furthermore, because these control-based success rate requirements may be significantly lower than traditional, high reliability communication demands, the proposed method is better suited to find scheduling configurations that can moreover meet the strict latency requirements imposed by the physical system. This is in contrast to the control-aware approach taken in \cite{GatsisEtal15}, which focuses on low-power and infrequent transmissions rather than the low-latency setting of interest in industrial control.

The paper is organized as follows. We formulate the wireless control system in which state information is communicated to the control over a wireless channel. Due to the potential for random packet drops, this is modeled as a switched dynamical system (Section \ref{sec_problem_formulation}). A Lyapunov function is used to evaluate the stability of the control state, and the uncertainty in this measurement grows the more consecutive packets are lost for a particular system.  We then discuss the scheduling parameters of the IEEE 802.11ax communication model (Section \ref{sec_comm_model}).
%, which allows for centralized scheduling of transmissions through the use of varying bandwidth resource units, time-slotted PPDUs, and date rate selection via MCS selection (Section \ref{sec_comm_model}).
From there, we derive a mathematical formulation of the optimal scheduling problem (Section \ref{sec_optimal}). This can be formulation by minimizing a control cost with an explicitly latency constraint (Section \ref{sec_optimal_a}) or minimizing transmission time with an explicit control performance constraint (Section \ref{sec_optimal_b}).

Using this formulation, we develop the control-aware low latency scheduling (CALLS) method (Section \ref{sec_calls}). The CALLS method uses current control states and channel conditions to derive dynamic packet success rates for each user (Section \ref{sec_psr}). In this way, control systems that are closest to instability will be given priority in the scheduling so that they may close their control loops. The scheduling procedure consists of a random user selection procedure to reduce the number of required PPDUs that incur significant overhead  (Section \ref{sec_rss}), followed by an assignment-method based scheduling of selected users to minimize total transmission time (Section \ref{sec_assignment}). The performance of the CALLS method is analyzed in a series of simulation experiments in which its performance is compared against a control-agnostic procedure (Section \ref{sec_simulation}). We demonstrate in numerous low-latency control systems that the control-aware method can support more users than the alternative and achieve more robust overall performance.

%The reliability and robustness of wireless communication systems has become a key component in the design and implementation of large scale systems in the Internet-of-Things (IoT). For these systems to operate successfully, it is important that the data or information collected by sensors can be communicated throughout the IoT network. While the most reliable form of data communication is through wired connections, the scale and mobility of modern IoT settings has rendered the cost of installing and maintaining a wired network a significant challenge~\cite{wollschlaeger2017future}. Consequently, the design of wireless control systems that can achieve reliable performance has become of great interest~\cite{varghese2014wireless, li2017review}. The primary challenge is then the inherent trade off that occurs between reliability and latency. Low-latency transmissions are a necessary feature in many wireless IoT systems, particularly those in industrial control~\cite{varghese2014wireless}. It is difficult to maintain high reliability while using higher data rates, however, due to the stochasticity of the wireless channel. It is necessary then to design resource allocation and scheduling protocols for wireless control systems that can both meet reliability \emph{and} latency requirements of the control system.
%
%
%In the wireless communications research and industry, many scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}. For instance, round-robin scheduling is a common approach due to its simplicity inherent fairness. Algorithms such as proportional fair (PF)~\cite{kim2005proportional}, max-SNR~\cite{knopp1995information} and their variations have been designed for maximizing the overall system throughput and/or fairness. However, these algorithms can perform very poorly with regards to ensuring that all of the admitted users' packet delivery rates and latency targets will be met. For time-sensitive applications such as industrial control and automation, the experiences of individual users/devices are highly dependent on the ability of the network to deliver their packets within a latency bound. If this objective is not met, the high overall system throughput or throughput-fairness means little to the end users. On the other hand, the generic delay-aware schedulers such as EDF~\cite{wu2014analysis} and WFQ~\cite{lu1999fair} do not properly adapt to wireless channel conditions and may make poor radio resource allocation decisions. These algorithms have also been modified to adapt to wireless channel conditions, such as in M-LWDF \cite{andrews2001providing}. However, all such wireless scheduling algorithms are unable to leverage the information on the control system dynamics and thus, perform inefficiently when deployed in time-sensitive wireless control systems.
%
%In the context of wireless control systems, there was been a range or works that incorporates control system information in the networking and communication policies.
%%
%The mechanisms usually examined are either static or dynamic. Typical examples of the first type are periodically protocols where the wireless devices transmit in a predefined repeating order, e.g., round-robin. Control system stability under such protocols can be analyzed -- see, e.g.,~\cite{Hespanha_survey, Schenato_foundations, Donkers_switched, Branicky_stability}. Periodic sequences leading to stability~\cite{Hristu_shared_feedback}, controllability and observability~\cite{Hristu_communication_control}, or optimizing control objectives~\cite{LeNy_resource_LQR,Meier_measurement_control, Scheduling_control_combinatorics} have been proposed.
%%
%The second type of schedulers, the dynamic ones, do not rely on a predefined sequence but decide access to the communication medium dynamically at each step. %, for example by dynamically assigning priorities to the competing tasks.
%Initial approaches abstract control performance requirements in the time/frequency domain, e.g., how often a task needs resource access, employing algorithms from real-time scheduling theory~\cite{Branicky_RM, Liu_Real_time_systems}. More recently scheduling approaches often depend on the current control system states, i.e., informally the subsystem with the largest state discrepancy is scheduled to communicate -- see, e.g.,~\cite{Donkers_switched, Walsh_stability, Hristu_Kumar_interrupt_based, Egerstedt_queue, Hirche_Scheduling_Price, Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
%%
%Alternatively scheduling can take into account current wireless channel conditions opportunistically to meet target control system reliability requirements~\cite{GatsisEtal15}. None of these approaches, however, are explicitly designed to minimize communications latency or achieve strong performance with specific low-latency requirements.
%
%
%
%In this paper, we develop a control-aware, high reliability, low-latency IEEE 802.11ax WiFi protocol~\cite{bellalta2016ieee} that is designed to reduce the total transmission time at each uplink cycle in the control loop. This is done through the mathematical formulation of the control system design goal in the form of a Lyapunov function that ensures stability of the control system. This formulation naturally induces a bound on the packet delivery rate each control system needs to achieve to meet the control-based goal. Such packet delivery rates depend upon current control and channel states and thus dynamically change over the course of the system life time. This can be viewed as an opportunistic protocol with respect to both the current control states \emph{and} channel conditions. Furthermore, because these control-based success rate requirements may be significantly lower than traditional, high reliability communication demands, the proposed method is better suited to find scheduling configurations that can moreover meet the strict latency requirements imposed by the physical system.
%
%The paper is organized as follows. We formulate the wireless control system in which state information is communicated to the control over a wireless channel. Due to the potential for random packet drops, this is modeled as a switched dynamical system (Section \ref{sec_problem_formulation}). A Lyapunov function is used to evaluate the stability of the control state, and the uncertainty in this measurement grows the more consecutive packets are lost for a particular system.  We then discuss the scheduling parameters of the IEEE 802.11ax communication model (Section \ref{sec_comm_model}).
%%, which allows for centralized scheduling of transmissions through the use of varying bandwidth resource units, time-slotted PPDUs, and date rate selection via MCS selection (Section \ref{sec_comm_model}). 
%From there, we derive a mathematical formulation of the optimal scheduling problem (Section \ref{sec_optimal}). This can be formulation by minimizing a control cost with an explicitly latency constraint (Section \ref{sec_optimal_a}) or minimizing transmission time with an explicit control performance constraint (Section \ref{sec_optimal_b}).
%
%Using the control-constrained optimal scheduling formulation, we develop the control-aware low latency scheduling (CALLS) method (Section \ref{sec_calls}). The CALLS method uses current control states and channel conditions to derive dynamic packet success rates for each user (Section \ref{sec_psr}). In this way, control systems that are closest to instability will be given priority in the scheduling so that they may close their control loops. The scheduling procedure consists of a random user selection procedure to reduce the number of required PPDUs that incur significant overhead  (Section \ref{sec_rss}), followed by an assignment-method based scheduling of selected users to minimize total transmission time (Section \ref{sec_assignment}). The performance of the CALLS method is analyzed in a series of simulation experiments in which its performance is compared against a control-agnostic procedure (Section \ref{sec_simulation}). We demonstrate in numerous low-latency control systems that the control-aware method can support more users than the alternative and achieve more robust overall performance.

### 2019-02-01 Low_latency_control/revised/conclusions.tex
Add files via upload

In this paper we proposed a novel control-communication co-design approach to solving the radio resource
allocation problem for time-sensitive wireless control systems. Given a channel state and control state, we mathematically derive a minimum packet delivery
rate a device must meet to maintain a control-orientated target,
as defined by a stability-inducing Lyapunov function. By
dynamically assigning variable packet delivery rate targets
to each device based on its current conditions, we are able to
more easily meet feasibility requirements of a latency-constrained
wireless control problem and maintain stability
and strong performance. We perform simulations on numerous
well-studied low-latency control problems to demonstrate the
benefits of using the control-aware approach, which can include
a 2x gain on number of devices that can be supported.

The results presented in this paper suggest an interesting
potential for control-aware resource allocation and scheduling, particularly in low-latency industrial systems. By considering the control-specific targets such as maintaining
stability or an error margin, we observe that the standard
high reliability targets considered (e.g. packet delivery rates $\geq 0.999$) can in some cases be substantially stricter than
necessary for adequate performance. Wireless control systems
with sufficiently slow dynamics can be kept stable with much
lower packet delivery rates, which in turn make low-latency
communications more achievable. Furthermore, in realistic industrial systems there will be many heterogeneous devices being controlled, whose variation in communication needs is well-served by control-aware opportunism proposed in this paper. This suggests the potential for wireless communications to be adopted using a smart control-communication co-design approach even
while ultra-reliable wireless system technology remains under development.

### 2019-02-01 Low_latency_control/revised/introduction.tex
Add files via upload

--- 2019-02-01 Low_latency_control/introduction.tex
@@
 Consequently, there is great interest in the design of wireless control systems that can achieve reliable performance~\cite{varghese2014wireless, li2017review}.
-Low-latency wireless transmissions are however a necessary feature in many wireless IoT systems, particularly those in industrial control~\cite{varghese2014wireless}.
+Low-latency wireless transmissions are however a necessary feature in many wireless IoT systems, particularly those in industrial control~\cite{varghese2014wireless,popovski2018wireless, bennis2018ultra}.
 The primary challenge is then the inherent trade-off that occurs between reliability and latency, as it is difficult to maintain high reliability while using higher data rates due to the stochasticity of the wireless channel.
 It is necessary to design resource allocation and scheduling protocols for wireless control systems that can both meet reliability \emph{and} latency requirements of the control system.
-
-%In the wireless communications research and industry, many radio resource allocation schems in the form of wireless scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}.
-For instance, round-robin scheduling is a common approach due to its simplicity inherent fairness.
-Algorithms such as proportional fair (PF)~\cite{kim2005proportional}, max-SNR~\cite{knopp1995information} and their variations have been designed for maximizing the overall system throughput and/or fairness.
-However, these algorithms can perform very poorly with regards to ensuring that all of the admitted users' packet delivery rates and latency targets will be met.
-For time-sensitive applications such as industrial control and automation, the experiences of individual users/devices are highly dependent on the ability of the network to deliver their packets within a latency bound.
-If this objective is not met, the high overall system throughput or throughput-fairness means little to the end users.
-On the other hand, the generic delay-aware schedulers such as EDF~\cite{wu2014analysis} and WFQ~\cite{lu1999fair} do not properly adapt to wireless channel conditions and may make poor radio resource allocation decisions.
-These algorithms have also been modified to adapt to wireless channel conditions, such as in M-LWDF \cite{andrews2001providing}.
-However, all such wireless scheduling algorithms are unable to leverage the information on the control system dynamics and thus, perform inefficiently when deployed in time-sensitive wireless control systems.
 

### 2019-02-01 Low_latency_control/revised/scheduling_alg.tex
Add files via upload

We develop a control-aware low-latency scheduling (CALLS) algorithm to approximately solve the control-constrained scheduling formulation in \eqref{eq_problem2}-\eqref{eq_c42}. Because this problem is combinatorial in nature, it is infeasible to solve exactly. Instead, we focus on a practical and efficient means of solving approximately. In particular, we identify sets of feasible points and use a heuristic approach towards minimizing the transmission time objective among the set of feasible points. Additionally, within the development of the CALLS method we identify and characterize new PDR requirements that are defined relative to the control system requirements; these are generally significantly less strict than the PDR requirements often considered in general high reliability communication systems without codesign. Overall, the CALLS method consists of (i) the derivation of dynamic control-based PDR targets, (ii) a principled random selection of devices to schedule to reduce latency, and (iii) the use of assignment based methods to find a low-latency schedule. We discuss these three components in detail in the proceeding subsections.

\subsection{Control-based dynamic PDR}\label{sec_psr}
Due to the complexity of the scheduling problem in \eqref{eq_problem2}-\eqref{eq_c42}, we first focus our attention on identifying scheduling parameters $\{\bbSigma_k, \bbmu_k, \bbalpha_k\}$ that are feasible, i.e. satisfy the constraints in \eqref{eq_c12}-\eqref{eq_c42}. In particular, the Lyapunov control constraint in \eqref{eq_c22} is of significant interest. Recall that the control cost function $J_i(\hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i)$ is itself determined by the PDR $q(\bbh_{i,k},\mu_i,\bbsigma_i)$, as per \eqref{eq_lyap_p}. Thus, the constraint in \eqref{eq_c22} can be seen as indirectly placing a constraint on the required PDR necessary to achieve a $\rho_i$-rate decrease in expectation. The equivalent condition on PDR $q(\bbh_{i,k},\mu_i,\bbsigma_i)$ is presented in the following proposition.

%%%%%%%%%%%%%%%%%%%
\begin{proposition}\label{prop_pdr_constraint}
Consider the Lyapunov control constraint in \eqref{eq_c22} and the definition of $J_i(\hbx^{(l_i)}_{i,k},\bbh_{i,k},\mu_i,\bbsigma_i)$ given in \eqref{eq_lyap_p}.  Define the closed-loop state transition matrix $\bbA^c_i := \bbA_i + \bbB_i \bbK_i$ and $j$-accumulated noise $\omega_i^j := \Tr[(\bbA_i^T\bbP^{1/j} \bbA_i)^{j} \bbW_i]$. The control constraint in \eqref{eq_c22} is satisfied for device $i$ if and only if the following condition on PDR $q(\bbh_{i,k},\mu_i,\bbsigma_i)$ holds, 
%
\begin{align}\label{eq_pdr_constraint}
&q(\bbh_{i,k},\mu_i,\bbsigma_i) \geq \tdq_i(\hbx^{(l_i)}_{i,k}) :=  \\ &\ \frac{1}{\Delta_i} \left[  \left\| (\bbA^c_i - \rho_i\bbI)\hbx^{(l_i)}_{i,k} \right\|_{\bbP^{\frac{1}{2}}}^2+ (1-\rho_i)\sum_{j=0}^{l_i-1} \omega_i^j + \omega_i^{l_i} - c_i  \right],  \nonumber
\end{align}
%
where we have further defined the constant
%
\begin{equation}\label{eq_noise_diff}
\Delta_i :=  \sum_{j=0}^{l_i-1}[\omega^{j+1}_i - \Tr(\bbA^{cT}_i (\bbA_i^T\bbP^{1/j}\bbA_i)^j \bbA^c_i\bbW_i)  ].
\end{equation}
\end{proposition}
%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
%%%% P R O O F %%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%
\begin{myproof}
 Consider the Lyapunov decrease constraint as written in \eqref{eq_c22}. As the same logic holds for all $i$ and $k$, for ease of presentation we remove all subscripts when presenting the details of this proof. We further introduce the simpler notation $q := q(\bbh,\mu,\bbsigma)$. Now, we may expand the left hand side of \eqref{eq_c22} be rewriting the definition in \eqref{eq_lyap_p} as
%
\begin{align}
J(\hbx^{(l)},\bbh,\mu,\bbsigma)  &= q \E_{\bbw}[ L(\bbA_c \bbx + \bbw)] \label{eq_p1}\\
&\quad + (1-q)  \E_{\bbw} [L(\bbA \bbx + \bbB\bbK\hbx + \bbw)]. \nonumber 
\end{align}
%
Recall the definition of the quadratic Lyapunov function $L(\bbx) := \bbx^T \bbP \bbx$ for some positive definite $\bbP$. Further recall the relation $\bbx = \hbx + \bbe$ as described by \eqref{eq_diff}. Combining these, we expand the right hand size of \eqref{eq_p1} as
%
\begin{align}
&J(\hbx^{(l)},\bbh,\mu,\bbsigma) = \label{eq_p2}\\
&\quad q \E_{\bbw} \left[ \bbA_c  \left(\hbx +\bbe \right) + \bbw \right]^T \bbP \left[ \bbA_c  \left(\hbx +\bbe \right) + \bbw \right]  \nonumber\\
&\ +  (1-q) \E_{\bbw} \left[ \bbA_c \hbx + \bbA\bbe +  \bbw \right]^T \bbP \left[ \bbA_c \hbx +\bbA\bbe +  \bbw\right]. \nonumber
\end{align}
%
To evaluate the expectations in \eqref{eq_p2}, recall the random noise $\bbw$ follows a Gaussian distribution with zero mean and covariance $\bbW$. Thus, the expectation can be evaluated over $\bbw$ and expanded as
%
\begin{align}
&J(\hbx^{(l)},\bbh,\mu,\bbsigma)= \\
&q \left[  \| \bbA_c \hbx\|^2_{\bbP^{\frac{1}{2}}} + \Tr(\bbP\bbW) + \sum_{j=0}^{l-1} \Tr( \bbA_c (\bbA^T\bbP^{\frac{1}{j}} \bbA)^j \bbA_c \bbW)\right] +  \nonumber\\
&   (1-q)\left[   \| \bbA_c \hbx\|^2_{\bbP^{\frac{1}{2}}} + \Tr(\bbP \bbW) + \sum_{j=1}^{l} \Tr((\bbA^T\bbP^{\frac{1}{j}} \bbA)^{j}\bbW) \right].\nonumber
\end{align}
%
From here, we rearrange terms and substitute the notation $\omega^j := \Tr[(\bbA^T\bbP^{1/j}\bbA)^{j} \bbW]$ to obtain that the control cost can be written as
%
\begin{align}
J(\hbx^{(l)},\bbh,\mu,\bbsigma)&=\left[  \| \bbA_c \hbx\|^2_{\bbP^{\frac{1}{2}}}+ \Tr(\bbP \bbW) + \sum_{j=1}^{l}\omega^{j} \right]  \label{eq_p4} \\
&\ + q  \sum_{j=0}^{l-1}[ \Tr( \bbA_c (\bbA^T\bbP^{\frac{1}{j}}\bbA)^j \bbA_c \bbW) - \omega^{j+1}]. \nonumber
\end{align}
%
With \eqref{eq_p4}, we have expanded the control cost in terms of the PDR $q$. Now, we return to the constraint in \eqref{eq_c22}. Recall the expansion for $\E [L(\bbx) \mid \hbx^{(l)}]$ via \eqref{eq_perf}. By combining this with the expansion in \eqref{eq_p4}, the terms in\eqref{eq_c22} can be rearranged to obtain the inequality in \eqref{eq_pdr_constraint}.
\end{myproof}
%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
%%%% %%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%

In Proposition \ref{prop_pdr_constraint} we establish a lower bound $\tdq_i(\hbx^{(l_i)}_{i,k})$ on the PDR of device $i$ that is dependent upon the current estimated state $\hbx^{(l_i)}_{i,k}$ and system dynamics determined by $\bbA^c_i, \bbA_i$, and $\bbW^i$. We may note the following intuitions about the constraint in \eqref{eq_pdr_constraint}. The PDR condition naturally grows stricter as the bound $\tdq_i(\hbx^{(l_i)}_{i,k})$ defined on the right hand side of \eqref{eq_pdr_constraint} gets larger. The first term on the right hand side reflects the current estimated channel state, and will become larger as the state gets larger. Similarly, the latter two terms on the right hand side together reflect the size of the noise that has accumulated by operating in open loop. When the noise variance $\bbW_i$ is high and when the last-update counter $l_i$ is large, these latter two noise terms will both be large. Thus, both the current magnitude of the control state and the growing uncertainty from infrequent transmissions together determine how large is the PDR requirement in \eqref{eq_pdr_constraint}.

We stress the value of the PDR condition in \eqref{eq_pdr_constraint} is both in its adaptability to the control system state and dynamics, as well as its identification of precise target delivery rates that are necessary to keep the control systems moving towards stability on average. Depending on the particular system dynamics as described in \eqref{eq_control_orig}, such PDR's may be, and often are considerably more lenient than the default target transmission success rates used in practical wireless systems (e.g. $q = 0.999$). Thus, through \eqref{eq_pdr_constraint} we make a claim that, with knowledge of the control system dynamics and targeted \emph{control performance}, we can effectively soften the targeted \emph{communication performance}---or ``reliability''--- accordingly to something more easily obtained in low-latency constrained systems.

\begin{remark}\normalfont
It is worthwhile to note that by placing a stricter Lyapunov decrease constraint with smaller  rate $\rho_i$ in \eqref{eq_c22}, then the first term on the right hand side of \eqref{eq_pdr_constraint} also grows larger and increases the necessary PDR. Generally, selecting a smaller $\rho$ will result in a faster convergence to stability but will require stricter communication requirements. In fact, we may use the inherent bound on the probability $q(\bbh_{i,k},\mu_i,\bbsigma_i) \leq 1$ to find a lower bound on the Lyapunov decrease rate $\rho_i$ that can be feasibly obtained based upon current control state and system dynamics. This bound, however, may not be obtainable in practice due to the scheduling constraints. In practice, we select $\rho_i$ to be in the interval $[0.90,0.1)$. 
\end{remark}

\subsection{Selective scheduling}\label{sec_rss}
We now proceed to describe the procedure with which we can find a set of feasible scheduling decisions $\{\bbSigma_k, \bbmu_k, \bbalpha_k\}$. To begin, we first consider a stochastically \emph{selective scheduling} protocol, whereby we do not attempt to schedule every device at each transmission cycle, but instead select a subset to schedule a principled random manner. Define by $\nu_{i,k} \in [0,1]$ the probability that device $i$ is included in the transmission schedule at time $k$ and further recall by $q(\bbh_{i,k},\mu_i,\bbsigma_i)$ to be the packet delivery rate with which it transmits. Then, we may consider the \emph{effective} packet delivery rate $\hat{q}$ as 
%
\begin{align}\label{eq_effective_pdr}
\hat{q}(\bbh_{i,k},\mu_i,\bbsigma_i) = \nu_{i,k} q(\bbh_{i,k},\mu_i,\bbsigma_i)
\end{align}
%
Selective scheduling is motivated by the ultimate goal of minimizing total transmit time as described in the objective in \eqref{eq_problem2}. As we consider a large number of total devices $m$, scheduling all such devices will require a larger number of PPDU slots---a maximum of 9 devices can transmit within a single PPDU. Recall in \eqref{eq_time_slot} that each additional PPDU requires unavoidable overhead in $\tau_0$, which in aggregation over multiple PPDUs may become a significant bottleneck in minimizing $\hat{\tau}$ or meeting a strict latency requirement $\tau_{max}$. Thus, by decreasing the amount of scheduled devices, we may decrease the number of total PPDUs and the overhead that is added to the total transmission time.

Observe that by introducing the term $\nu_i$ to the evaluation of effective PDR $\tdq_i$ in \eqref{eq_effective_pdr}, we would thus need to transmit with higher PDR $q(\bbh_{i,k},\mu_i,\bbsigma_i) \geq \tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k}$ to meet the condition in \eqref{eq_pdr_constraint}. While imposing a tighter PDR requirement will indeed require longer transmission times, this added time cost is generally less than the transmission overhead of additional PPDUs. In this work, we use the determine scheduling probability of device $i$ through its PDR requirement $\tdq_i(\hbx^{(l_i)}_{i,k})$ as 
%
\begin{align}\label{eq_prob_c}
\nu_{i,k} := e^{\tdq_i(\hbx^{(l_i)}_{i,k})-1} .
\end{align}
%

With \eqref{eq_prob_c}, the probability of scheduling device $i$ increases as the required PDR increases. Notice that, when a transmission is required, i.e. $\tdq_i(\hbx^{(l_i)}_{i,k}) = 1$, then device $i$ is included in the scheduling with probability $1$. In general, devices with very high PDR requirements, e.g. $>0.99$, will be scheduled with very high probability. Thus, the transmission time gains that are provided through selective scheduling using \eqref{eq_prob_c} would be minimal, if non-existent, in high-reliability settings in which PDR requirements remain high at all times. However, with the lower PDR requirement obtained through the control-aware scheduling in \eqref{eq_pdr_constraint}, selective scheduling as the potential to create significant time savings, as will be later shown in Section \ref{sec_simulation} of this paper.

\subsection{Assignment-based scheduling}\label{sec_assignment}

%We now proceed to discuss how the PDR requirements previously derived are used to schedule the devices during a TXOP. Rather than employing a greedy method as is commonly done in wireless scheduling problems, in the proposed method we use assignment-type methods. To begin, we must determine a set of schedules that satisfy the constraints in \eqref{eq_c12}-\eqref{eq_c42}. Consider that device $i$ is selected to be scheduled at cycle $k$ with probability $\nu_{i,k}$ and define the set of devices to be scheduled as $\ccalI_k \subseteq \{1,2,\hdots,m\}$ with size $| \ccalI_k|$. To specify the sets of RUs that we consider in our scheduling, we first define some notation necessary in the description. We first define $\hat{\ccalS} \subset \ccalS$ to be an arbitrary set of RUs that do not intersect over any frequency bands (i.e. satisfy the constraint in \eqref{eq_c12}). We further define $\hat{\ccalS}^s_n$ to be such a set with size $|\hat{\ccalS}_n| = n$ for PPDU slot $s$, with elements labeled as $\bbsigma_1^s,\hdots,\bbsigma_n^s$. For example, the set $\hat{\ccalS}^s_9$ includes 9 RUs, each containing a single different frequency band of 2MHz. Conversely, the set $\hat{\ccalS}^s_2$ will contain two non-intersecting RU's of size 8MHz. In general, the number of RU and PPDU combinations of size $|\ccalI_k|$ is very large and combinatorial in nature, thus making it infeasible to identify all such combinations. For the reasons mentioned above regarding overhead, in our algorithm we first keep $S_k$, i.e. the number of PPDUs in scheduling cycle $k$, small. We set $S_k = \lceil |\ccalI_k / 9| \rceil$, where the first $S_k - 1$ PPDUs slot 9 devices in the RUs in $\hat{\ccalS}_9$, with the remaining devices slotted into the $S_k$th PPDU allocate the RUs in $\hat{\ccalS}_{(|\ccalI_k| \mod 9)}$. The complete set of $|\ccalI_k|$ RU possibilities at time $k$ is notated as 
%\begin{align}\label{eq_ru_sets}
%\ccalS'_k :=\hat{\ccalS}^1_9 \cup \hat{\ccalS}^2_9 \cup \hdots \cup \hat{\ccalS}^{S_k-1}_9 \cup \hat{\ccalS}^{S_k}_{(|\ccalI_k| \mod 9)}.
%\end{align}

We now proceed to discuss how the PDR requirements previously derived are used to schedule the devices during a TXOP. Rather than employing a greedy method as is commonly done in wireless scheduling problems, in the proposed method we use assignment-type methods. In such assignment-type methods, we assign all scheduled devices to a PPDU and RU at the beginning of the TXOP rather than make scheduling decisions after each PPDU. To begin, we must determine a set of schedules that satisfy the constraints in \eqref{eq_c12}-\eqref{eq_c42}. Recall each device $i$ is selected to be scheduled at cycle $k$ with probability $\nu_{i,k}$ and define the set of $m_k$ devices to selected be scheduled as $\ccalI_k \subseteq \{1,2,\hdots,m\}$ where  $| \ccalI_k| = m_k$. To specify the sets of RUs that we consider in our scheduling, we first define some notation necessary in the description. We define $\hat{\ccalS}_{(n)} \subset \ccalS$ to be an arbitrary set of RUs that do not intersect over any frequency bands (i.e. satisfy the constraint in \eqref{eq_c12}) with exactly $n$ elements. To accommodate the $m_k$ devices to be scheduled, we consider a set of $S_k$ such sets  $\hat{\ccalS}_{(n_s)}$ with size $n_s$, whose combined elements total $\sum_{s=1}^{S_k} n_s = m_k$. In other words, we identify a set $S_k$ PPDUs in which the $s$th PPDU contains $n_s$ non-intersecting PPDUs. We define this full set of assignable RUs at cycle $k$ as
%
\begin{align}\label{eq_ru_sets}
\ccalS'_k :=\hat{\ccalS}^1_{(n_1)} \cup \hat{\ccalS}^1_{(n_2)} \cup \hdots \cup \hat{\ccalS}^{S_k}_{(n_{S_k})}.
\end{align}
%

Note that in \eqref{eq_ru_sets} we further superindex each set by a PPDU index $s$, in order to stress that elements are distinct between sets. That is, an RU $\bbsigma$ present in sets $\hat{\ccalS}^x_{(n_x)}$ and $\hat{\ccalS}^y_{(n_y)}$ is considered as two distinct elements in $\ccalS'_k$, denoted $\bbsigma^x$ and $\bbsigma^y$, respectively. In this way \eqref{eq_ru_sets} defines a complete set of  combinations of frequency-allocated RU and \emph{time}-allocated PPDUs to assign users during this cycle. We point out that there are numerous ways in which to define such sets of RUs in each PPDU that total $m_k$ assignments. There are various heuristic methods that may be employed to quickly identify a permissible assignment pool $\ccalS'_k$, and various simple heuristics may be developed to make this selection in a manner that reduces the overall latency of the transmission window. An example of the set $\ccalS'_k$ for scheduling $m_k = 14$ devices is shown in Table \ref{tab_rus}.

%%%%
\begin{table}[]
\centering
\begin{tabular}{|c|c|c|}
\hline 
\textbf{PPDU 1} & \textbf{PPDU 2} & \textbf{PPDU 3}        \\ \hline \hline
\multicolumn{1}{|c|}{RU 1} & \multicolumn{1}{c|}{\multirow{2}{*}{RU 10}} & \multicolumn{1}{c|}{\multirow{4}{*}{RU 13}} \\ \cline{1-1}
\multicolumn{1}{|c|}{RU 2} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \cline{1-2}
\multicolumn{1}{|c|}{RU 3} & \multicolumn{1}{c|}{\multirow{2}{*}{RU 11}} & \multicolumn{1}{c|}{}                       \\ \cline{1-1}
\multicolumn{1}{|c|}{RU 4} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \hline
\multicolumn{1}{|c|}{RU 5} & \multicolumn{1}{c|}{\multirow{4}{*}{RU 12}} & \multicolumn{1}{c|}{\multirow{4}{*}{RU 14}} \\ \cline{1-1}
\multicolumn{1}{|c|}{RU 6} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \cline{1-1}
\multicolumn{1}{|c|}{RU 7} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \cline{1-1}
\multicolumn{1}{|c|}{RU 8} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \hline
\multicolumn{1}{|c|}{RU 9} & \multicolumn{1}{c|}{}                       & \multicolumn{1}{c|}{}                       \\ \hline
\end{tabular}
\caption{Example of RU selection with $m_k= 14$ devices. There are a total of $S_k = 3$ PPDUs, given $n_1=9$, $n_2 = 3$, $n_3 = 2$ RUs, respectively.}
\label{tab_rus}
\end{table}
%%%%%

For all $i \in \ccalI_k$ and RU $\bbsigma \in \ccalS'_k$, define the largest affordable MCS given the \emph{modified} PDR requirement $\tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k}$ by
 %
 \begin{align}\label{eq_mcs_select}
 \mu_{i,k}(\bbsigma) := \begin{cases}
 \max \{\mu \mid q(\bbh_{i,k},\mu,\bbsigma) \geq \tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k}\} \\
 1,\quad  \text{ if } q(\bbh_{i,k},\mu,\bbsigma) < \tdq_i(\hbx^{(l_i)}_{i,k})/\nu_{i,k} \ \forall \mu
 \end{cases}
 \end{align}
 %
Observe in \eqref{eq_mcs_select} that, when no MCS achieves the desired PDR in a particular RU, this value is set to $\mu=1$ by default. The above adaptive MCS selection can be achieved based on channel conditions using the techniques outlined in~\cite{hoefel2016application}. This MCS selection subsequently then yields a corresponding time cost $\tau(\mu_{i,k}(\bbsigma), \bbsigma)$ for assigning device $i$ to RU $\bbsigma$. Further define an 3-D assignment tensor $V$---where $v^s_{ij}= 1$ when device $i$ is assigned to RU $\bbsigma^s_j$ and 0 otherwise---and $\ccalV$ as the set of all possible assignments. Recalling the form of the total transmission time given PPDU arrangements in \eqref{eq_time_slot}, the assignment that minimizes total transmission time is given by
%
\begin{align}\label{eq_assignment}
V^* = \argmin_{V \in \ccalV} \sum_{s=1}^S \max_{j} \left[v^s_{ij} \tau(\mu_{i,k}(\bbsigma^s_j), \bbsigma_j^s)\right].
\end{align}
%

The expression in \eqref{eq_assignment} can be identified as a particular form of the \emph{assignment problem}, a common combinatorial optimization problem in which the selection of mutually exclusive assignment of agents to tasks incurs some cost. Here, the cost is the total transmission time across all PPDUs necessary for scheduled devices to meet the target PDRs. Assignment problems are generally very challenging to solve---there are $m_k!$ combinations---although polynomial-time algorithms exist for simple cases. The Hungarian method \cite{kuhn1955hungarian}, for example, is a standard method for solving linear-cost assignment problems. While the cost we consider in \eqref{eq_assignment} is nonlinear, the Hungarian method may be used as an approximation. Alternatively, other heuristic assignment approaches may be designed to approximate the solution to \eqref{eq_assignment}. We note that, for the simulations performed later in this paper, we apply such a heuristic method, the details of which are left out for proprietary reasons.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%   A   L   G   O   R   I   T   H   M   %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%{\linespread{1.3}
%\begin{algorithm}[t] \begin{algorithmic}[1]
%\red{
%\STATE \textbf{Parameters:} Lyapunov decrease rate $\rho$
%\STATE \textbf{Input:} Initial states $\bbx_{i,0}$
%\FOR [main loop]{$k = 1,2,\hdots$}
%      \STATE Obtain current estimates of control states $\hbx^{(l_i)}_{i,k}$ for each device $i$ [cf. \eqref{eq_state_est}].
%      \STATE Obtain channel measurements $\bbH_k$ via pilot signals.
%      \STATE Compute target PDR $\tdq_i(\hbx^{(l_i)}_{i,k})$ for each device $i$ [cf. \eqref{eq_pdr_constraint}].
%      \STATE Select devices with prob. $\nu_{i,k} = e^{\tdq_i(\hbx^{(l_i)}_{i,k})-1}$ [cf. \eqref{eq_prob_c}]
%      \STATE Schedule selected devices via Algorthim \ref{alg_hungarian} or \ref{alg_maxmin}.
%\ENDFOR}
%\end{algorithmic}
%\caption{Hungarian Method \cite{kuhn1955hungarian}}\label{alg_hungarian} \end{algorithm}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{method}[Linear approximation via Hungarian Method]\label{method_hungarian} \normalfont
%While efficient methods to solve general assignment problems do not exist, the special case of linear cost assignment problems can be solved exactly in polynomial time. These cannot be applied directly to \eqref{eq_assignment} due to the nonlinearities introduced by the maximization operation in \eqref{eq_time_slot}. Consider, however, the following linear-cost surrogate assignment problem
%%
%\begin{align}\label{eq_assignment_linear}
%V^* = \argmin_{V \in \ccalV} \sum_{s=1}^S \sum_{j} \left[v_{ijs} \tau(\mu_{i,k}(\bbsigma^s_j), \bbsigma_j^s)\right].
%\end{align}
%%
%In \eqref{eq_assignment_linear}, we select the assignment that minimizes the sum of all individual transmission times, regardless of PPDU. With this approximation, we can perform the well-known Hungarian method \cite{kuhn1955hungarian}, which is known to solve linear-cost problems in $\ccalO(| \ccalI_k|^3)$ iterations \cite{edmonds1972theoretical}. While the cubic rate here may indeed become prohibitive when a very large number of devices are scheduled, the selective scheduling procedure will generally keep this runtime from becoming too large. We present in Algorithm \ref{alg_hungarian} the use of the Hungarian method on the linearly approximated scheduling problem.
%\end{method}
%%
%%
%\begin{method}[Max-min heursitic method]\label{method_maxmin} \normalfont
%\blue{Add here. Or do not include?}
%\end{method}
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%   A   L   G   O   R   I   T   H   M   %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\linespread{1.3}
\begin{algorithm}[t] \begin{algorithmic}[1]
\STATE \textbf{Parameters:} Lyapunov decrease rate $\rho$
\STATE \textbf{Input:} Channel conditions $\bbH_k$ and estimated states $\hbX_{k}$
\STATE Compute target PDR $\tdq_i(\hbx^{(l_i)}_{i,k})$ for each device $i$ [cf. \eqref{eq_pdr_constraint}].
\STATE Determine selection probabilities $\nu_{i,k}$ for each device [cf. \eqref{eq_prob_c}].
\STATE Select devices $\ccalI_k$ with probs. $\{\nu_{1,k},\hdots,\nu_{m,k}\}$
\STATE Determine set of RUs/PPDUs $\ccalS'_k$ [cf. \eqref{eq_ru_sets}].
\STATE Determine maximum MCS for each device/RU assignment [cf. \eqref{eq_mcs_select}].
\STATE Schedule selected devices via assignment method.
\STATE \textbf{Return:} Scheduling variables $\{\bbSigma_k, \bbmu_k, \bbalpha_k\}$ 
\end{algorithmic}
\caption{Control-Aware Low Latency Scheduling (CALLS) at cycle $k$}\label{alg_calls} \end{algorithm}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

By combining these methods with the control-based PDR targets and selective scheduling procedure, we obtain the complete control-aware low-latency scheduling (CALLS) algorithm. The steps as performed by the centralized AP/controller are outlined in Algorithm \ref{alg_calls}. At each cycle $k$, the AP determines the scheduling parameters based on the current channel states $\bbH_k$ (obtained via pilot signals) and the current estimated control states $\hbX_k$ (obtained via \eqref{eq_state_est} for each device $i$). With the current state estimates, the AP computes target PDRs  $\tdq_i(\hbx^{(l_i)}_{i,k})$ for each device via \eqref{eq_pdr_constraint} in Step 3. In Step 4, the target PDRs are used to establish selection probabilities $\nu_{i,k}$ for each agent with \eqref{eq_prob_c}. After randomly selecting devices $\ccalI_k$ with their associated probabilities in Step 5, the set of RUs and PPDUs $\ccalS'_k$ are determined in Step 6 as in \eqref{eq_ru_sets} , based upon the number of devices selected to be scheduled $|\ccalI_k|$. In Step 7, the associated MCS values are determined each possible assignment of device to RU via \eqref{eq_mcs_select}. Finally, in Step 8 the assignment is performed using either the Hungarian method \cite{kuhn1955hungarian} or other user-designed heuristic assignment method. The resulting assignment determines the scheduling parameters $\bbSigma_k, \bbmu_k, \bbalpha_k$ for the current cycle.

 \begin{remark}\label{remark_latency}\normalfont
 Observe that the CALLS method as outlined in Algorithm \ref{alg_calls} seeks to minimize the total latency of the transmission but does not explicitly prevent latency from exceeding some specific threshold $\tau_{\max}$. In practical systems, this limit may need to be enforced. In such a setting, the CALLS method can be modified so that all devices scheduled in PPDUs whose transmission end after $\tau_{\max}$ seconds do not transmit. 
 \end{remark}

### 2019-02-01 scheduling_policy.py
Delete scheduling_policy.py

@@ -1,244 +1,0 @@
-import numpy as np
-import tensorflow as tf
-import math
-import scipy
-import scipy.stats
-
-
-# slow implmenetation of running average
-class RunningStats(object):
-    def __init__(self, N):
-        self.N = N
-        self.vals = []
-        self.num_filled = 0
-
-    def push(self, val):
-        if self.num_filled == self.N:
-            self.vals.pop(0)
-            self.vals.append(val)
-        else:
-            self.vals.append(val)
-            self.num_filled += 1
-
-    def push_list(self, vals):
-        num_vals = len(vals)
-
-        self.vals.extend(vals)
-        self.num_filled += num_vals
-        if self.num_filled >= self.N:
-            diff = self.num_filled - self.N
-            self.num_filled = self.N
-            self.vals = self.vals[diff:]
-
-    def get_mean(self):
-        return np.mean(self.vals[:self.num_filled])
-
-    def get_std(self):
-        return np.std(self.vals[:self.num_filled])
-
-    def get_mean_n(self, n):
-        start = max(0, self.num_filled-n)
-        return np.mean(self.vals[start:self.num_filled])
-
-
-
-
-class SchedulingPolicy(object):
-    def __init__(self, state_dim, action_dim, constraint_dim, sess=None):
-        self.state_dim = state_dim
-        self.action_dim = action_dim
-        self.constraint_dim = constraint_dim
-        self.lambd = np.ones((constraint_dim, 1))
-
-
-
-        self.stats = RunningStats(64*100)
-
-        self._build_model(state_dim, action_dim)
-
-        if sess == None:
-            self.sess = tf.InteractiveSession()
-            tf.global_variables_initializer().run() 
-        else:
-            self.sess = sess
-
-
-    def _build_model(self, state_dim, action_dim):
-        with tf.variable_scope('policy'):
-            self.state_input = tf.placeholder(tf.float32, [None, state_dim], name='state_input')
-
-
-            layer1 = tf.contrib.layers.fully_connected(self.state_input,
-                32,
-                activation_fn=tf.nn.relu,
-                scope='layer1')
-
-            layer2 = tf.contrib.layers.fully_connected(layer1,
-                16,
-                activation_fn=tf.nn.relu,
-                scope='layer2')
-
-
-            # # gaussian distribution
[…]

### 2019-02-01 datatools.py
Delete datatools.py

@@ -1,332 +1,0 @@
-import gensim
-import sklearn, sklearn.datasets
-import sklearn.naive_bayes, sklearn.linear_model, sklearn.svm, sklearn.neighbors, sklearn.ensemble
-import matplotlib.pyplot as plt
-import scipy.sparse
-import numpy as np
-import time, re
-
-def indices_to_onehot(a,n):
-	n_samples = a.size # Number of samples
-	E = np.zeros((n,n_samples)) # Deltas to diffuse
-	for it in range(n_samples):
-		E[(a[it]-1).astype(int),it] = 1.
-	return E
-
-def create_samples(S,a,K):
-	n = S.shape[0] # Dimension of x
-	n_samples = a.size # Number of samples
-	E = indices_to_onehot(a,n)
-	t = np.floor(K*np.random.rand(n_samples)) # Diffusion times
-	# Storage:
-	x = np.zeros((n,n_samples))
-	#!!!
-	for it in range(n_samples):
-		St = np.linalg.matrix_power(S,t[it].astype(int))
-		x[:,it] = St.dot(E[:,it])
-	return x
-
-def my_eig(S):
-	d,V = np.linalg.eig(S)
-	order = np.argsort(-d)
-	d = d[order]
-	V = V[:,order]
-	D = np.diag(d)
-	VV = np.linalg.inv(V)
-	SS = V.dot(D.dot(VV))
-	diff = np.absolute(S-SS)
-	if diff.max() > 1e-6:
-		print("Eigendecomposition not good enough")
-	return V,D
-	
-# def print_matrix(X,t=3):
-# 	"""
-# 	Print matrix
-# 	X: matrix
-# 	t: number of decimals
-# 	"""
-# 	N,M = X.shape
-# 	assert t <= 5
-# 	print(' ')
-# 	#print('[', end = ' ', flush = True)
-# 	for i in range(N):
-# 		if i >= 1:
-# 			print(' ', end = ' ', flush = True)
-# 		for j in range(M):
-# 			if t == 0:
-# 				print('{:4.0f}'.format(X[i][j]), end = ' ', flush = True)
-# 			elif t == 1:
-# 				print('{:5.1f}'.format(X[i][j]), end = ' ', flush = True)
-# 			elif t == 2:
-# 				print('{:6.2f}'.format(X[i][j]), end = ' ', flush = True)
-# 			elif t == 3:
[…]

### 2019-02-01 graphtools.py
Delete graphtools.py

@@ -1,783 +1,0 @@
-import numpy as np
-import scipy.sparse
-import sklearn.metrics
-import sklearn.neighbors
-import matplotlib.pyplot as plt
-import scipy.sparse.linalg
-import scipy.spatial.distance
-import sys
-import datatools
-import tensorflow as tf
-
-def create_ER(n,p):
-	# Function provided in MATLAB(R) by Dr. S. Segarra.
-
-	FLAG_CONNECTED = 0
-	while FLAG_CONNECTED == 0:
-		G = np.random.rand(n,n)
-		G = G < p
-		G = G.astype(float)
-		G = np.triu(G,1)
-		G = G + G.conj().T
-		L = np.diag(np.sum(G,axis=0))-G
-		D,_ = np.linalg.eig(L)
-		n_connected_components = (np.absolute(D)<=1e-6)
-		if np.sum(n_connected_components.astype(float)) == 1.:
-			FLAG_CONNECTED = 1
-	return G
-
-def create_SBM(n,p):
-	n_c = n.shape[0] # Number of communities
-	FLAG_CONNECTED = 0
-	P = np.zeros([np.sum(n),np.sum(n)])
-	for it1 in range(n_c):
-		P_row1 = np.sum(n[0:it1])
-		P_row2 = np.sum(n[0:(it1+1)])
-		P[P_row1:P_row2,P_row1:P_row2] = np.tile(
-				p[it1,it1],(n[it1,0],n[it1,0]))
-		for it2 in range(it1+1,n_c):
-			P_col1 = np.sum(n[0:it2])
-			P_col2 = np.sum(n[0:(it2+1)])
-			P[P_row1:P_row2,P_col1:P_col2] = np.tile(
-					p[it1,it2], (n[it1,0],n[it2,0]))
-			P[P_col1:P_col2,P_row1:P_row2] = np.tile(
-					p[it1,it2], (n[it2,0],n[it1,0]))
-	while FLAG_CONNECTED == 0:
-		G = np.random.rand(np.sum(n),np.sum(n)) < P
-		G = G.astype(np.float32)
-		G = np.triu(G,1);
-		G = G + G.T
-		L = np.diag(np.sum(G,axis=0))-G
-		D,_ = np.linalg.eig(L)
-		n_connected_components = (np.absolute(D)<=1e-6)
-		if np.sum(n_connected_components.astype(float)) == 1.:
-			FLAG_CONNECTED = 1
-	return G
-	
[…]

### 2019-02-01 control_system.py
Add files via upload

@@ -1,4 +1,5 @@
 import numpy as np
+import pdb
 import scipy.io
 from functools import partial
 
@@ -36,8 +37,13 @@ class ProbabilisticSystem(object):
         samples2 = np.random.normal(0, 1, size=(batch_size, self.control_state_dim))
         return np.hstack((samples,samples2))
 
+    def _exponential_uniform_sample(self, batch_size):
+        samples = np.random.exponential(self.mu, size=(batch_size, self.channel_state_dim))
+        samples2 = np.random.uniform(-10, 10, size=(batch_size, self.control_state_dim))
+        return np.hstack((samples,samples2))
 
-class WirelessSchedulingSystem(ProbabilisticSystem):
+
+class WirelessSchedulingSystem_CC(ProbabilisticSystem):
     def __init__(self, num_users, p=1, Ao=None, Ac=None, W=None, rho = .95, mu=2, S=1, num_rus = 25,num_channels=9):
         super().__init__(num_users*num_rus + num_users*p, num_users, S*num_channels)
 
@@ -62,6 +68,207 @@ class WirelessSchedulingSystem(ProbabilisticSystem):
         self.rate_by_mcs = [1.6,2.4,3.3,4.9,6.5,7.3,8.1,9.8,10.8,12.2,13.5]
         self.rate_by_mcs = np.asarray(self.rate_by_mcs)
         self.bw_by_ru = [1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,4,4,4,4,4,4,10]
+        self.bw_by_ru = np.asarray(self.bw_by_ru)
+
+        self.bw_by_ru = [1,1,1,1,1,1,1,1,1]
+        self.bw_by_ru = np.asarray(self.bw_by_ru)
+
+        T = scipy.io.loadmat('snr_info.mat')
+        self.per_by_snr = T.get('mcs_snr_per_wcl_20B')
+
+        if Ac == None:
+            self.Ac = (0.95-0.35)*np.random.random_sample(num_users)+0.5
+        else:
+            if (Ac.shape != (1, num_users)):
+                raise Exception("Ac is not the correct shape")
+            self.Ac = Ac
+
+        if Ao == None:
+            self.Ao = (1.2-1.01)*np.random.random_sample(num_users)+1.01
+        else:
+            if (Ao.shape != (1, num_users)):
+                raise Exception("Ao is not the correct shape")
+            self.Ao = Ao
+
+        if W == None:
[…]

### 2019-02-01 main.py

@@ -26,20 +26,42 @@ def moving_average(data, window=10):
 
 def run_sys(sys, policies, num_iter, num_reruns=1, batch_size=64):
     history_list = []
+    runtime_list = []
+    num_iters_run = 500
     for i in range(num_reruns):
         print("Simulation " + str(i))
 
         history_dict = {}
+        runtime_dict = {}
+
         for policy_name, _ in policies.items():
             policy_dict = {'lambd': [],
                            'f0': [],
                            'f1': [],
-                           'L': []}
+                           'L': [],
+                           'p': [],
+                           'Ao': [],
+                           'Ac': [],
+                           'x': []}
+            policy_dict2 = {'f0': [],
+                           'f1': [],
+                           'f0_m': [],
+                           'f1_m': [],
+                           'x': [],
+                           'Ao':[],
+                           'Ac':[]}
             history_dict[policy_name] = policy_dict
+            history_dict[policy_name]['Ao'].append(sys.Ao)
+            history_dict[policy_name]['Ac'].append(sys.Ac)
+            runtime_dict[policy_name] = policy_dict2
+            runtime_dict[policy_name]['Ao'].append(sys.Ao)
+            runtime_dict[policy_name]['Ac'].append(sys.Ac)
 
         for k in range(num_iter):
             states = sys.sample(batch_size)
-            print("Iteration " + str(k))
+
+            if k%1000 == 0:
+                print("Iteration " + str(k))
 
             for policy_name, policy in policies.items():
                 actions = policy.get_action(states)
@@ -54,33 +76,55 @@ def run_sys(sys, policies, num_iter, num_reruns=1, batch_size=64):
                 history_dict[policy_name]['f1'].append(np.mean(f1,axis=0))
                 history_dict[policy_name]['L'].append(np.mean(L))
 
+                if k%1000 == 0:
+                    #print(actions)
+                    history_dict[policy_name]['x'].append(states)
+                    history_dict[policy_name]['p'].append(actions)
+
                 policy.learn(states, actions, f0, f1)
+        
+        for k in range(num_iters_run):
+            states = sys.sample(2)
+            print("Run Iteration " + str(k))
[…]

### 2019-02-01 reinforce_policy.py

@@ -3,6 +3,7 @@ import tensorflow as tf
 import math
 import scipy
 import scipy.stats
+import pdb
 import graphtools as gt
 import datatools as dt
 
@@ -134,7 +135,7 @@ class GaussianDistribution(ProbabilityAction):
         log_probs = -0.5 * tf.log(2*np.pi*var) - tf.square(selected_action - mean) / (2. * var)
         log_probs = tf.reduce_sum(log_probs, axis=1)
 
-        return log_probs, output
+        return log_probs, outputex
 
     def get_action(self, params):
         mean = params[:, :action_dim]
@@ -150,8 +151,8 @@ class GaussianDistribution(ProbabilityAction):
 
 class ClassificationDistribution(ProbabilityAction):
     #TODO: FIX THIS
-    def __init__(self, action_dim):
-        super().__init__(25, action_dim)
+    def __init__(self, action_dim, num_classes):
+        super().__init__(num_classes, action_dim)
 
     def log_prob(self, params, selected_action):
         log_probs = 0
@@ -165,13 +166,7 @@ class ClassificationDistribution(ProbabilityAction):
             output = tf.concat([output, ru_class_sig],axis=1)
             #output_list.append(ru_class_sig)
         #print(tf.stack(output_list).shape)
-        #output = tf.concat(ru_class_sig,axis=1)
-        #output = tf.concat([mean, std], axis=1)
-        # TODO: this can create NaNs when self.var is small
-        #log_probs = -0.5 * tf.log(2*np.pi*var) - tf.square(selected_action - mean) / (2. * var)
-        #log_probs = tf.reduce_sum(log_probs, axis=1)
 
-        #return log_probs, tf.stack(output_list)
         return log_probs, output[:,1:]
 
     def get_action(self, params):
@@ -182,23 +177,147 @@ class ClassificationDistribution(ProbabilityAction):
         for i in np.arange(self.action_dim):
             ru_class = np.take(params, np.array(range(i*self.num_param, (i+1)*self.num_param)), axis=1)
             ru_class_sig = 1 / (1 + np.exp(-ru_class))
-            #print(ru_class_sig.shape)
-            #print(ru_class_sig[1,:])
-            #print(np.where(np.random.multinomial(1,ru_class_sig[1,:]/np.sum(ru_class_sig[1,:])))[0][0])
             for j in np.arange(N):
[…]

### 2019-02-01 test_script.py
Delete test_script.py

@@ -1,55 +1,0 @@
-import numpy as np
-import scipy.io
-
-def snr_to_db(tx_power,noise_snr):
-    db = 10*np.log10(tx_power/noise_snr)
-    return db
-
-def db_to_col(snr_value):
-    snr_value = np.minimum(snr_value,35)
-    snr_value = np.maximum(snr_value,-35)
-
-    col = np.floor((snr_value+35)/0.1+1)-1
-    return col.astype(int)
-
-def mcs_to_time(rate_by_mcs,mcs,bw,packet_size):
-	time = 8*packet_size*0.000001/(rate_by_mcs[mcs]*np.tile(bw,(mcs.shape[0],1)))
-	return time
-
-
-if __name__ == '__main__':
-
-	num_users = 6
-	num_channels = 9
-	num_rus = 24
-	T = scipy.io.loadmat('snr_info.mat')
-	per_by_snr = T.get('mcs_snr_per_wcl_20B')
-	rate_by_mcs = np.array([1.6,2.4,3.3,4.9,6.5,7.3,8.1,9.8,10.8,12.2,13.5])
-	bw_by_ru = [1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,4,4,4,4,4,4,10]
-	N = 12
-
-	rho = 0.95
-
-	channel_state_mat = np.random.exponential(1,(N,num_users,num_rus))
-	pp = snr_to_db(2,channel_state_mat)
-	channel_state_col = db_to_col(pp)
-	mcs_per_user_per_ru = np.zeros((N,num_users,num_rus))
-	time_per_user_per_ru = np.zeros((N,num_users,num_rus))
-
-	Ac = 0.6*np.random.random_sample(num_users)
-	Ao = (1.4-1)*np.random.random_sample(num_users)+1
-
-	pp = 1 - per_by_snr[:,channel_state_col]
-
-	#print(pp[:,1,1,:])
-	for i in np.arange(num_users):
-		p_min = (np.power(Ao[i],2)-rho) / (np.power(Ao[i],2) - np.power(Ac[i],2))
-		print(p_min)
-		mcs_per_user_per_ru[:,i,:] = np.maximum(np.sum(pp[:,:,i,:]>=p_min,axis=0),1)-1
[…]

### 2019-02-04 Low_latency_control/response/response_main.tex
Add files via upload

@@
 
-\red{We thank the reviewer for the comments, feedback, and questions.
-The reviewer raised thoughtful questions about the modeling of the problem and some of the assumptions made in the paper.
-The reviewer also suggests to make the numerical results more thorough.
-Each concern is addressed individually below:}
+We thank the reviewer for the comments, feedback, and questions.
+The reviewer raised concerns about the placement of the work within the URLL control and communication framework and literature.
+The reviewer also brings up points of clarification regarding the problem formulation.
+Each concern is addressed individually below.
 
@@
 \issue{ It was unclear whether the focus is on communication for a given control or the other way around. Taht is currently it looks like the solution is based on sequential design.}
-\answer{\red{To be written}}
+\answer{We thank the reviewer for brining up this point of clarification.
+The focus of this work is designing a scheduling protocol for a given system and control.
+As such, the control is assumed to be given, or chosen agnostic to the communication, while the communication is designed relative to both the system dynamics and choice of control.
+\blue{We have added a sentence to Section II to clarfiy this point.} }
 
@@
 \issue{ Linear control was adopted (for tractability) but clearly the nonlinearity is needed}
-\answer{\red{To be written}}
+\answer{The reviewer brings up a reasonable point about the formulation of the wireless control problem.
+Indeed, many practical systems feature both nonlinear dynamics and potentially non-linear controls.
+As stated, the proposed approach focuses on linear systems and controls so that we may derive analytic expressions for the control-aware adaptive packet delivery rates.
+In practice, nonlinear systems may be linearized for the purposes of obtaining such packet delivery rate targets in (23) and performing the proposed CALLS method as written.
+\blue{To demonstrate how this may be done, in the revised manuscript we have included additional simulations in Section VI in which we control a non-linear system using the CALLS method to schedule plant communications.
+In these simulations, while the actual dynamics and control being applied are nonlinear in nature, a linearization of the dynamics is used solely to derive the necessary PDR targets as in expression (23).
+Here, we demonstrate that the proposed method is applicable even in nonlinear systems.}   }
 
@@
 \issue{Authors did not cite recent and fundamental articles within URLLC:}
-\answer{\red{To be written}}
+\answer{We thank the reviewer for pointing out these references.
+They have been added to the revised manuscript, as well as \blue{ included in an added discussion of URLLC in the introduction (Section I).} }
 
@@
 
-\red{We thank the reviewer for the comments, feedback, and questions.
-The reviewer raised thoughtful questions about the modeling of the problem and some of the assumptions made in the paper.
-The reviewer also suggests to make the numerical results more thorough.
-Each concern is addressed individually below:}
+We thank the reviewer for the comments, feedback, and questions.
+The reviewer raised concerns about the readability of the abstract and made some suggestions regarding the simulation results.
+Each concern is addressed individually below:
 
 \issue{ Currently, it is not easy to follow the abstract. It should be revised. The contributions of the this paper should be highlighted.}
-\answer{\red{To be written}}
+\answer{We thank the reviewer for raising this point. \blue{We have revised the abstract in the manuscript to be more clear and the highlight the specific contributions of the paper.}}
 
@@
 
-\red{We thank the reviewer for the comments, feedback, and questions.
-The reviewer raised thoughtful questions about the modeling of the problem and some of the assumptions made in the paper.
-The reviewer also suggests to make the numerical results more thorough.
-Each concern is addressed individually below:}
+We thank the reviewer for the comments, feedback, and questions.
+The reviewer raised points of clarification regarding the system setup and proposed additional comparisons to be made in the simulation results.
+Each concern is addressed individually below:
 
@@
 }
-\answer{\red{To be written}}
+\answer{We thank the reviewer for raising this point, as this aspect of the model is poorly worded in the previous version of the manuscript.
+We do indeed consider a fading channel to model the wireless channel conditions (specifically, IEEE Channel Model E).
+In our simulations, we use AWGN curves to evaluate packet error probabilities, but the curves are evaluated at the complete SNR values, which take into account the instantaneous fast fading channel conditions.
+In addition, we point out that IEEE Channel Model E is the current standard for measuring channel conditions in an indoor environment.
+More sophisticated models may be developed, but are outside the scope of this work.
+\blue{We have revised the manuscript to clarify these points regarding the wireless channel model.}}
 
 \issue{In your system, how do you obtain accurate channel state information? E.g. for AWGN channels, you need to test the noise variance. Is your scheme sensitive to channel estimation errors?}
-\answer{\red{To be written}}
+\answer{This is a good point raised by the reviewer. \red{To be written.}}
 

### 2019-02-05 Low_latency_control/response/response_main.tex
Add files via upload

@@
 As such, the control is assumed to be given, or chosen agnostic to the communication, while the communication is designed relative to both the system dynamics and choice of control.
-\blue{We have added a sentence to Section II to clarfiy this point.} }
+We have added a sentence to Section II to clarify this point.
+}
 
@@
 \issue{ Currently, it is not easy to follow the abstract. It should be revised. The contributions of the this paper should be highlighted.}
-\answer{We thank the reviewer for raising this point. \blue{We have revised the abstract in the manuscript to be more clear and the highlight the specific contributions of the paper.}}
+\answer{We thank the reviewer for raising this point. We have revised the abstract in the manuscript to be more clear and the highlight the specific contributions of the paper.}
 

### 2019-02-05 Low_latency_control/revised/problem_formulation.tex
Add files via upload

--- 2019-02-01 Low_latency_control/arxiv/problem_formulation.tex
@@
 \centering
-\input{tikz_wncs4}
+\input{../images/tikz_wncs4}
 \caption{Wireless control system with $m$ independent systems.
@@
 Thus, application of this control over time will drive the state $\bbx_{i,k} \rightarrow 0$ as $k \rightarrow \infty$.
-As the controller does not always have access to state information, we alternatively consider the estimate of state information of device $i$ known to the controller at time $k$ as
+\red{We assume that this choice of $\bbK$ is given---in other words, the controller is pre-designed with respect to ideal closed loop behavior.} As the controller does not always have access to state information, we alternatively consider the estimate of state information of device $i$ known to the controller at time $k$ as
 %
@@
 %\input{multiplex.tex}
-\includegraphics[width=0.5\textwidth]{schedule2.png}
+\includegraphics[width=0.5\textwidth]{../images/schedule2.png}
 \caption{Multiplexing of frequencies (RU) and time (PPDU) in IEEE 802.11ax transmission window (formally referred as Transmission Opportunity or TXOP in the standard.
@@
 \begin{table}
-\input{mcs.tex}
+\input{../images/mcs.tex}
 \caption{Data rates for MCS configurations in IEEE 802.11ax for 20MHz channel.

### 2019-02-05 Low_latency_control/revised/root.tex
Add files via upload

--- 2019-02-01 Low_latency_control/arxiv/root.tex
@@
 
-\input{mysymbol.sty}
-\input{my_sections}
-%\input{TikZGallery-master/Sty/TikZGallery-Preamble}
+\input{../mysymbol.sty}
+\input{../my_sections}
+\input{../TikZGallery-master/Sty/TikZGallery-Preamble}
 %\renewcommand{\baselinestretch}{.995}
@@
 
-%\renewcommand \blue[1]{}  \renewcommand \red[1]{}
+%\renewcommand \blue[1]{}  \renewcommand \red[1]{
 
@@
 \author{Mark Eisen$^*$ \quad Mohammad M. Rashid$^\dagger$ \quad Konstantinos Gatsis$^*$ \\ \textup{Dave Cavalcanti$^\dagger$\quad Nageen Himayat$^{\dagger}$ \quad Alejandro Ribeiro$^*$}
-\thanks{Supported by the Intel Science and Technology Center for Wireless Autonomous Systems and ARL DCIST CRA W911NF-17-2-0181.
+\thanks{Supported by the Intel Science and Technology Center for Wireless Autonomous Systems.
 The authors are with the ($^*$)Department of Electrical and Systems Engineering, University of Pennsylvania and ($^\dagger$)Wireless Communications Research, Intel Corporation.
@@
 \begin{abstract}
-%We consider the problem of allocating radio resources, such as frequency, bandwidth, data rates and time slots, over wireless communication links to control a series of independent wireless control systems.
-Low-latency transmissions are necessary in control systems that contain high sampling rates, thus necessitating fast transmissions.
-Compared to wired links, wireless links are highly susceptible to communication errors arising from channel noise and interference.
-Achieving fast data rates over wireless links thus comes at the cost of reliability in the form of high packet error rates.
-However, the effect of the communication link errors on the control system performance is not static but depends dynamically on the control system state.
-Traditional resource allocation methods over wireless links are unaware of control system states and thus are designed with static link error performance, which could be very inefficient if  error performance assumption is too conservative (e.g.
-too low error rate target), or unacceptably unreliable if the assumption is too relaxed.
-In this paper, we propose a novel control-communication co-design approach to this low-latency resource allocation problem.
-We develop methods that incorporate control state information to make scheduling decisions over time on frequency, bandwidth and data rates across the next-generation Wi-Fi based wireless communication links connecting the different components of the control systems.
-Control systems that are closer to instability or further from a desired range in a given control cycle are given priority in the scheduling.
-Rather than a simple priority ranking, we derive precise packet error rate targets for each system needed to satisfy respective Lyapunov control targets and make scheduling decisions to best meet targets while minimizing total transmission time.
-The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.
-We consider the problem of allocating radio resources over wireless communication links to control a series of independent wireless control systems.
-Low-latency transmissions are necessary in enabling time-sensitive control systems to operate over wireless links with high reliability.
-Achieving fast data rates over wireless links thus comes at the cost of reliability in the form of high packet error rates compared to wired links due to channel noise and interference.
-However, the effect of the communication link errors on the control system performance depends dynamically on the control system state.
-We propose a novel control-communication co-design approach to the low-latency resource allocation problem.
-We incorporate control and channel state information to make scheduling decisions over time on frequency, bandwidth and data rates across the next-generation Wi-Fi based wireless communication links that close the control loops.
+\red{We consider the problem of allocating radio resources over wireless communication links to control a series of independent wireless control systems.
+Low-latency transmissions are necessary in enabling time-sensitive control systems with high sampling rates to operate over wireless links.
+Enabling low-latency through fast data rates comes at the cost of reliability in the form of higher packet error rates due to channel noise and interference.
+However, the cost of such communication link errors on the control system performance depends dynamically on the control system state.
+We propose a novel control-aware communication design to the low-latency resource allocation problem.
+In our proposed method, we incorporate both control and channel state information to scheduling transmissions across time slots, frequency bands, and data rates using the next-generation Wi-Fi scheduling architecture.
 Control systems that are closer to instability or further from a desired range in a given control cycle are given higher packet delivery rate targets to meet.
-Rather than a simple priority ranking, we derive precise packet error rate targets for each system needed to satisfy stability targets and make scheduling decisions to meet such targets while reducing total transmission time.
-The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.
-
+Rather than a simple priority ranking, we derive precise adaptive packet error rate targets for each system needed to satisfy control-specific performance requirements.
+We use these adaptive rate targets to make scheduling decisions that reduce total transmission time.
+The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.}
 \end{abstract}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{wireless_ll_control,scheduling_control}
+\bibliography{../wireless_ll_control,../scheduling_control}
 

### 2019-02-05 Low_latency_control/response/response_main.tex
Add files via upload

@@
 \issue{Reliability as not addressed however it is intertwined with latency}
-\answer{\red{To be written}}
+\answer{This is a good point raised by the reviewer that deserves clarification.
+Reliability is a key component of a practical system that must be carefully managed in low latency systems.
+In our work, rather than aim for traditional, communication-based reliability targets, such as $0.999$ packet delivery rates, we consider the fact that the observed performance of a control system is more closely related to a \emph{control}-based reliability metric, such as stability or distance from origin.
+\blue{As discussed in Section...}, by defining reliability targets with respect to the Lyapunov cost function rather than with respect to a fixed packet delivery rate target, we are able to derive \emph{adaptive} packet delivery rate requirements that change with the current performance of each control system.
+When a control system is in a safe or stable range, it can remain in good performance with smaller packet delivery rates than a control system in a less desirable region.
+Thus, the proposed CALLS method aims to achieve reliability of control system performance, in other words an \emph{adaptive reliability}, in contrast to a standard high reliability framework.
+Such adaptive rates are critical in enabling low-latency transmissions in large-scale wireless settings, as observed our simulations in Section \blue{???}.
+We point out that the proposed control-adaptive framework is thus presented as an alternative to high reliability scheduling schemes in settings in which the control systems are well-managed and defined, such as in the case of industrial controls.
+\blue{We have added a discussion of this to the revised manuscript.}}
 

### 2019-02-12 Low_latency_control/revised/conclusions.tex
Add files via upload

@@
 
-The results presented in this paper suggest an interesting
+\red{The results presented in this paper suggest an interesting
 potential for control-aware resource allocation and scheduling, particularly in low-latency industrial systems. By considering the control-specific targets such as maintaining
 stability or an error margin, we observe that the standard
-high reliability targets considered (e.g. packet delivery rates $\geq 0.999$) can in some cases be substantially stricter than
+high reliability targets considered in URLCC (e.g. packet delivery rates $\geq 0.999$) can in some cases be substantially stricter than
 necessary for adequate performance. Wireless control systems
@@
 lower packet delivery rates, which in turn make low-latency
-communications more achievable.
-Furthermore, in realistic industrial systems there will be many heterogeneous devices being controlled, whose variation in communication needs is well-served by control-aware opportunism proposed in this paper.
+communications more achievable.} Furthermore, in realistic industrial systems there will be many heterogeneous devices being controlled, whose variation in communication needs is well-served by control-aware adaptivity proposed in this paper.
 This suggests the potential for wireless communications to be adopted using a smart control-communication co-design approach even

### 2019-02-12 Low_latency_control/revised/introduction.tex
Add files via upload

@@
-The reliability and robustness of wireless communication systems has become a key component in the design and implementation of large scale systems in the Internet-of-Things (IoT).
+%The reliability and robustness of wireless communication systems has become a key component in the design and implementation of large scale systems in the Internet-of-Things (IoT).
 For these systems to operate successfully, it is important that the data collected by sensors can be communicated throughout the IoT network.
@@
 
-In the wireless communications research and industry, many radio resource allocation schemes in the form of wireless scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}.
+%In the wireless communications research and industry, many radio resource allocation schemes in the form of wireless scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}.
 For instance, round-robin scheduling is a common approach due to its simplicity inherent fairness.
@@
 However, all such methods are unable to leverage the information on the control system dynamics and thus can make unsuitable radio resource allocation decisions when deployed in time-sensitive wireless control systems.
+
+\red{The reliability and robustness of wireless communication systems has become a key component in the design and implementation of large scale systems in the Internet-of-Things (IoT).
+For these systems to operate successfully, it is important that the data collected by sensors can be communicated throughout the IoT network.
+While the most reliable form of data communication is through wired connections, the scale and mobility of modern IoT settings has rendered the cost of installing and maintaining a wired network a significant challenge~\cite{zand2012wireless,wollschlaeger2017future}.
+Consequently, there is great interest in the design of wireless control systems that can achieve reliable performance~\cite{varghese2014wireless, li2017review}.
+Low-latency wireless transmissions are however a necessary feature in many wireless IoT systems, particularly those in industrial control~\cite{zand2012wireless,varghese2014wireless,weiner2014design,popovski2018wireless, bennis2018ultra}.
+The primary challenge in so-called ultra reliable low latency communications (URLLC) is then the inherent trade-off that occurs between reliability and latency, as it is difficult to maintain high reliability while using higher data rates due to the stochasticity of the wireless channel.
+It is necessary to design resource allocation and scheduling protocols for wireless control systems that can both meet reliability \emph{and} latency requirements of the control system.
+
+Many radio resource allocation schemes in the form of wireless scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{yaacoub2012survey}.
+Algorithms such as proportional fair (PF)~\cite{kim2005proportional}, max-SNR~\cite{knopp1995information} and their variations have been designed for maximizing the overall system throughput and/or fairness.
+For URLLC or time-sensitive applications such as industrial control and automation, the experiences of individual users/devices are highly dependent on the ability of the network to deliver their packets within a latency bound.
+Generic delay-aware schedulers such as EDF~\cite{wu2014analysis} and WFQ~\cite{lu1999fair} do not adapt to wireless channel conditions, which are considered in M-LWDF \cite{andrews2001providing}.
+Within the URLLC framework, low-latency communications have been enabled through achieving various forms of diversity \cite{swamy2015cooperative,popovski2018wireless,nielsen2018ultra}.
+However, all such methods are unable to leverage the information on the control system dynamics and thus can make unsuitable radio resource allocation decisions when deployed in time-sensitive wireless control systems.}
 
@@
 Initial approaches abstract control performance requirements in the time/frequency domain, e.g., how often a task needs resource access, employing algorithms from real-time scheduling theory~\cite{Branicky_RM, Liu_Real_time_systems}.
-More recent scheduling approaches often depend on the current control system states, i.e., informally the subsystem with the largest state discrepancy is scheduled to communicate -- see, e.g.,~\cite{Donkers_switched, Walsh_stability, Hristu_Kumar_interrupt_based, Egerstedt_queue, Hirche_Scheduling_Price, Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
+More recent scheduling approaches often depend on the current control system states, i.e., informally the subsystem with the largest state discrepancy is scheduled to communicate -- see, e.g.,~\cite{Donkers_switched,  Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
 %
 Alternatively scheduling can take into account current wireless channel conditions opportunistically to meet target control system reliability requirements~\cite{GatsisEtal15}.
-None of these approaches, however, are explicitly designed to achieve strong performance it low-latency scenarios.
+\red{None of these approaches, however, are explicitly designed to achieve strong performance it low-latency scenarios, which are critical in modern industrial control and factory settings \cite{brahmi2015deployment,yilmaz2015analysis}.}
 
@@
 Furthermore, because these control-based success rate requirements may be significantly lower than traditional, high reliability communication demands, the proposed method is better suited to find scheduling configurations that can moreover meet the strict latency requirements imposed by the physical system.
-This is in contrast to the control-aware approach taken in \cite{GatsisEtal15}, which focuses on low-power and infrequent transmissions rather than the low-latency setting of interest in industrial control.
+\red{In contrast to more standard URLLC approaches of fixing high reliability targets, our approach can be seen as \emph{adaptive} reliability with respect to control state and dynamics.
+}
 
@@
 The performance of the CALLS method is analyzed in a series of simulation experiments in which its performance is compared against a control-agnostic procedure (Section \ref{sec_simulation}).
-We demonstrate in numerous low-latency control systems that the control-aware method can support more users than the alternative and achieve more robust overall performance.
-
-%The reliability and robustness of wireless communication systems has become a key component in the design and implementation of large scale systems in the Internet-of-Things (IoT).
-For these systems to operate successfully, it is important that the data or information collected by sensors can be communicated throughout the IoT network.
-While the most reliable form of data communication is through wired connections, the scale and mobility of modern IoT settings has rendered the cost of installing and maintaining a wired network a significant challenge~\cite{wollschlaeger2017future}.
-Consequently, the design of wireless control systems that can achieve reliable performance has become of great interest~\cite{varghese2014wireless, li2017review}.
-The primary challenge is then the inherent trade off that occurs between reliability and latency.
-Low-latency transmissions are a necessary feature in many wireless IoT systems, particularly those in industrial control~\cite{varghese2014wireless}.
-It is difficult to maintain high reliability while using higher data rates, however, due to the stochasticity of the wireless channel.
-It is necessary then to design resource allocation and scheduling protocols for wireless control systems that can both meet reliability \emph{and} latency requirements of the control system.
-%
-%
-%In the wireless communications research and industry, many scheduling techniques have been proposed to provide reliability, or quality of service (QoS), to users across the network in the form of throughput, fairness and/or latency~\cite{cao2001scheduling,wongthavarawat2003packet,fattah2002overview,yaacoub2012survey}.
-For instance, round-robin scheduling is a common approach due to its simplicity inherent fairness.
-Algorithms such as proportional fair (PF)~\cite{kim2005proportional}, max-SNR~\cite{knopp1995information} and their variations have been designed for maximizing the overall system throughput and/or fairness.
-However, these algorithms can perform very poorly with regards to ensuring that all of the admitted users' packet delivery rates and latency targets will be met.
-For time-sensitive applications such as industrial control and automation, the experiences of individual users/devices are highly dependent on the ability of the network to deliver their packets within a latency bound.
-If this objective is not met, the high overall system throughput or throughput-fairness means little to the end users.
-On the other hand, the generic delay-aware schedulers such as EDF~\cite{wu2014analysis} and WFQ~\cite{lu1999fair} do not properly adapt to wireless channel conditions and may make poor radio resource allocation decisions.
-These algorithms have also been modified to adapt to wireless channel conditions, such as in M-LWDF \cite{andrews2001providing}.
-However, all such wireless scheduling algorithms are unable to leverage the information on the control system dynamics and thus, perform inefficiently when deployed in time-sensitive wireless control systems.
-%
-%In the context of wireless control systems, there was been a range or works that incorporates control system information in the networking and communication policies.
-%%
-%The mechanisms usually examined are either static or dynamic.
-Typical examples of the first type are periodically protocols where the wireless devices transmit in a predefined repeating order, e.g., round-robin.
-Control system stability under such protocols can be analyzed -- see, e.g.,~\cite{Hespanha_survey, Schenato_foundations, Donkers_switched, Branicky_stability}.
-Periodic sequences leading to stability~\cite{Hristu_shared_feedback}, controllability and observability~\cite{Hristu_communication_control}, or optimizing control objectives~\cite{LeNy_resource_LQR,Meier_measurement_control, Scheduling_control_combinatorics} have been proposed.
-%%
-%The second type of schedulers, the dynamic ones, do not rely on a predefined sequence but decide access to the communication medium dynamically at each step.
-%, for example by dynamically assigning priorities to the competing tasks.
-%Initial approaches abstract control performance requirements in the time/frequency domain, e.g., how often a task needs resource access, employing algorithms from real-time scheduling theory~\cite{Branicky_RM, Liu_Real_time_systems}.
-More recently scheduling approaches often depend on the current control system states, i.e., informally the subsystem with the largest state discrepancy is scheduled to communicate -- see, e.g.,~\cite{Donkers_switched, Walsh_stability, Hristu_Kumar_interrupt_based, Egerstedt_queue, Hirche_Scheduling_Price, Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
-%%
-%Alternatively scheduling can take into account current wireless channel conditions opportunistically to meet target control system reliability requirements~\cite{GatsisEtal15}.
-None of these approaches, however, are explicitly designed to minimize communications latency or achieve strong performance with specific low-latency requirements.
-%
-%
-%
-%In this paper, we develop a control-aware, high reliability, low-latency IEEE 802.11ax WiFi protocol~\cite{bellalta2016ieee} that is designed to reduce the total transmission time at each uplink cycle in the control loop.
-This is done through the mathematical formulation of the control system design goal in the form of a Lyapunov function that ensures stability of the control system.
-This formulation naturally induces a bound on the packet delivery rate each control system needs to achieve to meet the control-based goal.
-Such packet delivery rates depend upon current control and channel states and thus dynamically change over the course of the system life time.
-This can be viewed as an opportunistic protocol with respect to both the current control states \emph{and} channel conditions.
-Furthermore, because these control-based success rate requirements may be significantly lower than traditional, high reliability communication demands, the proposed method is better suited to find scheduling configurations that can moreover meet the strict latency requirements imposed by the physical system.
-%
-%The paper is organized as follows.
-We formulate the wireless control system in which state information is communicated to the control over a wireless channel.
-Due to the potential for random packet drops, this is modeled as a switched dynamical system (Section \ref{sec_problem_formulation}).
-A Lyapunov function is used to evaluate the stability of the control state, and the uncertainty in this measurement grows the more consecutive packets are lost for a particular system.
-We then discuss the scheduling parameters of the IEEE 802.11ax communication model (Section \ref{sec_comm_model}).
-%%, which allows for centralized scheduling of transmissions through the use of varying bandwidth resource units, time-slotted PPDUs, and date rate selection via MCS selection (Section \ref{sec_comm_model}).
-%From there, we derive a mathematical formulation of the optimal scheduling problem (Section \ref{sec_optimal}).
-This can be formulation by minimizing a control cost with an explicitly latency constraint (Section \ref{sec_optimal_a}) or minimizing transmission time with an explicit control performance constraint (Section \ref{sec_optimal_b}).
-%
-%Using the control-constrained optimal scheduling formulation, we develop the control-aware low latency scheduling (CALLS) method (Section \ref{sec_calls}).
-The CALLS method uses current control states and channel conditions to derive dynamic packet success rates for each user (Section \ref{sec_psr}).
-In this way, control systems that are closest to instability will be given priority in the scheduling so that they may close their control loops.
-The scheduling procedure consists of a random user selection procedure to reduce the number of required PPDUs that incur significant overhead  (Section \ref{sec_rss}), followed by an assignment-method based scheduling of selected users to minimize total transmission time (Section \ref{sec_assignment}).
-The performance of the CALLS method is analyzed in a series of simulation experiments in which its performance is compared against a control-agnostic procedure (Section \ref{sec_simulation}).
-We demonstrate in numerous low-latency control systems that the control-aware method can support more users than the alternative and achieve more robust overall performance.
+\red{We demonstrate in numerous control systems that the control-aware, adaptive reliability approach can support more users than the alternative and achieve more robust overall performance.}

### 2019-02-12 Low_latency_control/response/response_main.tex
Add files via upload

@@
 \issue{ The authors need to recast the framework within ultra-reliable and low-latency control and communication.}
-\answer{\red{To be written}}
+\answer{We thank the reviewer for this important comment and suggestion.
+We have revised the introduction of the manuscript substantially to contextualize the work being done here within the framework of URLLC.
+In doing so, we have added the following important references on the topic to the disucssion in the introduction.
+\begin{itemize}
+\item M. Weiner, M. Jorgovanovic, A. Sahai, and B. Nikolie, “Design of a
+low-latency, high-reliability wireless communication system for control
+applications,” in Communications (ICC), 2014 IEEE International
+Conference on. IEEE, 2014, pp. 3829–3835.
+\item P. Popovski, J. J. Nielsen, C. Stefanovic, E. de Carvalho, E. Strom,
+K. F. Trillingsgaard, A.-S. Bana, D. M. Kim, R. Kotaba, J. Park et al.,
+“Wireless access for ultra-reliable low-latency communication: Principles
+and building blocks,” IEEE Network, vol. 32, no. 2, pp. 16–23, 2018.
+\item M. Bennis, M. Debbah, and H. V. Poor, “Ultra-reliable and lowlatency
+wireless communication: Tail, risk and scale,” arXiv preprint
+arXiv:1801.01270, 2018.
+\item V. N. Swamy, S. Suri, P. Rigge, M. Weiner, G. Ranade, A. Sahai, and
+B. Nikolic, “Cooperative communication for high-reliability low-latency
+wireless control,” in Communications (ICC), 2015 IEEE International
+Conference on. IEEE, 2015, pp. 4380–4386.
+\item J. J. Nielsen, R. Liu, and P. Popovski, “Ultra-reliable low latency communication
+using interface diversity,” IEEE Transactions on Communications,
+vol. 66, no. 3, pp. 1322–1334, 2018.
+\end{itemize}
+In addition to this more comprehensive review of URLLC techniques, we have the introduction to highlight the fact that our control-aware scheduling approach can be seen as a means of defining an \emph{adaptive}-reliability framework for achieving low-latency communications, in contrast to fixed ultra reliability targets.
+This adaptive reliability framework considers the particular control system dynamics to determine the reliability necessary to keep all plants in good conditions.
+Our numerical simulations show how this alternative framework can improve the scale of low-latency wireless control systems by loosening reliability requirements where applicable.
+}
 
@@
 In our work, rather than aim for traditional, communication-based reliability targets, such as $0.999$ packet delivery rates, we consider the fact that the observed performance of a control system is more closely related to a \emph{control}-based reliability metric, such as stability or distance from origin.
-\blue{As discussed in Section...}, by defining reliability targets with respect to the Lyapunov cost function rather than with respect to a fixed packet delivery rate target, we are able to derive \emph{adaptive} packet delivery rate requirements that change with the current performance of each control system.
+As discussed in Section III and IV, by defining reliability targets with respect to the Lyapunov cost function rather than with respect to a fixed packet delivery rate target, we are able to derive \emph{adaptive} packet delivery rate requirements that change with the current performance of each control system.
 When a control system is in a safe or stable range, it can remain in good performance with smaller packet delivery rates than a control system in a less desirable region.
 Thus, the proposed CALLS method aims to achieve reliability of control system performance, in other words an \emph{adaptive reliability}, in contrast to a standard high reliability framework.
-Such adaptive rates are critical in enabling low-latency transmissions in large-scale wireless settings, as observed our simulations in Section \blue{???}.
+Such adaptive rates are critical in enabling low-latency transmissions in large-scale wireless settings, as observed our simulations in Section V.
 We point out that the proposed control-adaptive framework is thus presented as an alternative to high reliability scheduling schemes in settings in which the control systems are well-managed and defined, such as in the case of industrial controls.
-\blue{We have added a discussion of this to the revised manuscript.}}
+We have added a discussion of this to the revised manuscript.}
 
@@
 \answer{We thank the reviewer for pointing out these references.
-They have been added to the revised manuscript, as well as \blue{ included in an added discussion of URLLC in the introduction (Section I).} }
+They have been added to the revised manuscript, as well as  included in an added discussion of URLLC in the introduction (Section I).
+The references added are listed in a previous response to this reviewer.
+}
 

### 2019-02-12 Low_latency_control/revised/simulations.tex
Add files via upload

--- 2019-02-01 Low_latency_control/arxiv/simulations.tex
@@
 The parameters for the simulation setup are provided in Table \ref{tab_simulation}.
-The packet delivery rate function $q(\bbh,\mu,\bbsigma)$ is computed using the standard AWGN noise curves for wireless channels.
-The transmission time $\tau(\mu,\bbsigma)$ is computed in the simulations using the associated data rates of an MCS in Table \ref{tab_mcs} for a 100 byte packet and overhead (e.g.
+\red{The wireless fading channel modeled using  IEEE Indoor Channel Model E, while the packet delivery rate function $q(\bbh,\mu,\bbsigma)$ is computed using the corresponding SNR values via an AWGN noise curve.} The transmission time $\tau(\mu,\bbsigma)$ is computed in the simulations using the associated data rates of an MCS in Table \ref{tab_mcs} for a 100 byte packet and overhead (e.g.
 TFs) of the 802.11ax specifications.
@@
 \begin{figure}
-%\input{pendulum.tex}
-\centering
-\includegraphics[height=.2\textheight]{pendulum.PNG}
+\input{../images/pendulum.tex}
 \caption{Inverted pendulum-cart system $i$.
@@
 \centering
-\includegraphics[width=.45\textwidth]{ip_dist2.pdf}
+\includegraphics[width=.45\textwidth]{../images/ip_dist.eps}
 \caption{Average pendulum distance to center vertical for $m=25$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max}= 1$ ms latency threshold.
@@
 \centering
-\includegraphics[width=.45\textwidth]{bbb_dist.pdf}
+\includegraphics[width=.45\textwidth]{../images/bbb_dist.eps}
 \caption{Average ball distance to center for $m=50$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold.
@@
 \centering
-\includegraphics[width=.45\textwidth]{bbb_psr.pdf}
+\includegraphics[width=.45\textwidth]{../images/bbb_psr.eps}
 \caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold.

### 2019-02-18 Low_latency_control/revised/introduction.tex
Add files via upload

@@
 Alternatively scheduling can take into account current wireless channel conditions opportunistically to meet target control system reliability requirements~\cite{GatsisEtal15}.
-\red{None of these approaches, however, are explicitly designed to achieve strong performance it low-latency scenarios, which are critical in modern industrial control and factory settings \cite{brahmi2015deployment,yilmaz2015analysis}.}
+\red{None of these approaches, however, are explicitly designed to achieve strong performance in large scale low-latency systems, which are critical in modern industrial control and factory settings \cite{brahmi2015deployment,yilmaz2015analysis}.}
 

### 2019-02-18 Low_latency_control/revised/scheduling_alg.tex
Add files via upload

@@
  \end{remark}
+
+\red{
+ \begin{remark}\label{remark_estimation}\normalfont
+In practical systems, the channel state information $\bbH_k$ is often obtained with some estimation errors, which may impact the channel aware scheduling approach taken here.
+Observer, however, that the computation of adaptive PDR targets in \eqref{eq_pdr_constraint} does not depend upon channel state information.
+Thus, the primary component of the CALLS method---namely Steps 3-6 in Algorithm \ref{alg_calls}--- is unaffected by inaccurate channel estimation.
+Likewise, the assignment method-based scheduling in Step 8 does also not directly depend upon channel information---see the formulation of the assignment problem in \eqref{eq_assignment}.
+The only component that may be negatively impacted by estimation errors is Step 7, i.e.
+the maximum MCS selection in \eqref{eq_mcs_select}.
+Here, estimation errors may result in the selection of an MCS that cannot meet the target PDR targets.
+Such an effect can be mitigated by selecting a more conservative MCS selection to account for channel estimation errors---e.g.
+transmitting at an MCS $\mu_{i,k}(\bbvarsigma) - 1$ rather than $\mu_{i,k}(\bbvarsigma)$.
+Such a conservative scheme would tradeoff latency to the benefit of reliability or robustness.
+ \end{remark}
+ }

### 2019-02-18 Low_latency_control/revised/simulations.tex
Add files via upload

@@
 \begin{bmatrix}
-1 & 0 & 0 & 0 \\
+1 & 0.037 & 3.477 & 0.042 \\
 0 & 2.055 & -0.722 & 4.828 \\
@@
     enlargelimits=0.15,
-    legend style={at={(0.5,0.85)},
+    legend style={at={(0.5,0.85),size=\footnotesize},
       anchor=north,legend columns=-1},
@@
     ]
-\addplot coordinates {(0.5 ms,9) (1 ms,19) (1.5 ms,36)};
-\addplot coordinates {(0.5 ms,12) (1 ms,28) (1.5 ms,55)};
+\addplot[blue!40!black,fill=blue!20!white] coordinates {(0.5 ms,9) (1 ms,19) (1.5 ms,36)};
+%\addplot[green!40!black,fill=green!20!white] coordinates {(0.5 ms,9) (1 ms,19) (1.5 ms,36)};
+\addplot[red!40!black,fill=red!20!white] coordinates {(0.5 ms,12) (1 ms,28) (1.5 ms,55)};
 \legend{Fixed PDR,  CALLS}
@@
 
+\red{
+In the simulations performed with the balancing board system, in addition to making comparisons of the CALLS method to a fixed PDR low latency scheduling scheme, we perform additional comparisons to a standard control-aware scheduling approach---namely, the event-triggered scheduling approach \cite{Cervin_event_scheduling, mamduhi2014event}.
+In event triggered scheduling, we schedule devices only when its estimated control state goes above some threshold value.
+When such an event occurs, this device is scheduled with a fixed high reliability PDR using a low-latency assignment based scheduling method.
+This, in effect, combines the selective scheduling approach of CALLS with fixed high reliability PDR targets commonly used in URLLC.
+
 In Figure \ref{fig_bbb_dist} we show the results of a representative simulation of the control of $m=50$ balancing board ball systems with a latency bound of $\tau_{\max} = 10^{-3}$ seconds.
-Observe that, in this system, even with a large number of users the CALLS method can keep all systems very close to the center of the board, while the fixed PDR scheduler loses a few of the balls due to the agnosticism of the scheduler.
+Observe that, in this system, even with a large number of users, both the event-triggered scheduling and the CALLS method can keep all systems very close to the center of the board, while the fixed PDR scheduler loses a few of the balls due to the agnosticism of the scheduler.
 
  To dive deeper into the benefits provided by control aware scheduling, we present in Figure \ref{fig_bbb_psr} a histogram of the actual  packet delivery rates each of the devices achieved over the representative simulation.
-It is interesting to observe that, in the CALLS method, the achieved PDRs are closely concentrated, ranging from 0.3 to 0.44.
+It is interesting to observe that, in both the event triggered and CALLS method, the achieved PDRs are closely concentrated, ranging from 0.3 to 0.44.
 On the other hand, using a fixed PDR scheduling scheme, the non-variable rates are too strict for the low-latency system to support, and without control-aware scheduling the achieved PDRs range wildly from close to 0 to close to 1.
@@
 This is comparison to the inverted pendulum system, in which the pendulums were kept stable with PDRs in the range 0.6-0.75.
+ }
 
@@
 \includegraphics[width=.45\textwidth]{../images/bbb_dist.eps}
-\caption{Average ball distance to center for $m=50$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold.
+\caption{Average ball distance to center for $m=50$ devices using (top) CALLS, (middle) event-triggered, and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold.
 The proposed control aware scheme keeps all balancing balls close to center, while fixed-PDR scheduling cannot.}
@@
 \includegraphics[width=.45\textwidth]{../images/bbb_psr.eps}
-\caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold.
+\caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) fixed PDR, (middle) event-triggered, and (bottom) CALLS scheduling with $\tau_{\max} =1$ ms latency threshold.
 The proposed control aware scheme achieves similar PDRs for all devices, while the control agnostic scheduling results in large variation in packet delivery.}
@@
 
+\red{
 We present in Figure \ref{fig_bbb_bar} the final capacity results obtained over all the simulations for the balancing board ball system.
@@
 The even greater improvement here relative to the inverted pendulum simulations can be attributed to the slower dynamics of the balancing board ball, which allows for even more gains using control-aware PDRs due to the lower PDR requirements of the system.
+Likewise, the Lyapunov-based adaptive PDR requirements allow for even greater scalability than the more standard event-triggered approach, which can service only 17 and 50 users with 0.5 and 1 ms latency thresholds, respectively.}
 
@@
     enlargelimits=0.15,
-    legend style={at={(0.5,0.85)},
+    legend style={at={(0.4,.85),font=\footnotesize},
       anchor=north,legend columns=-1},
@@
     ]
-\addplot coordinates {(0.5 ms,9) (1 ms,27)};
-\addplot coordinates {(0.5 ms,25) (1 ms,60)};
-\legend{Fixed PDR, CALLS}
+\addplot[blue!40!black,fill=blue!20!white] coordinates {(0.5 ms,9) (1 ms,27)};
+\addplot[green!40!black,fill=green!20!white] coordinates {(0.5 ms,17) (1 ms,50)};
+\addplot[red!40!black,fill=red!20!white] coordinates {(0.5 ms,25) (1 ms,60)};
+\legend{Fixed PDR, Event-Trigged, CALLS}
 \end{axis}
 \end{tikzpicture}
-\caption{Total number of balancing ball board devices that can be controlled using Fixed-PDR and CALLS scheduling for various latency thresholds.}
+\caption{Total number of balancing ball board devices that can be controlled using Fixed-PDR, Event-Triggered, and CALLS scheduling for various latency thresholds.}
 \label{fig_bbb_bar}

### 2019-02-18 Low_latency_control/response/response_main.tex
Add files via upload

@@
 
-\red{
 We thank the associate editor and the reviewers for the time and effort invested in providing constructive suggestions.
 The requests most common among the reviewers concerned numerous points of clarification that add to the understandability and readability of the manuscript.
-Two reviewers raised questions and comments regarding the derivation of the problem in Example 1.
-Reviewers also raised key questions regarding some assumptions made in the analysis.
+Two reviewers requested additional comparisons in the numerical simulations.
+Reviewers also raised key questions regarding the placement of the proposed work in the context of ultra reliability low latency communications and control (URLLC).
 A more detailed account of the major changes implemented in the manuscript follows.
+These changes are additionally highlighted in the manuscript in \red{red} font.
 
@@
 
-\item[{\bf[C1]}] Reviewers 1 and 3 expressed concerns about the derivation of Example 1 and the stability analysis in Section V-C.
-We have thus expanded upon Example 1 that makes the derivations more explicit.
-We have revised Section V-C to give further details on the stability claim.
-We have also identified an omission in the previous version that has been corrected.
-More specifically, assuming we have a statistically accurate dual variable solution, using Proposition 2 and Theorem 2 we argue that the actual packet success rate selected by the algorithm can be made sufficiently close to the optimal one.
-Since the latter leads to system stability then the packet success rate selected by our algorithm also leads to stability.
-We thank the reviewers for these comments.
+\item[{\bf[C1]}] Reviewers 1 and 2 expressed concerns about the completeness the numerical simulations and felt that it is necessary to include a comparison against another control-aware scheduling method that is commonly used in wireless control systems.
+We believe this is a good point to be raised, and have made the following changes.
+Although there is not scheduling method to our knowledge that explicitly addresses the challenge of latency using control state information, we have adapted the broad ideas of event-triggered control to the low-latency scheduling framework developed here.
+We have added to the revised manuscript a comparison of the performance of the proposed CALLS method against an event-triggered low-latency scheduling alternative and show that the proposed method is able to outperform the standard event-triggered approach.
 
-\item[{\bf[C2]}] Reviewer 1 has suggested some additional numerical experiments to explore the effects of different parameters in the performance.
-Per the reviewer's suggestion, we have added Figure 4 to the paper, which compares both the suboptimality and constraint violation of the resource allocations generated by Newton's method over a non-stationary channel using two different accuracies $\hat{V}$.
-The results demonstrate an interesting effect that can occur if the choice of $\hat{V}$ is too strict.
-We thank the reviewer for this suggestion.
+\item[{\bf[C2]}] Reviewer 1 has suggested some additional clarification and context of the work within the framework of URLLC.
+This was an important point to bring up and we thank the reviewer for this suggestion.
+To address this concern, we have revised the introduction of the manuscript to add important references on the topic of URLLC, as well as clarify the alternative approach taken in this work: namely, by using control system knowledge to derive \emph{adaptive} reliability rate targets, rather that fixed high or ultra reliability targets commonly used.
+These adaptive targets are shown to still allow for successful control of these systems and further enable low-latency by loosening the constraints on reliability.
 
-\item[{\bf[C3]}] Reviewers raised numerous questions regarding some of the practical considerations in Section VI, such as backtracking parameter selection as well as satisfying the non-stationary assumption in Assumption 4.
-To address these concerns, we have added additional details to Section VI to clarify many of these points, and an additional remark after Assumption 4.
-We thank the reviewers for raising these questions.
+\item[{\bf[C3]}] Reviewers raised numerous points of clarification regarding the various system models, such as the wireless control model and the fading channel model.
+We have individually addressed each concern by adding clarifying sentences to the revised manuscript.
 
-\end{itemize}}
+\end{itemize}
 
@@
 We have revised the introduction of the manuscript substantially to contextualize the work being done here within the framework of URLLC.
-In doing so, we have added the following important references on the topic to the disucssion in the introduction.
+In doing so, we have added the following important references on the topic to the discussion in the introduction.
 \begin{itemize}
@@
 \issue{ The article would improve a lot if specific use cases are defined.}
-\answer{\red{To be written}}
+\answer{We thank the reviewer for this suggestion.
+We have added a sentence to the introduction clarifying the use cases of particular interest in this work, namely the industrial control and factory settings, where there is significant interest in closing many low-latency control over wireless links.}
 
 \issue{ Linear control was adopted (for tractability) but clearly the nonlinearity is needed}
-\answer{The reviewer brings up a reasonable point about the formulation of the wireless control problem.
-Indeed, many practical systems feature both nonlinear dynamics and potentially non-linear controls.
-As stated, the proposed approach focuses on linear systems and controls so that we may derive analytic expressions for the control-aware adaptive packet delivery rates.
-In practice, nonlinear systems may be linearized for the purposes of obtaining such packet delivery rate targets in (23) and performing the proposed CALLS method as written.
-\blue{To demonstrate how this may be done, in the revised manuscript we have included additional simulations in Section VI in which we control a non-linear system using the CALLS method to schedule plant communications.
-In these simulations, while the actual dynamics and control being applied are nonlinear in nature, a linearization of the dynamics is used solely to derive the necessary PDR targets as in expression (23).
-Here, we demonstrate that the proposed method is applicable even in nonlinear systems.}   }
+\answer{The reviewer brings up a good point about the consideration of nonlinear control in additional the linear control that is discussed in this paper.
+We agree that a more complete discussion of this topic should include non-linear controllers.
+However, we first point out that many of the most commonly used controller in practice, e.g.
+LQR and PID control, are linear in nature and can be formulated in the manner discussed in the manuscript.
+Such controllers are used in both industrial and robotic systems.
+Therefore, we feel that the consideration of linear control not only allows for the precise and analytic derivation of PDR targets used in this work,  but also encompasses a large variety of practical control systems.
+The reviewer is correct, however, that more nonlinear control is also an important thing to consider.
+As the non-linearities further complicate the analysis performed here, we aim to study their use in this context in future work.
+We have added a sentence to this effect in the Conclusion section of the revised manuscript.
+}
 
@@
 \issue{  In simulation part, it would be great if the authors could compare the proposed CALLS method with other methods if possible, since there are also some papers discussing communication resource allocation for multiple control loops.}
-\answer{\red{To be written}}
+\answer{We thank the reviewer for this suggestion.
+Indeed, an interesting comparison would be to compare the proposed CALLS method with other scheduling methods for wireless control systems.
+We first point out that existing control-aware scheduling methods do not explicitly attempt to enable low latency or are developed for the IEEE 802.11ax architecture.
+In any case, we can adapt the existing typical scheduling approach of \emph{event-triggered control} for the low latency setting.
+In such a scheduling scheme, a device is scheduled to transmit only if its estimated state goes beyond some threshold value.
+Should this event occur, the device is scheduled with a fixed, high reliability packet delivery rate using a low latency scheduling method (i.e.
+assignment-based scheduling).
+In a sense, this scheme combines the selective scheduling approach of CALLS method with the standard fixed PDR of URLLC.
+
+In the revised manuscript, we include a comparison against an event-triggered scheduling protocol for the balancing board ball experiments in Section V-B.
+As can be seen, event-triggered scheduling indeed outperforms the fixed PDR scheduling but does achieve the same scale in terms of number of devices supported as the CALLS method, which uses \emph{adaptive} PDR targets that make explicit use of control system dynamics to schedule devices.
+We lastly point out that these comparisons were also performed with the inverted pendulum system.
+However, the event triggered scheduling was not able to adapt to the faster dynamics of the inverted pendulum to control many devices, and such comparisons are therefore left out of the revised manuscript.}
 
@@
 \issue{In your system, how do you obtain accurate channel state information? E.g. for AWGN channels, you need to test the noise variance. Is your scheme sensitive to channel estimation errors?}
-\answer{This is a good point raised by the reviewer. \red{To be written.}}
+\answer{This is a good point raised by the reviewer.
+In our simulations, we consider that channel state information is obtained via pilot signals and there is no estimation errors.
+In practice, however, such estimation errors may indeed exist.
+We first point out that the computation of adaptive packet delivery rate targets in equation (23) does not depend upon channel state information, only the estimated control state information.
+Therefore, the main control-awareness component of the proposed scheduling protocol---namely Steps 3-6 in Algorithm 1--- is unaffected by inaccurate CSI estimation.
+Likewise, the assignment method-based scheduling in Step 8 does also not directly depend upon CSI information, as can be seen in the formulation of the assignment problem in equation (33).
+The only component that may be negatively impacted by estimation errors is Step 7,  the maximum MCS selection in equation (32).
+Here, the potential impact would be selecting an MCS that cannot meet the target PDR $\tdq_i$ due to an error in estimating $\bbh_{i,k}$.
+Such an effect can be mitigated by selecting a more conservative MCS selection to account for channel estimation errors---e.g.
+transmitting at an MCS $\mu_{i,k}(\bbvarsigma) - 1$ rather than $\mu_{i,k}(\bbvarsigma)$.
+Such a conservative scheme would tradeoff latency to the benefit of reliability or robustness.
+A detailed numerical study of this effect is outside the scope of this work, however, and left as a study of future work.
+In the revised manuscript, we have added a remark to the end of Section IV discussing this point.
+}
 
 \issue{The proposed system shall be compared to more existed typical scheduling scheme employed in control system.}
-\answer{\red{To be written}}
+\answer{We thank the reviewer for this suggestion.
+Indeed, an interesting comparison would be to compare the proposed CALLS method with other scheduling methods for wireless control systems.
+We first point out that existing control-aware scheduling methods do not explicitly attempt to enable low latency or are developed for the IEEE 802.11ax architecture.
+In any case, we can adapt the existing typical scheduling approach of \emph{event-triggered control} for the low latency setting.
+In such a scheduling scheme, a device is scheduled to transmit only if its estimated state goes beyond some threshold value.
+Should this event occur, the device is scheduled with a fixed, high reliability packet delivery rate using a low latency scheduling method (i.e.
+assignment-based scheduling).
+In a sense, this scheme combines the selective scheduling approach of CALLS method with the standard fixed PDR of URLLC.
+
+In the revised manuscript, we include a comparison against an event-triggered scheduling protocol for the balancing board ball experiments in Section V-B.
+As can be seen, event-triggered scheduling indeed outperforms the fixed PDR scheduling but does achieve the same scale in terms of number of devices supported as the CALLS method, which uses \emph{adaptive} PDR targets that make explicit use of control system dynamics to schedule devices.
+We lastly point out that these comparisons were also performed with the inverted pendulum system.
+However, the event triggered scheduling was not able to adapt to the faster dynamics of the inverted pendulum to control many devices, and such comparisons are therefore left out of the revised manuscript.}
 

### 2019-04-02 Learning_llc/globecom/problem_formulation.tex
Add files via upload

%!TEX root = root.tex

%%%%%%%%%%%%%%%%% F I G U R E %%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\input{../images/tikz_wncs4}
%\caption{Wireless control system with $m$ independent systems. Each system contains a sensor that measure state information, which is transmitted to the controller over a wireless channel. The state information is used by the controller to determine control policies for each of the systems.The communication is assumed to be wireless in the uplink and ideal in the downlink.}
%\label{fig_wcs}
%\end{figure}

We consider a series of $m$ control systems operating over a shared wireless channels. The state of plant $i$ at control cycle index $k$ is given by the variable $\bbx^k_i \in \reals^p$. At each control/scheduling cycle, the sensor measures the state $\bbx^k_i$ and transmits it over a wireless channel to a common base station (BS) that is co-located with the controller. Given the state information, the controller determines the necessary control input is fed back to the plant. This is referred to as the closed-loop configuration of the control cycle. Given the noisy nature of the wireless channel, there is the potential for the communications packet containing the state information to be dropped, resulting in an open-loop configuration of the control cycle. We may model the linear dynamics of the wireless control system for plant $i$ as
%
\begin{equation}\label{eq:system}
	\bbx_i^{k+1} = \left\{ \begin{array}{ll} \hbA_i \bbx^k_i + \bbw^k &\text{if packet received} \\ \mathring{\bbA}_i \bbx^k_i + \bbw^k & \text{otherwise} \end{array} \right.,
\end{equation}
%
where $\hbA_i \in \reals^{p\times p}$ is the closed loop gain, $\mathring{\bbA}_i \in \reals^{p\times p}$ is the open loop gain, and $\bbw^k \in \reals^p$ is zero-mean i.i.d. disturbance process with covariance $\bbW$. The closed loop and open loop gains may reflect, e.g., controlled dynamics using accurate and estimated state information, respectively. We assume that the closed loop gains are preferable to the open loop gain, i.e. $\lambda_{\max}(\hbA_i) < \lambda_{\max}(\mathring{\bbA}_i)$. Further note this model restricts its attention to wireless connections in uplink of the control loop, while downlink is assumed to occur over an ideal channel---i.e. no packet drops.
%

Given this dynamical model of the wireless control systems, the communications goal is to allocate radio resources among the various plants to maintain strong performance across all the systems. To do so, we present a generic frequency and time division multiplexing scheduling architecture with which the BS allocates scheduling resources to the plants. A scheduling window occupies the uplink of a single cycle in the control loop; the total length of this scheduling window is subject to a tight low-latency bound. Transmissions are scheduled by the BS across $n$ different channels occupying different frequency bands. Multiple transmissions scheduled in the same channel will occur in sequence, while transmissions scheduled in different channels may occur simultaneously. Denote by $\bbvarsigma_i \in \{0,1\}^n$ a binary vector whose $j$th element $\varsigma_{i,j}$ is 1 if the $i$th device transmits in the $j$th channel, and 0 otherwise. Further denote for each device a data rate selection $\mu_i \in [\mu_{\min}, \mu_{\max}]$. These two scheduling parameters together define the scheduling decision made for the $i$ plant. This architecture reflects generalizes many practical scheduling-based multiple access wireless protocols, such as those used in LTE \cite{sesia2011lte} , 5G \cite{agiwal2016next}, and next-generation WiFi IEEE 802.11ax \cite{liu2014ieee}. An illustration of $m=4$ users making multiple transmission across $n=3$ channels is shown in Figure \ref{fig_multi}.

The achieved communications performance by a given scheduling decision can be formulated as follows. We first define $\bbh^k_{i} \in \reals^n_+$ to be the set of fading channel states experienced by device $i$ at cycle $k$, where the $j$ element $h^k_{i,j}$ is the fading channel gain in channel $j$. We assume that these channel conditions do not change over the course of a scheduling window. In any given channel with fading state $\bbh$, we define a function $q(h,\mu)$ that returns the packet delivery rate (PDR), or the probability of successful transmission of the packet, when transmitting with data rate $\mu$. Likewise, we define a function $\tau(\mu)$ that returns the transmission time to transmit a packet of fixed length with data rate $\mu$. These two functions play a critical role in designing low-latency wireless control systems, as they allow us to explore the trade-off between PDR and transmission time and the resulting effect on control system performance. We may consider that the functions $q(h, \mu)$ and $\tau(\mu)$ both get smaller as we increase data rate $\mu$, i.e.
%
\begin{equation}
\mu' > \mu \implies q(h, \mu) \leq q(h, \mu'), \quad \tau(\mu') \leq \tau(\mu).
\end{equation}
%
Thus, by increasing the data rate we may reduce the transmission time to satisfy latency constraints, but at the cost of control system performance, as illustrated by the switched dynamics in \eqref{eq:system}.

\subsection{Optimal scheduling design}\label{sec_opt}

We are interested in designing scheduling policies that optimize control performance, subject to the strict low latency constraints of the system. To do so, we first formulate the global control-based performance given a scheduling decision. Collect in the matrix $\bbSigma \in \{0,1\}^{n \times m}$ all of the channel transmission vectors $\bbvarsigma_i$ for $i=1,\hdots,m$ and collect in the vector $\bbmu \in [\mu_{\min},\mu_{\max}]^m$ the data rates $\mu_i$ for $i=1,\hdots,m$. Given that a device may transmit in multiple channels within a single scheduling cycle, the probability of successful transmission can be given as the probability that the transmission was successful in at least one channel, i.e.
%
\begin{equation}\label{eq_psr}
\tdq(\bbh_i, \bbvarsigma_i, \mu_i) := 1 - \prod_{j=1}^n \left(1 - \varsigma_{i,j} q(h_{i,j}, \mu_i)\right). 
\end{equation}
%

The total delivery rate in \eqref{eq_psr} can be viewed as the probability of receiving the packet and experiencing the closed loop dynamics in \eqref{eq:system}. Now, to evaluate the performance of a given plant at a particular state $\bbx$, define a quadratic Lyapunov function $L_i(\bbx) := \bbx^T \bbP_i \bbx$ with some positive definite matrix $\bbP_i \in \reals^{p \times p}$. Such a function can be used to evaluate performance or stability of the control system. Because the control system evolves in a random manner, the cost of a given scheduling decision $\{\bbvarsigma_i, \mu_i\}$ for the $i$th plant can be formulated as the \emph{expected future Lyapunov cost} under such a schedule. As the probability of closing the loop in \eqref{eq:system} is given by $\tdq(\bbh^k_i, \bbvarsigma_i, \mu_i)$, we may write this expected future cost as 
%
\begin{align}\label{eq_ex_cost}
J_i(\bbx_i, \bbh_i, \bbvarsigma_i, \mu_i) :=& \quad \E \left[ L_i(\bbx^{k+1}_i) \mid \bbx^k_i = \bbx_i, \bbh^k_i = \bbh_i \right] \\
=&  \quad  \tdq(\bbh_i, \bbvarsigma_i, \mu_i) (\hbA_i \bbx_i)^T \bbP_i (\hbA_i \bbx_i) \quad +  \nonumber   \\
 &(1- \tdq(\bbh_i, \bbvarsigma_i, \mu_i))  (\mathring{\bbA}_i \bbxi)^T \bbP_i (\mathring{\bbA}_i \bbx_i) \nonumber \\
 &+  \quad \Tr(\bbP_i \bbW_i). \nonumber
\end{align}
%
Observe that the local control cost for the $i$th plant $J_i(\bbx^k_i, \bbh^k_i, \bbvarsigma_i, \mu_i)$ is a function of both the system \emph{states}---the fading channel $\bbh^k_i$ and control state $\bbx^k_i$---and the scheduler \emph{actions}---channel selection $\bbvarsigma_i$ and data rate $\mu_i$. The objective is to choose the actions $\bbvarsigma_i$ and $\mu_i$ that minimizes the cost relative to states $\bbh^k_i$ and $\bbx^k_i$.

In addition to minimizing a control cost, we must make scheduling decisions that respect the low-latency requirements of the system. To formulate this constraint, consider the \emph{total} time of a global scheduling decision $\bbSigma, \bbmu$ of channel $j$ as the sum of all active transmissions, i.e.
%
\begin{equation} \label{eq_total_time}
\tilde{\tau}_j(\bbSigma, \bbmu) :=   \sum_{i=1}^m \varsigma_{i,j} \tau(\mu_i).
\end{equation}
%

Combining all the local costs for plants $i=1,\hdots,m$ in \eqref{eq_ex_cost} with the latency constraints for all channels $j=1,\hdots,n$ in \eqref{eq_total_time}, we may define the optimal scheduling design problem. Because we are interested in long-term, or average, performance across random channels and control states, we optimize with respect to expected costs and probabilistic constraints. Collect all channel vectors $\bbh_i$ in a matrix $\bbH \in \reals_+^{n \times m}$ and states $\bbx_i$ in a matrix $\bbX \in \reals^{p \times n}$. Consider a scheduling policy $\bbp(\bbH, \bbX) := \{ \bbSigma, \bbmu\}$ that, given a set of channel states $\bbH$ and control states $\bbX$, returns a schedule defined by the channel selection matrix $\bbSigma$ and data rate selection vector $\bbmu$. The optimal low-latency constrained scheduling policy for the wireless control systems is the one which solves the program
%
\begin{alignat}{2} \label{eq_problem}
   J^* := &  \min_{\bbp(\bbH, \bbX)}  \E_{\bbH,\bbX} \left[ \sum_{i=1}^m J_i(\bbx_i,\bbh_i, \bbvarsigma_i, \mu_i) \right],             \\
        &  \st           \quad           \mathbb{P}_{\bbH,\bbX} \left( \tilde{\tau}_j(\bbSigma, \bbmu) \leq t_{\max} \right)   \geq 1-  \delta \qquad \forall j,   \nonumber \\
        &     \                       \bbp(\bbH,\bbX) := \{ \bbSigma \in  \{0,1\}^{n \times m}, \bbmu \in [\mu_{\min},\mu_{\max}]^m \}.   \nonumber%   
\end{alignat}
% 
In \eqref{eq_problem}, we minimize the average cost over the distribution of channel and control states, subject to the condition that the probability of violating the latency constraint over the distribution of states is less than some small value $\delta$. The above scheduling problem can be viewed as a constrained statistical learning problem---a connection made for a more generic class of resource allocation problems in \cite{eisen2018learninga}. While such a problem characterizes the optimal scheduling decision for the latency-constraint wireless control system, finding solutions to such a problem is a significant challenge. This is due to a number of complexities in \eqref{eq_problem}, namely: (i) it requires functional optimization, (ii) it contains explicit constraints, and (iii) we typically do not have analytic forms for the functions and distributions in \eqref{eq_problem}. The first of these complexities can be resolved using a standard technique in statistical learning, discussed next in Section \ref{sec_param}. The latter two of these complexities are discussed and resolved later in Sections \ref{sec_primal_dual} and \ref{sec_model_free}, respectively.

### 2019-04-02 Learning_llc/globecom/root.tex
Add files via upload

\documentclass[conference,10pt]{IEEEtran}
 \IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\graphicspath{{fig/}{jpeg/}}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subcaption}
\usepackage[outdir=./]{epstopdf}
%\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage{comment}
\input{../mysymbol.sty}
\input{../my_sections.tex}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float,environ}
\usetikzlibrary{matrix} % for block alignment
\usetikzlibrary{decorations.markings} % for arrow heads
\usetikzlibrary{calc} % for manimulation of coordinates
\usetikzlibrary{decorations.pathmorphing,snakes, positioning, decorations.pathreplacing}

\usetikzlibrary{fit}
%\input{figures/tikh_styles}

\usepackage{pgfplots}
\usepackage{color}
%\usepackage{setspace}
%\usepackage{hyperref}
\usepackage{multirow}
%\usepackage{rotating}
%\usepackage{comment}
%\usepackage{flushend}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage[affil-it]{authblk}
%\usepackage{bm}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{enumerate}
%\usepackage{morefloats}
%\usepackage{setspace}

\makeatletter
\newsavebox{\measure@tikzpicture}
\NewEnviron{scaletikzpicturetowidth}[1]{%
  \def\tikz@width{#1}%
  \def\tikzscale{1}\begin{lrbox}{\measure@tikzpicture}%
  \BODY
  \end{lrbox}%
  \pgfmathparse{#1/\wd\measure@tikzpicture}%
  \edef\tikzscale{\pgfmathresult}%
  \BODY
}
\makeatother

\newcommand{\QED}{\hfill\ensuremath{\blacksquare}}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}

%\IEEEoverridecommandlockouts 
\def \vneg {\vspace{-.5cm}}

\def\whp{\text{w.h.p}}
\def\E{\mathbb{E}}
\def\P{\mathbb{P}}
\def\supp{\text{support}}
\def\Tr{\text{Tr}}
\def\vec2{\text{vec}}

% Resource allocation functions
\def \fph  {\bbf\big(\bbp(\bbh), \bbh\big)}
\def \Efph {\mbE\big[\fph\big]}
%
% Optimal power allocation functions
\def \fphstar  {\bbf\big(\bbp^*(\bbh),\bbh\big)}
\def \Efphstar {\mbE\big[\fphstar\big]}
%
% Slater's condition power allocation functions
\def \fphzero  {\bbf\big(\bbp_0(\bbh),\bbh\big)}
\def \Efphzero {\mbE\big[\fphzero\big]}
%
% Parametrized power allocation functions
\def \fphih  {\bbf\big(\bbphi(\bbh,\bbtheta),\bbh\big)}
\def \Efphih {\mbE\big[\fphih\big]}
%
% Parametrized power allocation functions indexed by sample index k
\def \fpihk {\bbf\big(\bbh_k,\bbphi(\bbh_k,\bbtheta_k)\big)}
%
% Lagrangian
\def \Lagphi {\ccalL_{\bbphi}(\bbtheta,\bbx, \bblambda, \bbmu)}

%\addtolength{\textwidth}{10mm}
%\addtolength{\evensidemargin}{-5mm}
%\addtolength{\oddsidemargin}{-5mm}
%\addtolength{\textheight}{10mm} 
%\addtolength{\topmargin}{-5mm}

%\renewcommand \blue[1]{}  \renewcommand \red[1]{}

\newtheorem{assumption}{\hspace{0pt}\bf Assumption}
\newtheorem{lemma}{\hspace{0pt}\bf Lemma}
\newtheorem{proposition}{\hspace{0pt}\bf Proposition}
\newtheorem{example}{\hspace{0pt}\bf Example}
\newtheorem{observation}{\hspace{0pt}\bf Observation}
\newtheorem{theorem}{\hspace{0pt}\bf Theorem}
\newtheorem{corollary}{\hspace{0pt}\bf Corollary}
\newtheorem{fact}{\hspace{0pt}\bf Fact}
\newtheorem{remark}{\hspace{0pt}\bf Remark}
\newtheorem{test}{\hspace{0pt}\it Test Case}
\newtheorem{definition}{\hspace{0pt}\bf Definition}
\newtheorem{conj}{\hspace{0pt}\bf Conjecture}

% Title.
% ------
\begin{document} 
%
% Single address.
% ---------------

\title{Primal-Dual Learning for Resource Allocation \\ in Low Latency Wireless Systems}
\author{Mark Eisen$^*$ \quad Mohammad M. Rashid$^\dagger$ \quad 
			Dave Cavalcanti$^\dagger$ \quad Alejandro Ribeiro$^*$
\thanks{{Supported by Intel Science and Technology Center for Wireless Autonomous Systems. The authors are with the $(*)$Department of Electrical and Systems Engineering, University of Pennsylvania and $(\dagger)$Wireless Communications Research, Intel Corporation. Email: maeisen@seas.upenn.edu, mamun.rashid@intel.com, dave.cavalcanti@intel.com, aribeiro@seas.upenn.edu}.}}

\maketitle

%\thispagestyle{empty}
%
\begin{abstract}
We consider the problem of scheduling transmissions over low-latency wireless communication links to control various control systems. Low-latency requirements are critical in developing wireless technology for industrial control, but are inherently challenging to meet while also maintaining reliable performance. An alternative to ultra reliable low latency communications (URLLC) is a framework in which reliability is adapted to control system demands. We formulate the control-aware scheduling problem as a constrained statistical optimization problem in which the optimal scheduler is a function of current control and channel states. The scheduler is parameterized with a deep neural network, and the constrained problem is solved using techniques from primal-dual learning. Furthermore, these updates are model-free in that they do not require explicit knowledge of channels models or performance metrics. The resulting control-aware deep scheduler is evaluated in empirical simulations and strong performance is shown relative to other heuristic scheduling methods.
\end{abstract}
%

\begin{IEEEkeywords}
wireless control, low-latency, deep learning, primal-dual
\end{IEEEkeywords}

\section{Introduction}

The recent advances in wireless technology and automation have given rise to efforts in integrating wireless communications in autonomous environments, particularly in industrial control settings where the scale of wired networks is proving increasingly costly \cite{zand2012wireless, wollschlaeger2017future}. The analysis of control systems operating over wireless communication links is thus an integral apart in enabling these wireless industrial automation applications. However, the performance specifications of these applications demands the design of wireless networks that can meet both the high reliability and low latency demands of the system \cite{zand2012wireless, varghese2014wireless, weiner2014design, popovski2018wireless, bennis2018ultra}. The problem of ultra reliable low latency communications (URLLC) is inherently challenging as the physical medium of wireless communication trades off these two performance criteria, making it hard to meet both reliability and latency demands.

One promising direction in enabling low latency communications involves specific developments in radio resource allocation, or scheduling. For low latency applications, traditional delay-aware schedulers ~\cite{wu2014analysis, lu1999fair, andrews2001providing} have been employed, in addition to more recent URLLC techniques based on various forms of diversity \cite{swamy2015cooperative, popovski2018wireless, nielsen2018ultra, ashraf2018dynamic}---all of which are agnostic to the control system. However, due to the physical limitations of the wireless channel, it is often necessary to use information from the control system to make proper use of scheduling resources in meeting latency requirements. While there exist numerous ways in which control system information is incorporated into ``control-aware'' scheduling methods \cite{Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal, GatsisEtal15,demirel2018deepcas,leong2018deep}, these are agnostic to latency requirements of the system. More recent work \cite{eisen2018control} looks at heuristic based scheduling methods that are both control and latency aware.

Aside from traditional heuristic based scheduling methods, tools from machine learning are being incorporated into making intelligent scheduling and resource allocation decisions. The work in \cite{eisen2018learninga} builds a framework for solving a generic set of resource allocation problems by interpreting resource allocation as a constrained statistical learning problem. This leads a natural use of learning models, such as deep neural networks (DNNs), for designing schedulers. Recent advancements apply techniques from both reinforcement learning and deep learning for control-aware scheduling in simple systems \cite{demirel2018deepcas,leong2018deep}. The goal in this work is to use the framework of \cite{eisen2018learninga} to employ so-called primal-dual deep learning to tackle develop novel scheduling techniques that are both control-aware and latency constrained.

This paper is organized as follows. We formulate the wireless control system in which state information is communicated to the control over a wireless channel as a switched dynamical system (Section \ref{sec_problem}). We further discuss the scheduling architecture under which we can allocate channels and data rates between plants. With this formulation, we formulate the optimal scheduling problem that minimizes a control cost under latency constraints (Section \ref{sec_opt}). As this problem is a statistical learning problem, it can be parameterized with a deep neural network (Section \ref{sec_param}). To handle the constraints, we convert the problem to the Lagrangian dual problem and present a primal-dual learning method to find the optimal parameters (Section \ref{sec_primal_dual}). We further discuss ways in which the primal-dual method can be approximated without explicit model knowledge (Section \ref{sec_model_free}). The performance of the learned control-aware scheduling method is analyzed in a numerical simulation of a control system with latency constrained and compared other heuristic scheduling methods (Section \ref{sec_numerical_results}).

%\input{Problem_formulation_icassp}

\section{Wireless Control Systems}\label{sec_problem}

%
\begin{figure}
\centering
\input{../images/multiplex.tex}
\caption{Scheduling architecture of $m=4$ users---colored in green, blue, red, and yellow---across $n=3$ channels.}
\label{fig_multi}
\end{figure}
%

\input{problem_formulation}

\subsection{Deep learning parameterization}\label{sec_param}

The scheduling problem in \eqref{eq_problem} is computationally challenging because it requires finding a policy---or \emph{function}---$\bbp(\bbH,\bbX)$. In statistical learning, or regression, problems the regression function is replaced by some given parameterization $\bbphi(\bbH,\bbX,\bbtheta)$ that is defined with some finite dimensional parameter $\bbtheta \in \reals^q$. Their exist a wide variety of choices of this parameterization, but in modern machine learning problems the \emph{deep neural network (DNN)} is commonly employed. This is due to the fact the DNN can be shown both empirically and analytically to contain strong representative power and generalization ability, meaning that it can approximate almost any function well. A DNN is defined as a composition of $L$ layers, each of which consisting of a linear operation followed by a point-wise nonlinearity---also known as an activation function.  More specifically, the layer $l$ is defined by the linear operation $\bbW_l \in \reals^{q_{l-1} \times q_l}$ followed by a non-linear activation function $\mathbb{\sigma}_{l}: \reals^{q_l} \rightarrow \reals^{q_l}$. Common choices of activation functions $\mathbb{\sigma}_{l}$ include a sigmoid function or a rectifier function (commonly referred to as ReLu). Given an input from the $l-1$ layer $\bbw_{l-1} \in \reals^{q_{l-1}}$, the resulting output $\bbw_{l} \in \reals^{q_l}$ is then computed as $\bbw_l := \mathbb{\sigma}_{l}(\bbW_l \bbw_{l-1})$.  The full DNN-parameterization of the scheduling policy is then defined as an $L$-layer DNN whose input at the initial layer is the concatenation of states $\bbw_0 := [\vec2(\bbH); \vec2(\bbX)]$, i.e.
%
\begin{equation}\label{eq_policy_dnn}
\bbphi(\bbh,\bbtheta) := \mathbb{\sigma}_{L}(\bbW_{L} (\mathbb{\sigma}_{L-1}(\bbW_{L-1}(\hdots(\mathbb{\sigma}_{1}(\bbW_{1}\bbw_0)))))).
\end{equation}
%

The parameter vector $\bbtheta \in \reals^q$ that defines the DNN is then the entries of $\{ \bbW_l \}_{l=1}^L$ with $q = \sum_{l=1}^{L-1} q_l q_{l+1} $. Further note that we can easily construct an activation function at the final layer $\mathbb{\sigma}_L$---or the \emph{output layer}---such that the outputs $\bbphi(\bbh,\bbtheta)$ are in the space $\{0,1\}^{n \times m} \times [\mu_{\min},\mu_{\max}]$ that contains possible schedules. With this DNN parameterization, the control-aware scheduling problem can be rewritten as
%
\begin{alignat}{2} \label{eq_param_problem}
   J_{\bbphi}^* := &  \min_{\bbtheta \in \reals^q}  \E_{\bbH,\bbX} \left[ \sum_{i=1}^m J_i(\bbx_i,\bbh_i, \bbalpha_i, \mu_i) \right],             \\
        &  \st           \quad           \mathbb{P}_{\bbH,\bbX} \left( \tilde{\tau}_j(\bbSigma, \bbmu) \leq t_{\max} \right)   \geq 1-  \delta \qquad \forall j,   \nonumber \\
        &     \                       \bbphi(\bbH,\bbX,\bbtheta) := \{ \bbSigma \in  \{0,1\}^{n \times m}, \bbmu \in [\mu_{\min},\mu_{\max}]^m \}.   \nonumber%   
\end{alignat}
% 
Observe in \eqref{eq_param_problem} that the optimization is performed over $\bbtheta$ rather than the scheduling policy directly. In other words, we look for the interlayer weights that define a DNN that minimizers the total control cost while satisfying the latency constraints. We proceed then to discuss a learning method that can find solutions to the constrained optimization problem in \eqref{eq_param_problem}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%  S  E  C  T  I  O  N     %%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Primal-Dual Learning}\label{sec_primal_dual}

\input{scheduling_alg}

 \begin{figure*}[t]
\centering
\includegraphics[width=0.45\linewidth, height=.18\textheight, keepaspectratio]{../images/cost_plot.eps} \qquad
\includegraphics[width=0.45\linewidth, height=.18\textheight, keepaspectratio]{../images/time_plot.eps} 
\caption{Convergence of (left) objective function value, and (right) transmission time for a low-latency, control aware scheduling policy over the learning process. The DNN parameterized scheduling policy obtains feasible latency-contained schedules ($t_{\max} = 5 \times 10^{-4}$) that converges to a local minimum.}\label{fig_simple_results}
\end{figure*}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   S   E   C   T   I   O   N   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \section{Simulation Results}\label{sec_numerical_results}

We perform a series of simulations on latency-constrained wireless control systems to evaluate the performance the learning method in and the resulting control-aware scheduling policies. We generate a series $m=20$ plants with closed-loop gains $\hbA_i \sim \text{Uniform}(0.8,0.95)$ and open-loop gains $\mathring{\bbA}_i \sim \text{Uniform}(1.01,1.3)$. The variance for all system noise $\bbw_i$ is set to be $W=1$. All such plants send their state information over a shared wireless channel with $n=5$ independent channels with a total latency constraint of $t_{\max} = 0.5$ ms. A latency bound of this order is typical of industrial control systems such as printing machines and presses \cite{ashraf2016ultra}. We further assume that the states of the plants are confined to the box $[-10,10]$. The DNN is given an architecture of 4 layers of sizes 256, 128, 64, and 32, all using a ReLU activation function.

With the scheduling architecture given in Figure \ref{fig_multi} for 5 channels and 20 plants, at each control scheduling interval each plant is given a data rate $\mu_i$ and a set of channels to transmit on. In our simulations, we use the modulation and coding schemes (MCS) of the next-generation IEEE 802.11ax Wi-Fi protocol as a representative architecture for data rate selection and packet error rate computation. As such, the continuous data rates $\mu_i$ are selected in an interval of $[1.6, 8]$ and rounded down to the nearest discrete MCS selection given in 802.11ax---see \cite{liu2014ieee} for details on the MCS tables given in this protocol. The corresponding transmission time $\tau(\mu)$ is then calculated assuming a fixed packet size of 100 bytes and the packet delivery rate $q(h,\mu)$ is computed using the associated AWGN error curve (scaled by the effective SNR given channel conditions).

 \begin{figure}
\centering
\includegraphics[height=.22\textheight]{../images/system_simulation.eps}
\caption{Evolution of plant states over 1000 control cycles using both the (top) DNN-based control aware scheduler and (b) a round robin scheduler.}\label{fig_system}
\end{figure}

In Figure \ref{fig_simple_results} we show the training process for the control-aware scheduler using the primal-dual scheduler given in Algorithm \ref{alg:learning}. In the left figure, we show the aggregate Lyapunov control cost over the course of $40,000$ learning iterations and in the right figure, show the transmission time of each channel (shown in different colored lines). As can be seen the in the figure, the policies initially output infeasible schedules that are beyond the latency limit but eventually converge to scheduling decisions that respect latency requirements. The control objective continues to decrease after feasibility is obtained and eventually converges to a local minimum value.

We proceed to simulate the performance of the control systems using the scheduling policy obtained at the end of the learning process and compare against a standard, round-robin scheduling policy---i.e. control-agnostic. Because the constraint in \eqref{eq_problem} is satisfied only with probability $1-\delta$, we ensure latency requirements are met at runtime by restricting all transmissions that are scheduled after the latency bound. In Figure \ref{fig_system} we show the state values over 1000 control cycles for the DNN-based scheduler (top figure) and the round robin scheduler (bottom figure). It can be observed that, overall, the plants are kept in better states using the DNN scheduler than with the round robin scheduler. This representative simulation highlights the ability of control-aware scheduling to better meet strict latency demands and, furthermore, highlight the ability of a DNN to properly encode such a constraints and objective in its dense architecture.

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   S   E   C   T   I   O   N   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \section{Conclusion}
In this paper we develop a control-communication control-aware approach towards scheduling for low-latency wireless control systems. To handle the challenge of achieving high reliability performance with limited scheduling resources, we formulate a control-aware scheduling problem in which reliability is adapted to control and channel states and plant dynamics. This problem takes the form of a constrained statistical learning problem, in which solutions can be found by parameterized the scheduling policy with a deep neural network and applying a primal-dual learning framework to obtain the optimal weights of the neural network. Numerical simulations showcase the ability of the primal-dual method to find scheduling policies that outperform simple control-agnostic scheduling procedures.

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\urlstyle{same}
\bibliographystyle{../IEEEbib}
\bibliography{../wireless_ll_control,../scheduling_control}

\end{document}

### 2019-04-02 Learning_llc/globecom/scheduling_alg.tex
Add files via upload

%!TEX root = root.tex

Finding the DNN layer weights $\bbtheta$ that provide good solutions to \eqref{eq_param_problem} requires the solving of a constraint learning problem. The standard approach of gradient-based optimization methods cannot be applied directly here to the presence of the latency constraints. To proceed then, we must formulate an unconstrained problem that captures the form of \eqref{eq_param_problem}. A naive penalty-based reformulation will introduce a similar but fundamentally different problem, so we thus opt for constructing a Lagrangian dual problem. For notational convenience, moving forward we employ the following shorthands for the state variables, aggregate Lyapunov function, latency constraint functions, respectively:
%
\begin{align}
\bbw &:= [\vec2(\bbH); \vec2(\bbX)], \\
f(\bbphi(\bbw,\bbtheta), \bbw) :&= \sum_{i=1}^m J_i(\bbx_i, \bbh_i, \bbvarsigma_i, \mu_i), \\
g_j(\bbphi(\bbw,\bbtheta),\bbw) &:=  \mathbb{I}\left[ \tdtau_j(\bbSigma,\bbmu) \leq t_{\max}\right] - (1 - \delta)
\end{align}
%
We introduce the nonnegative dual variables $\bblambda \in \reals_+^n$ associated with the vector of constraint functions $\bbg(\bbp(\bbw,\bbtheta),\bbw) := [g_1(\cdot); \hdots; g_n(\cdot)]$, and form the Lagrangian as
%
\begin{align}\label{eq_param_lagrangian}
   \ccalL(\bbtheta,\bblambda) &:=   \E_{\bbw} \left[ f(\bbphi(\bbw,\bbtheta), \bbw) - \bblambda^T \bbg(\bbphi(\bbw,\bbtheta),\bbw) \right].
\end{align}
%
The Lagrangian in \eqref{eq_param_lagrangian} penalizes constraint violation through the second term. Note, however, that the penalty is scaled by the dual parameter $\bblambda$. The so-called Lagranian dual problem is one in which both the primal variable $\bbtheta$ is simultaneously minimized while the dual parameter $\bblambda$ is maximized. Such a problem can be written with the saddle point formulation
%
\begin{align}\label{eq_param_dual0}
   D_{\bbphi}^* &:= \max_{\bblambda \geq \bb0} \min_{\bbtheta}\ccalL(\bbtheta,\bblambda).
\end{align}
%
The dual optimum $D_{\bbphi}^*$ is the best approximation of the form in \eqref{eq_param_lagrangian} we can have of $J_{\bbphi}^*$. In fact, under some standard assumptions on the problem and assuming a sufficiently dense DNN architecture, we can formally bound the difference between $D_{\bbphi}^*$ and $J^*$ to be proportional to the approximation capacity of the DNN $\bbphi(\bbH,\bbX,\bbtheta)$---see \cite{eisen2018learninga} for details on this result. Thus, we may say that, up to some approximation, solving the unconstrained problem in \eqref{eq_param_dual0} is equivalent to solving the constrained problem in \eqref{eq_param_problem}.

With the unconstrained saddle point problem in \eqref{eq_param_dual0}, we may perform standard gradient-based optimization methods to obtain solutions. The max-min structure necessitates the use of a \emph{primal-dual} learning method, in which we iteratively update both the primal and dual variable in \eqref{eq_param_lagrangian} to find a local stationary point of the KKT conditions of \eqref{eq_param_problem}. Consider a learning iteration index $t=0,1,\hdots$ over which we define a sequence of primal variables $\{\bbtheta_t\}$ and dual variables $\{\bblambda_t\}$. At index $t$, we determine the value of next primal iterate $\bbx_{t+1}$ by adding to the current iterates the corresponding partial gradients of the Lagrangian in \eqref{eq_param_lagrangian} $\nabla_{\bbtheta} \ccalL$, i.e., 
%
\begin{align}
%\bbtheta_{t+1} &= \bbtheta_t + \alpha_t \left[  \nabla_{\bbtheta}\E_{\bbw} f(\bbphi(\bbw,\bbtheta_t),\bbw) \right. \nonumber \\
%&\qquad  \left. - \ \nabla_{\bbtheta}\E_{\bbw} \bbg(\bbphi(\bbw,\bbtheta_t),\bbw)\bblambda_t \right] \label{eq_pd_update1},
\bbtheta_{t+1} \! &= \bbtheta_t \! - \! \alpha_t  \nabla_{\bbtheta}\E_{\bbw}  \left[ f(\bbphi(\bbw,\bbtheta_t),\bbw) \! - \! \bblambda_t^T \bbg(\bbphi(\bbw,\bbtheta_t),\bbw) \right] \label{eq_pd_update1},
\end{align}
%
where we introduce $\alpha_t >0$ as a scalar step size. We subsequently perform a corresponding partial gradient update to compute the dual iterate $\bblambda_{t+1}$, i.e.
%
\begin{align}
\bblambda_{t+1} &= \left[ \bblambda_t - \beta_t \E_{\bbw} \bbg(\bbphi(\bbw,\bbtheta_{t+1}),\bbw) \right]_+,\label{eq_pd_update2}
\end{align}
%
with associated step size $\beta_t >0$. Observe in \eqref{eq_pd_update2} that we additionally project onto the positive orthant to maintain the nonnegative constraint on $\bblambda$. The gradient primal-dual updates in \eqref{eq_pd_update1} and \eqref{eq_pd_update2} successively move the primal and dual variables towards maximum and minimum points of the Lagrangian function, respectively.

\subsection{Model-free updates}\label{sec_model_free}

The updates in \eqref{eq_pd_update1}-\eqref{eq_pd_update2} cannot, in general, be applied exactly. To see this, observe that computing the gradients in \eqref{eq_pd_update1} requires computing the gradient of $J_i(\cdot)$---which depends on PDR function $\tdq(\cdot)$---and the gradient of an indicator of  transmission length function $\tdtau(\cdot)$. In practical systems, we do not typically have easily available analytic forms for these functions to take gradients. Furthermore, both the updates in \eqref{eq_pd_update1} and \eqref{eq_pd_update2} require to take the expectation over the distribution of states $\bbx$ and $\bbh$. These, too, are often unknown in practice. However, there exist standard ways of approximating the updates with stochastic, \emph{model-free} updates that do not require such knowledge. Most popular among these is the policy gradient approximation \cite{sutton2000policy}.

To compute a policy gradient update, we consider the scheduling parameters $\bbSigma$ and $\bbmu$ are drawn stochastically from a distribution with given form $\pi_{\bbphi(\bbw,\bbtheta)}$ whose parameters are given by the output of the DNN $\bbphi(\bbw,\bbtheta)$---e.g. the mean and variance of a normal distribution. Using such a stochastic policy, it can be shown that an unbiased estimators of the gradients in \eqref{eq_pd_update1} and \eqref{eq_pd_update2} can be formed as,
%
\begin{align}\label{eq_policy_gradient_s}
\widehat{\nabla_{\bbtheta}} \mathbb{E}_{\bbw} f(\bbphi(\bbw,\bbtheta),\bbw) \!&=\!  f( \hbp_{\bbtheta},\hbw) \nabla_{\bbtheta} \log \pi_{\bbphi(\hbw,\bbtheta)}(\hbp_{\bbtheta} ) \\
\widehat{\nabla_{\bbtheta}} \mathbb{E}_{\bbw} \bbg(\bbphi(\bbw,\bbtheta),\bbw)\! &= \! \bbg( \hbp_{\bbtheta},\hbw) \nabla_{\bbtheta} \log \pi_{\bbphi(\hbw,\bbtheta)}(\hbp_{\bbtheta} )^T \\
\widehat{ \mathbb{E}}_{\bbw} \bbg(\bbphi(\bbw,\bbtheta),\bbw) &= \bbg(\hbp_{\bbtheta},\hbw), \label{eq_pg_3}
\end{align}
%
where $\hbw$ is a sampled state and $\hbp_{\bbtheta}$ is a sample drawn from the distribution $\pi_{\bbphi(\hbw,\bbtheta)}$. In practice, we may reduce the variance of these unbiased estimates by taking $B$ samples and averaging. Note that the updates here only require taking gradients of the log likelihoods rather than of the functions themselves. Thus, we can replace the updates in \eqref{eq_pd_update1} and \eqref{eq_pd_update2} with their model free counterparts, i.e.
%
\begin{align}
\bbtheta_{t+1} \! &= \bbtheta_t \! - \! \alpha_t  \widehat{\nabla_{\bbtheta}}\E_{\bbw}  \left[ f(\bbphi(\bbw,\bbtheta_t),\bbw) \! - \! \bblambda_t^T \bbg(\bbphi(\bbw,\bbtheta_t),\bbw) \right] \label{eq_spd_update1}, \\
\bblambda_{t+1} &= \left[ \bblambda_t - \beta_t \widehat{\E_{\bbw}} \bbg(\bbphi(\bbw,\bbtheta_{t+1}),\bbw) \right]_+.\label{eq_spd_update2}
\end{align}
%
%
The complete primal-dual learning algorithm is summarized in Algorithm \ref{alg:learning}. We conclude with a brief remark on state sampling.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%   A   L   G   O   R   I   T   H   M   %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\linespread{1.3}
\begin{algorithm}[t] \begin{algorithmic}[1]
\STATE \textbf{Parameters:} Policy model $\bbphi(\bbh,\bbtheta)$ and distribution form $\pi_{\bbh,\bbtheta}$ 
\STATE \textbf{Input:} Initial states $\bbtheta_0, \bblambda_0$
\FOR [main loop]{$t = 0,1,2,\hdots$}
      \STATE Draw samples $\{  \hbtheta, \hbh \}$, or in batches of size $B$
      \STATE Compute policy gradients  [ c.f. \eqref{eq_policy_gradient_s}-\eqref{eq_pg_3}]
      \STATE Update primal and dual variables  \ 
      \footnotesize
         \begin{align}
	\mkern-36mu \bbtheta_{t+1} \! &= \bbtheta_t \! - \! \alpha_t  \widehat{\nabla_{\bbtheta}}\E_{\bbw}  \left[ f(\bbphi(\bbw,\bbtheta_t),\bbw) \! - \! \bblambda_t^T \bbg(\bbphi(\bbw,\bbtheta_t),\bbw) \right] , [cf. \eqref{eq_spd_update1}]\nonumber \\
	\mkern-36mu \bblambda_{t+1} &= \left[ \bblambda_t - \beta_t \widehat{\E_{\bbw}} \bbg(\bbphi(\bbw,\bbtheta_{t+1}),\bbw) \right]_+ \!\![cf. \eqref{eq_spd_update2}]\	\nonumber 
\end{align}
\ENDFOR

\end{algorithmic}
\caption{Model-Free Primal-Dual Learning}\label{alg:learning} \end{algorithm}}

\begin{remark}\label{remark}\normalfont
In the gradient estimations in, e.g. \eqref{eq_policy_gradient_s}, we sample both the control states $\bbx$ and channel states $\bbh$. This assumes that such samples can be drawn i.i.d. While this may generally be true for the channel states $\bbh$, it will not be generally be true for the control states $\bbx$ in practice, due to the fact that the states evolve based on the switched dynamics in \eqref{eq:system}, which itself depends on the scheduling actions taken. A more precise way to model the statistics of the control states would be with a Markov decision process (MDP). The generalization of the presented techniques for this setting make up what is known as \emph{reinforcement learning} algorithms. In this work, we nonetheless assume that $\bbx$ can also be drawn i.i.d. from an approximate distribution and leave the full MDP formulation as the study of future work.
\end{remark}

### 2019-04-02 Learning_llc/images/multiplex.tex
Add files via upload

\begin{tikzpicture}
\draw[black,very thick] (-4,-1) rectangle (1,0);
\draw[black,very thick] (-4,0) rectangle (1,1);
\draw[black,very thick] (-4,1) rectangle (1,2);
\draw [step=.3, thick, draw= black, fill=black!10!green!30] (-4,-1) rectangle  (0,0);
\draw [step=.3, thick, draw= black, fill=black!10!blue!30] (0,-1) rectangle  (1,0);
\draw [step=.3, thick, draw= black, fill=black!10!red!30] (-4,0) rectangle  (-2,1);
\draw [step=.3, thick, draw= black, fill=black!10!blue!30] (-2,0) rectangle  (-1.2,1);
\draw [step=.3, thick, draw= black, fill=black!10!green!30] (-1.2,0) rectangle  (1,1);
\draw [step=.3, thick, draw= black, fill=black!10!yellow!30] (-4,1) rectangle  (1,2);
\node[align=center, scale =0.9] at (-1.5,-1.5) {Frame length};
\node[align=center, scale =0.9] at (-5,-.5) {Channel 3};
\node[align=center, scale =0.9] at (-5,.5) {Channel 2};
\node[align=center, scale =0.9] at (-5,1.5) {Channel 1};
\end{tikzpicture}

### 2019-04-02 Learning_llc/images/tikz_wncs4.tex
Add files via upload

--- 2019-02-01 Low_latency_control/images/tikz_random_access.tex
@@
-%!TEX root = root.tex
-\documentclass[]{standalone}
-
-\usepackage{amsmath} % nice math symbols
-
-\usepackage{color} % change text color        
-\usepackage{tikz}
-\usetikzlibrary{matrix} % for block alignment
-\usetikzlibrary{arrows, decorations.markings} % for arrow heads
-\usetikzlibrary{calc} % for manimulation of coordinates
-\usetikzlibrary{decorations.pathmorphing}
-\usetikzlibrary{shapes}
-
-\usetikzlibrary{fit}
-
-%\usetikzlibrary{backgrounds}
-
-\begin{document}
-% TikZ styles for drawing
-
 \pgfdeclarelayer{bg0}    % declare background layer
@@
 \tikzstyle{dash_block} = [draw,rectangle,dashed,minimum height=1cm,minimum width=1cm]
-\tikzstyle{smallblock} = [draw,rectangle,minimum height=0.5cm,minimum width=0.5cm,fill= blue!30, font =  \scriptsize]
+\tikzstyle{smallblock} = [draw,rectangle,minimum height=0.5cm,minimum width=0.5cm,fill= green!30, font =  \scriptsize]
 \tikzstyle{smallcircle} = [draw,ellipse,minimum height=0.1cm,minimum width=0.3cm,fill= yellow!40, font =  \scriptsize ]
@@
 
-\tikzset{>=latex}
-
-\begin{tikzpicture}[scale=1, blocka/.style ={rectangle,text width=0.6cm,text height=0.5cm, outer sep=0pt}]
+\begin{tikzpicture}[scale=1, blocka/.style ={rectangle,text width=0.9cm,text height=0.6cm, outer sep=0pt}]
  \small
@@
     % node placement with matrix library: 5x4 array
-    \matrix(M)[ampersand replacement=\&, row sep=0.25cm, column sep=0.25cm] {
+    \matrix(M)[ampersand replacement=\&, row sep=2.0cm, column sep=10pt] {
 
-    \node[smallblock, align=center, fill= green!30] (P1) {Plant {1}};\&
-
-		\node[smallblock, align=center] (S1) {Sensor {1}};\&
-		\&
-		\&
-		\node[smallblock, align=center, fill= green!30] (P2) {Plant {2}};\&
-
-		\node[smallblock, align=center] (S2) {Sensor {2}};\&
-
+    %\&
+    \node[smallblock, align=center] (CS1) {Control \\ System {1}};\&\&
+    \node[smallblock, align=center] (CS2) {Control \\ System {2}};\&\&\&
 %    \&
-    %\node(d0) {$\cdots$};\&
+    \node(d1) {$\cdots$};\&
 %    \&
-
+    \node[smallblock, align=center] (CSm) {Control \\ System \textit{m}};\&
     \\
-
-    %\node(d1) {$\cdots$};\&
-%    \&
-
+    %
+    \node[blocka] (R1) {};\&\&
+    \node[blocka] (R2) {};\&\&\&
+%    \node[smallcircle] (R2) {R2};\&
+    \node[blocka] (d3) {};\&
+    \node[blocka] (Rm) {};\&
     \\
-		\\
-		\\
-		\\
-		\\
-		\\
-		\\
-
-		\&
-		\node[smallblock, align=center] (C1) {Controller {1}};\&
-		\&
-		\&
-		\&
-		\node[smallblock, align=center] (C2) {Controller {2}};\&
-
-		\\
-    %%
-    %\node[blocka] (R1) {};\&
-%%    \node[smallcircle] (R2) {R2};\&
-    %%\node[blocka] (d3) {};\&
-    %\node[blocka] (R2) {};\&
-    %\\
     };
 
-    %\node[block] (outer) [fit=(R1.north west)  (R2.south east)] {};
-    %
-    %\node[align=center, font =  \scriptsize] at (outer.center) {Access Point/ \\Controller};
+    \node[block] (outer) [fit=(R1.north west) (d3) (Rm.south east)] {};
 
-    \draw [->,  red] (S1.south) -- node[left, pos = 0.1]{ $\alpha_1$ } (C1.north);
+    \node[align=center, scale =0.9] at (outer.center) {Access Point/ \\Controller};
+
+    \draw [->, thick, red] (CS1) -- node[left]{} (R1);
+    \draw [->, thick, red] (CS2) -- node[left]{} (R2);
 %    \draw [->, thick, magenta] (T2) -- node[left]{ \scriptsize $h_2$} (R2);
-    \draw [->, red] (S2.south) -- node[right, pos = 0.1]{ $\alpha_2$ } (C2.north);
+    \draw [->, thick, red] (CSm) -- node[left]{} (Rm);
 %
 
-		\draw [->, dashed,  red] (S1.south) -- node[right, pos=0.2]{ } (C2.north);
-
-		\draw [->, dashed, red] (S2.south) -- node[left, pos=0.2]{ } (C1.north);
-
-		%\begin{pgfonlayer}{bg0}    % select the background layer
-		%\draw [->, black] (C1.west) |- ($(C1) + (-25pt, 0pt)$) node(down_right){} 
-		%-- ($(P1) + (-25pt,+0pt)$) node(up_right){} -- (P1);
-		%\end{pgfonlayer}
+		\begin{pgfonlayer}{bg0}    % select the background layer
+		\draw [->, dashed, black] (R1) |- ($(R1) + (+35pt,-20pt)$) node(down_right){} 
+		-- ($(CS1) + (+35pt,+20pt)$) node(up_right){} -| (CS1);
+		\end{pgfonlayer}
 
 		\begin{pgfonlayer}{bg0}    % select the background layer
-		\draw [->, black] (C1.west) -|   (P1.south);
-		\end{pgfonlayer}
-
-		%\begin{pgfonlayer}{bg0}
-		%\draw [->,  black] (R2) |- ($(R2) + (-25pt,-15pt)$) node(down_right){} 
-		%--($(P2) + (-25pt,+0pt)$) node(up_right){} -- (P2);
-		%\end{pgfonlayer}
-
-		\begin{pgfonlayer}{bg0}    % select the background layer
-		\draw [->, black] (C2.west) -|   (P2.south);
+		\draw [->, dashed, black] (R2) |- ($(R2) + (+35pt,-20pt)$) node(down_right){} 
+		-- ($(CS2) + (+35pt,+20pt)$) node(up_right){} -| (CS2);
 		\end{pgfonlayer}
@@
 		\begin{pgfonlayer}{bg0}
-		\draw [->, black] (P1) -- (S1);
-		\end{pgfonlayer}
-
-		\begin{pgfonlayer}{bg0}
-		\draw [->,  black] (P2) -- (S2);
+		\draw [->, dashed, black] (Rm) |- ($(Rm) + (+35pt,-20pt)$) node(down_right){} 
+		--($(CSm) + (+35pt,+20pt)$) node(up_right){} -| (CSm);
 		\end{pgfonlayer}
@@
 		%\begin{scope}[on background layer]
-		\node(shared) [fill=red!10, fit={
-		($(S1.south) + (-32pt, -12pt)$) 
-		($(S2.south) + (+8pt, -12pt)$)
-		($(C1.north) + (-32pt, +10pt)$)
-		($(C2.north) + (+8pt, +10pt)$)
+		\node(shared) [fill=red!10, fit={($(CS1.south) + (-15pt, -10pt)$) 
+		($(CS2.south) + (-10pt, -10pt)$)
+		($(CSm.south) + (+20pt, -10pt)$)
+		($(R1.north) + (-15pt, +10pt)$)
+		($(R2.north) + (-10pt, +10pt)$)
+		($(Rm.north) + (+20pt, +10pt)$)
 		}] {};
@@
 
-		\node[align=center, red!40, font =  \footnotesize](shared_medium) at ($(shared.center)-(50pt, 0)$) {Shared \\ Wireless \\ Medium};
+		\node[align=center, red!50](shared_medium) at (shared.center) {Shared \\ Wireless \\ Medium};
 
@@
 	\pgfresetboundingbox
-   \useasboundingbox ($(FIRST SW) + (+52pt,+3pt)$) rectangle ($(FIRST NE) + (-10pt,-2pt)$);
+   \useasboundingbox ($(FIRST SW) + (+30pt,0)$) rectangle (FIRST NE);
 
 \end{tikzpicture}
-\end{document}

### 2019-04-02 control_system.py
Add files via upload

@@ -42,7 +42,17 @@ class ProbabilisticSystem(object):
         samples2 = np.random.uniform(-bound, bound, size=(batch_size, self.control_state_dim))
         return np.hstack((samples,samples2))
 
+    def _exponential_constant_sample(self, batch_size, state):
+        samples = np.random.exponential(self.mu, size=(batch_size, self.channel_state_dim))
+        samples2 = np.tile(state,[batch_size, 1])
+        return np.hstack((samples,samples2))
 
+
+#######################################################################################3
+########################################################################################
+################## W I F I (CC) ##############################################################
+##########################################################################################
+###########################################################################################
 class WirelessSchedulingSystem_CC(ProbabilisticSystem):
     def __init__(self, num_users, p=1, Ao=None, Ac=None, W=None, rho = .95, mu=2, S=1, num_rus = 25,num_channels=9):
         super().__init__(num_users*num_rus + num_users*p, num_users, S*num_channels)
@@ -241,258 +251,55 @@ class WirelessSchedulingSystem_CC(ProbabilisticSystem):
 
         return lhs
 
+
+#######################################################################################3
+########################################################################################
+################## W I F I (LC) ##############################################################
+##########################################################################################
+###########################################################################################
 class WirelessSchedulingSystem_LC(ProbabilisticSystem):
-    def __init__(self, num_users, p=1, Ao=None, Ac=None, W=None, tmax=0.001, mu=2, S=1, num_rus = 25,num_channels=9):
-        super().__init__(num_users*num_rus + num_users*p, num_users, S*num_channels)
[…]

### 2019-04-02 main.py

@@ -6,8 +6,10 @@ import scipy.io
 from scipy import sparse
 
 from control_system import *
-from reinforce_policy_T import ReinforcePolicy
-from reinforce_policy_T import *
+from reinforce_policy import ReinforcePolicy
+from reinforce_policy import *
+
+
 
 #######################
 ## UTILITY FUNCTIONS ##
@@ -26,7 +28,7 @@ def moving_average(data, window=10):
 #########################################################################
 ######################## RUN SIMULATION ################################
 ########################################################################
-def run_sys(sys, policy, num_iter_train, num_iter_exec, batch_size=64):
+def run_sys(sys, policy, num_iter_train, num_iter_exec, batch_size=64, explore_decay=0.999):
 
 
 
@@ -37,7 +39,9 @@ def run_sys(sys, policy, num_iter_train, num_iter_exec, batch_size=64):
                    'p': [],
                    'Ao': [],
                    'Ac': [],
-                   'x': []}
+                   'x': [],
+                   'tt': [],
+                   'g': []}
     runtime_dict = {'f0': [],
                    'f0_2': [],
                    'f0_3': [],
@@ -47,6 +51,7 @@ def run_sys(sys, policy, num_iter_train, num_iter_exec, batch_size=64):
                    'p3': [],
                    'p4': [],
                    'T': [],
+                   'T2': [],
                    'x': [],
                    'x2': [],
                    'x3': [],
@@ -62,6 +67,7 @@ def run_sys(sys, policy, num_iter_train, num_iter_exec, batch_size=64):
                    'p3': [],
                    'p4': [],
                    'T': [],
+                   'T2': [],
                    'x': [],
                    'x2': [],
                    'x3': [],
@@ -80,30 +86,50 @@ def run_sys(sys, policy, num_iter_train, num_iter_exec, batch_size=64):
     ##### T R A I N I N G #####################################
     ########################################################3##
     print("::::::::TRAINING PHASE:::::::")
+    eps = 1.0
     for k in range(num_iter_train):
         state0 = sys.sample(batch_size)
 
         if k%1000 == 0:
             print("Iteration " + str(k))
 
-        states, actions = policy.get_action(state0) # generate policy rollout over time horizon
[…]

### 2019-04-02 reinforce_policy.py

@@ -1,5 +1,6 @@
 import numpy as np
 import tensorflow as tf
+from tensorflow import keras
 import math
 import scipy
 import scipy.stats
@@ -89,81 +90,28 @@ class ProbabilityAction(object):
         raise Exception("Not Implemented")
 
 
-class GammaDistribution2(ProbabilityAction):
-    """
-    Gamma Distribution with fixed alpha=2
-    """
-    def __init__(self, action_dim, lower_bound, upper_bound):
-        self.alpha = 2
-        super().__init__(1, action_dim)
-
-    def log_prob(self, params, selected_action):
-        beta = tf.nn.sigmoid(params) * 4
-
-        log_probs = self.alpha * tf.log(beta) + \
-                (self.alpha - 1) * tf.log(selected_action) - \
-                beta * selected_action - \
-                np.log(scipy.special.gamma(self.alpha))
-        log_probs = tf.reduce_sum(log_probs, axis=1)
-
-        return log_probs, beta
-
-    def get_action(self, params):
-        # params is beta
-        action = np.random.gamma(shape=self.alpha, scale=1.0/params)
-        return action
-
-
-class GaussianDistribution(ProbabilityAction):
-    #TODO: FIX THIS
-    def __init__(self, action_dim, lower_bound, upper_bound):
-        super().__init__(2, action_dim)
-
-    def log_prob(self, params, selected_action):
-        mean = tf.gather(params, np.array(range(self.action_dim)), axis=1, name='mean')
-        var = tf.gather(params, np.array(range(self.action_dim, 2*self.action_dim)), axis=1, name='var')
-
-        mean = tf.nn.sigmoid(mean) * 10
-        var = tf.nn.sigmoid(var) * 1
-        output = None
-
-        # TODO: this can create NaNs when self.var is small
-        log_probs = -0.5 * tf.log(2*np.pi*var) - tf.square(selected_action - mean) / (2. * var)
-        log_probs = tf.reduce_sum(log_probs, axis=1)
-
-        return log_probs, outputex
-
-    def get_action(self, params):
-        mean = params[:, :action_dim]
-        var = params[:, action_dim:]
-
-        print("mean: " + str(mean))
-        print("var: " + str(var))
-
-        action = np.random.normal(mean, var) # TODO: this is not actually var, its std
-        # # restrict actions to be at least some small positive value
-        # action[action < 0.05] = 0.05
-        return action
-
-class ClassificationDistribution(ProbabilityAction):
[…]
